### 2023-11-01 R_code/paper_analysis.Rmd
new changes

---
title: "R Notebook"
output: html_notebook
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.

```{r}
library(scales)
library(tidyverse)
library(caret)
library(ggpubr)
library(pROC)
```

```{r}
prefix = "/Users/noa/Workspace/bootstrap_repo/ML_pipeline"

fasttree_all_models_performance = read_tsv(paste(prefix,"/fasttree/all_models_performance.tsv", sep = ""))
fasttree_final_models_performance = read_tsv(paste(prefix,"/fasttree/final_model_performance.tsv", sep = ""))
fasttree_group_performance = read_tsv(paste(prefix,"/fasttree/groups_performance.tsv", sep = ""))
fasttree_vi =  read_tsv(paste(prefix,"/fasttree/model_standard_vi.tsv", sep = ""))
fasttree_enriched_test = read_tsv(paste(prefix,"/fasttree/test_model_standard_enriched.tsv", sep = ""))
#####
iqtree_all_models_performance = read_tsv(paste(prefix,"/iqtree/all_models_performance.tsv", sep = ""))
iqtree_final_models_performance = read_tsv(paste(prefix,"/iqtree/final_model_performance.tsv", sep = ""))
iqtree_group_performance = read_tsv(paste(prefix,"/iqtree/groups_performance.tsv", sep = ""))
iqtree_vi =  read_tsv(paste(prefix,"/iqtree/model_standard_vi.tsv", sep = ""))
iqtree_enriched_test = read_tsv(paste(prefix,"/iqtree/test_model_standard_enriched.tsv", sep = ""))
#####
raxml_all_models_performance = read_tsv(paste(prefix,"/raxml/all_models_performance.tsv", sep = ""))
raxml_final_models_performance = read_tsv(paste(prefix,"/raxml/final_model_performance.tsv", sep = ""))
raxml_group_performance = read_tsv(paste(prefix,"/raxml/groups_performance.tsv", sep = ""))
raxml_vi =  read_tsv(paste(prefix,"/raxml/model_standard_vi.tsv", sep = ""))
raxml_enriched_test = read_tsv(paste(prefix,"/raxml/test_model_standard_enriched.tsv", sep = ""))

```

Choose specific program
```{r}
program = 'raxml'
if (program=='fasttree') {
    all_models_performance = fasttree_all_models_performance
    final_models_performance = fasttree_final_models_performance
    group_performance = fasttree_group_performance
    vi = fasttree_vi
    enriched_test = fasttree_enriched_test

} else if (program=='iqtree') {
    all_models_performance = iqtree_all_models_performance
    final_models_performance = iqtree_final_models_performance
    group_performance = iqtree_group_performance
    vi = fasttree_vi
    enriched_test = iqtree_enriched_test
} else if (program=='raxml') {
    all_models_performance = raxml_all_models_performance
    final_models_performance = raxml_final_models_performance
    group_performance = raxml_group_performance
    vi = fasttree_vi
    enriched_test = raxml_enriched_test
}
```

General performance
```{r}
final_models_performance  %>% filter (dataset=='test', metric_type=='all_data') %>% dplyr::select (analysis_type,name,balanced_accuracy_score, mcc)
```

```{r}

```

Variable importance 
```{r}
vi %>% arrange (-`Gini-importance`)
```

Confusion matrix

```{r}
ggplotConfusionMatrix <- function(m){
  mytitle <- ""#paste("Accuracy", percent_format()(m$overall[1]))
                  # "Kappa", percent_format()(m$overall[2]))
  p <-
    ggplot(data = as.data.frame(m$table) ,
           aes(x = Reference, y = Prediction)) +
    geom_tile(aes(fill = log(Freq)), colour = "white") +
    scale_fill_gradient(low = "white", high = "steelblue") +
    geom_text(aes(x = Reference, y = Prediction, label = scales::comma(Freq))) +
    theme(legend.position = "none") +
    ggtitle(mytitle)+theme(text = element_text(size = 13))
  return(p)
}
conf_data<-enriched_test %>% dplyr::select (prob_predictions, true_label) %>% mutate (pred_status = prob_predictions>0.5)

obs<- factor(conf_data %>% pull (true_label))
pred<- factor(conf_data %>% pull (pred_status))

confusion_plot<-ggplotConfusionMatrix(confusionMatrix(data = pred, reference= obs))+theme(legend.position = "none")

```
Prediction metrics

```{r}
library(predtools)
library(MASS)
library(caret)

error_vs_size<- all_models_performance %>% filter (metric_type=='all_data', dataset =='test')
error_vs_size_plt = error_vs_size %>% mutate(n_MSAs = sample_frac) %>% ggplot(aes(x=n_MSAs, y = AUC))+geom_line()+ expand_limits(y = 0.5)+xlab("Number of MSAs")+ theme(text = element_text(size = 13))+ylab("AUC")+geom_point()

rocobj <- roc(enriched_test$true_label, enriched_test$prob_predictions)
auc <- round(auc(enriched_test$true_label, enriched_test$prob_predictions),3)

global_max_auc_plt<-ggroc(rocobj,colour = 'steelblue', size = 2)+
  ggtitle(paste0('AUC = ', auc))+theme(text = element_text(size = 13))+xlab("Specificity")

ggarrange(global_max_auc_plt,confusion_plot,error_vs_size_plt, labels = c("A","B","C"),align = "h", nrow = 2, ncol = 2 )

```
Group performance

```{r}
test_group_performance<- group_performance %>% filter (dataset=='test')

Pypythia_error<-test_group_performance %>% filter (grouping_col_name =='msa_difficulty_group') %>% ggplot(aes(y=grouping_col, x=Balanced_Accuracy))+geom_col()+ expand_limits(y = c(0.5,1))+ylab('MSA\ndifficulty')+ xlab('Balanced\naccuracy')+theme(text = element_text(size = 11))

n_seq_error<-test_group_performance %>% filter (grouping_col_name =='n_seq_group')  %>% ggplot(aes(y=grouping_col, x=Balanced_Accuracy))+geom_col()+ expand_limits(y = c(0.5,1))+ylab('Number\nof sequences')+xlab('Balanced\naccuracy')+theme(text = element_text(size = 11))

#>%  mutate(grouping_col = factor(grouping_col, levels = c('(99.999, 219.0]','(219.0, 546.0]','(546.0, 1519.0]','(1519.0, 16115.0]')))

n_loci_error<-test_group_performance %>% filter (grouping_col_name =='feature_msa_n_loci') %>% ggplot(aes(y=grouping_col, x=Balanced_Accuracy))+geom_col()+ expand_limits(y = c(0.5,1))+ylab('Number\nof positions')+xlab('Balanced\naccuracy')+theme(text = element_text(size = 11))

Pypythia_scatter<-enriched_test %>% dplyr::select (msa_path,prob_predictions,  feature_msa_pypythia_msa_difficulty) %>% group_by(msa_path,feature_msa_pypythia_msa_difficulty) %>% summarise(mean_prob = mean(prob_predictions)) %>% ggplot(aes(x=feature_msa_pypythia_msa_difficulty, y = mean_prob))+geom_point(size = 0.7)+xlab("MSA\ndifficulty")+ylab("Predicted\nprobability")+theme(text = element_text(size = 11))

seq_scatter<-enriched_test %>% dplyr::select (msa_path,prob_predictions, feature_msa_n_seq, feature_msa_pypythia_msa_difficulty) %>% group_by(msa_path,feature_msa_n_seq) %>% summarise(mean_prob = mean(prob_predictions)) %>% ggplot(aes(x=feature_msa_n_seq, y = mean_prob))+geom_point(size = 0.7)+xlab("Number of\nsequences")+ylab("Predicted\nprobability")+theme(text = element_text(size = 11))

n_loci_scatter<-enriched_test %>% dplyr::select (msa_path,prob_predictions, feature_msa_n_loci) %>% group_by(msa_path,feature_msa_n_loci) %>% summarise(mean_prob = mean(prob_predictions)) %>% ggplot(aes(x=feature_msa_n_loci, y = mean_prob))+geom_point(size = 0.7)+xlab("Number of\n positions")+ylab("Predicted\nprobability")+theme(text = element_text(size = 11))

ggarrange(seq_scatter,Pypythia_scatter,n_loci_scatter, n_seq_error,Pypythia_error, n_loci_error,labels = c("A","C","E","B","D","F"),align = "h", nrow = 3, ncol = 3, legend = "bottom",vjust= 1)
#error_vs_group_5_5_including_plt

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

### 2023-11-01 ML_pipeline/ML_boot_edit.py

@@ -139,26 +139,17 @@ def main():
     parser.add_argument('--working_dir', type = str, default = os.getcwd())
     parser.add_argument('--cpus_per_main_job', type = int, default=1)
     parser.add_argument('--sample_fracs', type = str, default='0.5_1')
-    parser.add_argument('--main_data_path',type = str, default = '/Users/noa/Workspace/bootstrap_results/job_raw_data_with_features.tsv')
-    parser.add_argument('--validation_data_path_raxml',type = str, default ='/Users/noa/Workspace/bootstrap_results/simulations_df_raxml.tsv')
-    parser.add_argument('--validation_data_path_iqtree', type=str,
-                        default='/Users/noa/Workspace/bootstrap_results/simulations_df_iqtree.tsv')
-    parser.add_argument('--validation_data_path_fasttree', type=str,
-                        default='/Users/noa/Workspace/bootstrap_results/simulations_df_fasttree.tsv')
+    parser.add_argument('--main_data_folder',type = str, default = '/Users/noa/Workspace/bootstrap_results/job_raw_data_with_features.tsv')
+    parser.add_argument('--validation_data_folder', type=str,
+                        default='/Users/noa/Workspace/bootstrap_results/job_raw_data_with_features.tsv')
+
     args = parser.parse_args()
-    full_data = pd.read_csv(args.main_data_path, sep = '\t').sample(n = 2000)
 
-    validation_raxml = pd.read_csv(args.validation_data_path_raxml, sep = '\t').sample(n=200)
-    transform_data(validation_raxml)
-    validation_iqtree = pd.read_csv(args.validation_data_path_iqtree, sep='\t').sample(n=200)
-    transform_data(validation_iqtree)
-    validation_fasttree = pd.read_csv(args.validation_data_path_fasttree, sep='\t').sample(n=200)
-    transform_data(validation_fasttree)
-    validation_dict = {'raxml': {'val':validation_raxml}, 'iqtree': {'val':validation_iqtree}, 'fasttree': {'val':validation_fasttree}}
-
-    transform_data(full_data)
-    for program in full_data['program'].unique():
-        program_data = full_data.loc[full_data.program == program]
+    for program in ['iqtree','fasttree','raxml']:
+        program_data = pd.read_csv(os.path.join(args.main_data_folder,f'simulation_df_{program}.tsv'))
+        transform_data(program_data)
[…]

### 2023-11-01 ML_pipeline/ML_boot_edit.py
new defenition changes

@@ -1,3 +1,12 @@
+import sys
+if sys.platform == "linux" or sys.platform == "linux2":
+    PROJECT_ROOT_DIRECRTORY = "/groups/pupko/noaeker/bootstrap_repo"
+else:
+    PROJECT_ROOT_DIRECRTORY = "/Users/noa/Workspace/bootstrap_repo"
+sys.path.append(PROJECT_ROOT_DIRECRTORY)
+
+
+
 from ML_utils.ML_algorithms_and_hueristics import *
 from side_code.file_handling import create_or_clean_dir, create_dir_if_not_exists
 from side_code.MSA_manipulation import get_MSA_seq_names
@@ -133,23 +142,22 @@ def transform_data(df):
     df.drop(columns = ['feature_n_unique_seq'], inplace= True)
     #+[col for col in df.columns if 'msa_entropy' in col]
 
-
 def main():
     parser = argparse.ArgumentParser()
     parser.add_argument('--working_dir', type = str, default = os.getcwd())
+    parser.add_argument('--RFE', action='store_true', default= False)
+    parser.add_argument('--full_grid', action='store_true', default=False)
     parser.add_argument('--cpus_per_main_job', type = int, default=1)
     parser.add_argument('--sample_fracs', type = str, default='0.5_1')
     parser.add_argument('--main_data_folder',type = str, default = '/Users/noa/Workspace/bootstrap_results/job_raw_data_with_features.tsv')
     parser.add_argument('--validation_data_folder', type=str,
                         default='/Users/noa/Workspace/bootstrap_results/job_raw_data_with_features.tsv')
-
     args = parser.parse_args()
-
     for program in ['iqtree','fasttree','raxml']:
         program_data = pd.read_csv(os.path.join(args.main_data_folder,f'simulation_df_{program}.tsv'))
         transform_data(program_data)
-        validation_data = pd.read_csv(os.path.join(args.validation_data_folder,f'simulation_df_{program}.tsv'))
-        transform_data(validation_data)
+        program_validation_data = pd.read_csv(os.path.join(args.validation_data_folder,f'simulation_df_{program}.tsv'))
+        transform_data(program_validation_data)
         working_dir = os.path.join(args.working_dir, program)
         create_dir_if_not_exists(working_dir)
         bootstrap_cols = get_bootstrap_col(program)
@@ -158,11 +166,11 @@ def main():
         sample_fracs = [float(frac) for frac in (args.sample_fracs).split('_')]
[…]

### 2023-11-01 ML_pipeline/ML_boot_edit.py
new changes

@@ -153,8 +153,11 @@ def main():
     parser.add_argument('--validation_data_folder', type=str,
                         default='/Users/noa/Workspace/bootstrap_results/job_raw_data_with_features.tsv')
     args = parser.parse_args()
+    log_file_path = os.path.join(args.working_dir, "ML.log")
+    logging.basicConfig(filename=log_file_path, level=logging.DEBUG)
     for program in ['iqtree','fasttree','raxml']:
-        program_data = pd.read_csv(os.path.join(args.main_data_folder,f'simulation_df_{program}.tsv'))
+        logging.info(f"Program = {program}")
+        program_data = pd.read_csv(os.path.join(args.main_data_folder,f'simulations_df_{program}.tsv'))
         transform_data(program_data)
         program_validation_data = pd.read_csv(os.path.join(args.validation_data_folder,f'simulation_df_{program}.tsv'))
         transform_data(program_validation_data)
@@ -166,9 +169,11 @@ def main():
         sample_fracs = [float(frac) for frac in (args.sample_fracs).split('_')]
         all_model_merics = pd.DataFrame()
         for sample_frac in sample_fracs:
+            logging.info(f"Sample frac = {sample_frac}")
             curr_model_metrics, groups_analysis = ML_pipeline(program_data, bootstrap_cols, args.cpus_per_main_job, working_dir, sample_frac,compare_to_bootstrap_models= False,subsample_train = True,do_RFE = args.RFE, large_grid = False, name = f"frac_{sample_frac}", validation_dict = validation_dict[program])
             all_model_merics = pd.concat([all_model_merics,curr_model_metrics])
         all_model_merics.to_csv(os.path.join(working_dir, 'all_models_performance.tsv'), sep=CSV_SEP)
+        logging.info(f"Generating optimized final model")
         final_model_metrics,groups_analysis = ML_pipeline(program_data, bootstrap_cols, args.cpus_per_main_job, working_dir, sample_frac = -1,subsample_train = False, do_RFE=args.RFE,
                     large_grid=args.full_grid, name=f"final_model", validation_data = program_validation_data, compare_to_bootstrap_models= True, extract_predictions = True)
         final_model_metrics.to_csv(os.path.join(working_dir, 'final_model_performance.tsv'), sep=CSV_SEP)

### 2023-11-02 ML_pipeline/ML_boot_edit.py
adding logging and fixing \t

@@ -110,6 +110,7 @@ def ML_pipeline(program_data, bootstrap_cols, cpus_per_main_job, working_dir, sa
     y_test = test["true_binary_support"]
 
     groups = train["tree_id"]
+    logging.info("Training ML model")
     model = ML_model(X_train, groups, y_train, n_jobs=cpus_per_main_job, path=os.path.join(working_dir, f'model_{sample_frac}'),
                      classifier=True, model='lightgbm', calibrate=True, name=name, large_grid=large_grid, do_RFE=do_RFE,
                      n_cv_folds=3)
@@ -119,11 +120,13 @@ def ML_pipeline(program_data, bootstrap_cols, cpus_per_main_job, working_dir, sa
         validation_data = validation_dict[val_name]
         data_dict[val_name]={'X': validation_data[[col for col in validation_data.columns if col in features]],'y': validation_data["true_binary_support"], 'full_data': validation_data}
     all_models_performance = pd.DataFrame()
+    logging.info("Evaluating model performance")
     model_performance,group_performance = overall_model_performance_analysis(working_dir, model,data_dict, name=f"model_standard",extract_predictions = extract_predictions)
     model_performance["analysis_type"] = "standard"
     all_models_performance = pd.concat([all_models_performance, model_performance])
 
     if compare_to_bootstrap_models:
+        logging.info("Comparing to bootstrap models")
         for bootstrap_col in bootstrap_cols:
             bootstrap_models_performance = bootstrap_model_pipeline(working_dir,train, y_train,test, y_test,features,bootstrap_col,groups,cpus_per_main_job,sample_frac,do_RFE , large_grid, name)
             all_models_performance = pd.concat([all_models_performance, bootstrap_models_performance])
@@ -157,9 +160,9 @@ def main():
     logging.basicConfig(filename=log_file_path, level=logging.DEBUG)
     for program in ['iqtree','fasttree','raxml']:
         logging.info(f"Program = {program}")
-        program_data = pd.read_csv(os.path.join(args.main_data_folder,f'simulations_df_{program}.tsv'))
+        program_data = pd.read_csv(os.path.join(args.main_data_folder,f'simulations_df_{program}.tsv'),sep='\t')
         transform_data(program_data)
-        program_validation_data = pd.read_csv(os.path.join(args.validation_data_folder,f'simulation_df_{program}.tsv'))
[…]

### 2023-11-02 ML_pipeline/ML_boot_edit.py
small fix

@@ -162,7 +162,7 @@ def main():
         logging.info(f"Program = {program}")
         program_data = pd.read_csv(os.path.join(args.main_data_folder,f'simulations_df_{program}.tsv'),sep='\t')
         transform_data(program_data)
-        program_validation_data = pd.read_csv(os.path.join(args.validation_data_folder,f'simulation_df_{program}.tsv'),sep='\t')
+        program_validation_data = pd.read_csv(os.path.join(args.validation_data_folder,f'simulations_df_{program}.tsv'),sep='\t')
         transform_data(program_validation_data)
         working_dir = os.path.join(args.working_dir, program)
         create_dir_if_not_exists(working_dir)

### 2023-11-02 ML_pipeline/ML_boot_edit.py
fix

@@ -171,14 +171,15 @@ def main():
         program_data = program_data.dropna(axis=0)
         sample_fracs = [float(frac) for frac in (args.sample_fracs).split('_')]
         all_model_merics = pd.DataFrame()
+        validation_dict = {'val': program_validation_data}
         for sample_frac in sample_fracs:
             logging.info(f"\n#Sample frac = {sample_frac}")
-            curr_model_metrics, groups_analysis = ML_pipeline(program_data, bootstrap_cols, args.cpus_per_main_job, working_dir, sample_frac,compare_to_bootstrap_models= False,subsample_train = True,do_RFE = args.RFE, large_grid = False, name = f"frac_{sample_frac}", validation_dict = validation_dict[program])
+            curr_model_metrics, groups_analysis = ML_pipeline(program_data, bootstrap_cols, args.cpus_per_main_job, working_dir, sample_frac,compare_to_bootstrap_models= False,subsample_train = True,do_RFE = args.RFE, large_grid = False, name = f"frac_{sample_frac}", validation_dict = validation_dict)
             all_model_merics = pd.concat([all_model_merics,curr_model_metrics])
         all_model_merics.to_csv(os.path.join(working_dir, 'all_models_performance.tsv'), sep=CSV_SEP)
         logging.info(f"Generating optimized final model")
         final_model_metrics,groups_analysis = ML_pipeline(program_data, bootstrap_cols, args.cpus_per_main_job, working_dir, sample_frac = -1,subsample_train = False, do_RFE=args.RFE,
-                    large_grid=args.full_grid, name=f"final_model", validation_data = program_validation_data, compare_to_bootstrap_models= True, extract_predictions = True)
+                    large_grid=args.full_grid, name=f"final_model", validation_dict = validation_dict, compare_to_bootstrap_models= True, extract_predictions = True)
         final_model_metrics.to_csv(os.path.join(working_dir, 'final_model_performance.tsv'), sep=CSV_SEP)
         groups_analysis.to_csv(os.path.join(working_dir, 'groups_performance.tsv'), sep=CSV_SEP)
 

### 2023-11-02 ML_utils/ML_algorithms_and_hueristics.py
small fix

@@ -175,8 +175,8 @@ def overall_model_performance_analysis(working_dir,model, data_dict, name,extrac
     all_group_metrics = pd.DataFrame()
     for dataset in data_dict:
         if model:
-            prob_predictions = model['best_model'].predict_proba(data_dict[dataset]["X"])[:, 1]
-            predictions = model['best_model'].predict(data_dict[dataset]["X"])
+            prob_predictions = model['best_model'].predict_proba((model['selector']).transform(data_dict[dataset]["X"]))[:, 1]
+            predictions = model['best_model'].predict((model['selector']).transform(data_dict[dataset]["X"]))
         else:
             raw_bootstrap_values = data_dict[dataset]["X"].drop(['ignore'],axis=1).iloc[:,0]
             prob_predictions = raw_bootstrap_values / 100

### 2023-11-04 R_code/paper_analysis.Rmd
new changes

@@
 ```
+```{r}
+validation_data_prefix = "/Users/noa/Workspace/bootstrap_results/validation_data"
+fasttree_validation_data = read_tsv(paste(validation_data_prefix,"/simulations_df_fasttree.tsv", sep = ""))
+raxml_validation_data = read_tsv(paste(validation_data_prefix,"/simulations_df_raxml.tsv", sep = ""))
+iqtree_validation_data = read_tsv(paste(validation_data_prefix,"/simulations_df_iqtree.tsv", sep = ""))
+```
 
@@
 fasttree_enriched_test = read_tsv(paste(prefix,"/fasttree/test_model_standard_enriched.tsv", sep = ""))
+validation_data_fasttree = 
 #####
@@
 ```{r}
-program = 'raxml'
+program = 'fasttree'
 if (program=='fasttree') {
@@
     enriched_test = fasttree_enriched_test
+    validation_data = fasttree_validation_data
 
@@
     enriched_test = iqtree_enriched_test
+    validation_data = iqtree_validation_data
 } else if (program=='raxml') {
@@
     enriched_test = raxml_enriched_test
+    validation_data = raxnk_validation_data
 }
@@
 ```
+Validation_data
+```{r}
+validation_data
+```
 

### 2023-11-04 simulation_edit/bootstrap_edit.py

@@ -323,18 +323,21 @@ def msa_path_bootstrap_analysis(msa_path,curr_run_dir, mle_path, true_tree_path,
             statistics["feature_n_unique_seq"] = len(get_MSA_seq_names(curr_pruned_msa_path))
             nni_neighbors = get_nni_neighbors(mle_with_internal_path, node.name)
             neighbors_per_site_ll = []
-            for neighbor in  nni_neighbors:
+            st_ll = time.time()
+            for neighbor in nni_neighbors:
                 neighbor.write(format=1, outfile=neighbors_tmp_path)
                 neighbor = Tree(neighbors_tmp_path, format=1)
                 curr_pruned_tree_path, curr_pruned_msa_path = get_pruned_tree_and_msa(curr_run_dir, msa_path,
                                                                                       bootstrap_tree_details_dict[
                                                                                           "model_short"], neighbor)
 
-                st_ll = time.time()
                 neighbor_ll = raxml_compute_tree_per_site_ll(garbage_dir, curr_pruned_msa_path, curr_pruned_tree_path, ll_on_data_prefix = "nni_neighbors", model = bootstrap_tree_details_dict["model_short"], opt = True)
-                end_ll = time.time()
-                neig_ll_evaluation_time+= end_ll-st_ll
+
                 neighbors_per_site_ll.append(np.array(neighbor_ll))
+
+            end_ll = time.time()
+            statistics["partition_ll_eval_time_test"] = end_ll-st_ll
+            neig_ll_evaluation_time += end_ll - st_ll
             orig_tree_ll = np.sum(per_site_original_tree_ll)
             nei1_ll = np.sum(neighbors_per_site_ll[0])
             nei2_ll = np.sum(neighbors_per_site_ll[1])
@@ -347,10 +350,11 @@ def msa_path_bootstrap_analysis(msa_path,curr_run_dir, mle_path, true_tree_path,
                               }
             statistics.update(nni_statistics)
             statistics.update(bootstrap_tree_details_dict)
-            msa_splits = msa_splits.append(statistics, ignore_index= True)
+            msa_splits = msa_splits.append(statistics, ignore_index=True)
             create_or_clean_dir(garbage_dir)
 
-    return msa_splits,feature_extraction_time,neig_ll_evaluation_time
+    logging.info(f"Feature extraction time = {feature_extraction_time} total neig evaluation time = {neig_ll_evaluation_time}")
+    return msa_splits, feature_extraction_time, neig_ll_evaluation_time
 
 
 def main():

### 2023-11-04 simulations_generation/simulations_argparser.py

@@ -17,8 +17,8 @@ def main_parser():
     parser.add_argument('--msa_type', type = str, default= 'DNA')
     parser.add_argument('--nb_iqtree',type = int, default= 1000)
     parser.add_argument('--nb_fasttree', type=int, default=10)
-    parser.add_argument('--calc_features', action='store_true', default=False)
-    parser.add_argument('--pre_chosen_tree_ids',action='store_true', default= False)
+    parser.add_argument('--calc_features', action='store_true', default=True)
+    parser.add_argument('--pre_chosen_tree_ids',action='store_true', default= True)
     parser.add_argument('--n_k_means_clusters',type=int, default =3)
     return parser
 

### 2023-11-06 R_code/paper_analysis.Rmd
new changes to ML pipeline

--- 2023-11-01 R_code/paper_analysis.Rmd
@@
 ```
+```{r}
+validation_data_prefix = "/Users/noa/Workspace/bootstrap_results/validation_data"
+fasttree_validation_data = read_tsv(paste(validation_data_prefix,"/simulations_df_fasttree.tsv", sep = ""))
+raxml_validation_data = read_tsv(paste(validation_data_prefix,"/simulations_df_raxml.tsv", sep = ""))
+iqtree_validation_data = read_tsv(paste(validation_data_prefix,"/simulations_df_iqtree.tsv", sep = ""))
+```
 
 ```{r}
-prefix = "/Users/noa/Workspace/bootstrap_repo/ML_pipeline"
+prefix = "/Users/noa/Workspace/ML_results"
 
@@
 ```{r}
-program = 'raxml'
+program = 'iqtree'
 if (program=='fasttree') {
@@
     enriched_test = raxml_enriched_test
+    validation_data = raxml_validation_data
 }
@@
 ```{r}
-final_models_performance  %>% filter (dataset=='test', metric_type=='all_data') %>% dplyr::select (analysis_type,name,balanced_accuracy_score, mcc)
-```
-
-```{r}
-
+final_models_performance  %>% filter (dataset=='test', metric_type=='all_data') %>% dplyr::select (analysis_type,name,balanced_accuracy_score, mcc, AUC)
 ```
@@
 ```
+Validation_data
+```{r}
+validation_data %>% dplyr::select (boot_run_time, extraction_running_time,nni_ll_eval_running_time)
+
+```
 

### 2023-11-06 ML_pipeline/ML_boot_edit.py

@@ -68,12 +68,12 @@ def bootstrap_model_pipeline(working_dir,train, y_train,test, y_test,features,bo
     X_train_inc_boot = train[[col for col in train.columns if col in features] + [bootstrap_col]]
     X_test_inc_boot = test[[col for col in train.columns if col in features] + [bootstrap_col]]
     model_only_boot = ML_model(X_train_only_boot, groups, y_train, n_jobs=cpus_per_main_job,
-                               path=os.path.join(working_dir, f'model_{sample_frac}_only_boot'),
+                               path=os.path.join(working_dir, f'model_only_boot'),
                                classifier=True, model='lightgbm', calibrate=True, name=name,
                                large_grid=large_grid, do_RFE=do_RFE,
                                n_cv_folds=3)
     model_inc_boot = ML_model(X_train_inc_boot, groups, y_train, n_jobs=cpus_per_main_job,
-                              path=os.path.join(working_dir, f'model_{sample_frac}_inc_boot'
+                              path=os.path.join(working_dir, f'model_inc_boot'
                               f''),
                               classifier=True, model='lightgbm', calibrate=True, name=name, large_grid=large_grid,
                               do_RFE=do_RFE,
@@ -94,46 +94,77 @@ def bootstrap_model_pipeline(working_dir,train, y_train,test, y_test,features,bo
     return bootstrap_models_performance
 
 
+
+def standard_model_pipeline(train,test, y_train, y_test, groups,features, cpus_per_main_job, working_dir,do_RFE , large_grid, name, validation_dict,extract_predictions):
+    X_train = train[[col for col in train.columns if col in features]]
+    X_test = test[[col for col in train.columns if col in features]]
+
+    logging.info("Training ML model")
+    model = ML_model(X_train, groups, y_train, n_jobs=cpus_per_main_job,
+                     path=os.path.join(working_dir, f'model_stadard'),
+                     classifier=True, model='lightgbm', calibrate=True, name=name, large_grid=large_grid, do_RFE=do_RFE,
+                     n_cv_folds=3)
+
+    data_dict = {'test': {'X': X_test, 'y': y_test, 'full_data': test},
+                 'train': {'X': X_train, 'y': y_train, 'full_data': train}}
+    for val_name in validation_dict:
+        validation_data = validation_dict[val_name]
+        data_dict[val_name] = {'X': validation_data[[col for col in validation_data.columns if col in features]],
[…]

### 2023-11-06 programs/FastTree.py

@@ -51,7 +51,8 @@ def fasttree_pipeline(curr_run_dir, results_folder, msa_path, msa_type, nb):
     with open(bootstrap_trees_file,'w') as BF:
         BF.write("\n".join(bootstrap_trees))
     bootstrap_output_path = os.path.join(results_folder,'fasttree_standard_bootstrap')
-    get_booster_tree(output_tree_file, bootstrap_trees_file, out_path ="booster.nw")
+
+    get_booster_tree(output_tree_file, bootstrap_trees_file, out_path =bootstrap_output_path)
     return {'sh_bootstrap': output_tree_file, 'standard_bootstrap': bootstrap_output_path,'boot_run_time': bootsrtap_time}
 
 #t = fasttree_pipeline(curr_run_dir= os.path.join(os.getcwd(), 'trash'),msa_path="/Users/noa/Workspace/bootstrap_results/test/job_0/raxml_tree_0/25117/iqtree_msa_0/sim_msa.fa", msa_type='DNA', nb = 30)

### 2023-11-06 simulation_edit/bootstrap_edit.py

@@ -224,7 +224,7 @@ def get_bootstrap_and_tree_groups(program, bootstrap_tree_details,mle_path,garba
         add_internal_names(aBayes_ete)
         extra_boot = {'aLRT_iqtree': aLRT_ete, 'aBayes_iqtree': aBayes_ete}
     elif program == 'fasttree':
-        standard_bootstrap = bootstrap_tree_details['sh_bootstrap']
+        standard_bootstrap = bootstrap_tree_details['standard_bootstrap']
         standard_ete = Tree(newick=standard_bootstrap, format=0)
         add_internal_names(standard_ete)
         extra_boot = {'standard_fasttree_boot': standard_ete}

### 2023-11-06 simulations_generation/simulations_argparser.py

@@ -18,7 +18,7 @@ def main_parser():
     parser.add_argument('--nb_iqtree',type = int, default= 1000)
     parser.add_argument('--nb_fasttree', type=int, default=10)
     parser.add_argument('--calc_features', action='store_true', default=True)
-    parser.add_argument('--pre_chosen_tree_ids',action='store_true', default= True)
+    parser.add_argument('--pre_chosen_tree_ids',action='store_true', default= False)
     parser.add_argument('--n_k_means_clusters',type=int, default =3)
     return parser
 

### 2023-11-06 ML_pipeline/ML_boot_edit.py
fix

@@ -183,7 +183,7 @@ def main():
     parser.add_argument('--full_grid', action='store_true', default=False)
     parser.add_argument('--cpus_per_main_job', type = int, default=1)
     parser.add_argument('--sample_fracs', type = str, default='0.5_1')
-    parser.add_argument('--inc_sample_fracs', action='store_true', default=False)
+    parser.add_argument('--inc_sample_fracs', action='store_true', default=True)
     parser.add_argument('--main_data_folder',type = str, default = '/Users/noa/Workspace/bootstrap_results/remote_results')
     parser.add_argument('--validation_data_folder', type=str,
                         default='/Users/noa/Workspace/bootstrap_results/remote_results')
@@ -192,9 +192,9 @@ def main():
     logging.basicConfig(filename=log_file_path, level=logging.DEBUG)
     for program in ['fasttree']: #,'fasttree','raxml'
         logging.info(f"Program = {program}")
-        program_data = pd.read_csv(os.path.join(args.main_data_folder,f'simulations_df_{program}.tsv'),sep='\t').sample(frac=0.1)
+        program_data = pd.read_csv(os.path.join(args.main_data_folder,f'simulations_df_{program}.tsv'),sep='\t')
         transform_data(program_data)
-        program_validation_data = pd.read_csv(os.path.join(args.validation_data_folder,f'simulations_df_{program}.tsv'),sep='\t').sample(frac=0.1)
+        program_validation_data = pd.read_csv(os.path.join(args.validation_data_folder,f'simulations_df_{program}.tsv'),sep='\t')
         transform_data(program_validation_data)
         working_dir = os.path.join(args.working_dir, program)
         create_dir_if_not_exists(working_dir)

### 2023-11-06 ML_pipeline/ML_boot_edit.py
small fix

@@ -190,7 +190,7 @@ def main():
     args = parser.parse_args()
     log_file_path = os.path.join(args.working_dir, "ML.log")
     logging.basicConfig(filename=log_file_path, level=logging.DEBUG)
-    for program in ['fasttree']: #,'fasttree','raxml'
+    for program in ['fasttree','iqtree','raxml']: #,'fasttree','raxml'
         logging.info(f"Program = {program}")
         program_data = pd.read_csv(os.path.join(args.main_data_folder,f'simulations_df_{program}.tsv'),sep='\t')
         transform_data(program_data)

### 2023-11-06 ML_pipeline/ML_boot_edit.py
new fix

@@ -132,9 +132,9 @@ def ML_pipeline(program_data, bootstrap_cols, cpus_per_main_job, working_dir, sa
     groups = train["tree_id"]
     logging.info(f"Number of different trees is {len(program_data['tree_id'].unique())}")
     full_features = [col for col in program_data.columns if
-                'feature' in col and col not in bootstrap_cols]  # +['partition_branch_vs_mean','partition_branch','partition_size','partition_size_ratio','partition_divergence','divergence_ratio']
+                'feature' in col and 'msa_entropy' not in col and col not in bootstrap_cols]  # +['partition_branch_vs_mean','partition_branch','partition_size','partition_size_ratio','partition_divergence','divergence_ratio']
 
-    logging.info("Evaluating full standard model- including nni feautres")
+    logging.info(f"Evaluating full standard model- including nni feautres, number of features is {len(full_features)}")
     full_model_working_dir = os.path.join(working_dir,'full_model')
     create_dir_if_not_exists(full_model_working_dir)
     model_performance_full,group_performance_full = standard_model_pipeline(train,test, y_train, y_test, groups,full_features, cpus_per_main_job, full_model_working_dir,do_RFE , large_grid, name, validation_dict,extract_predictions)
@@ -143,10 +143,9 @@ def ML_pipeline(program_data, bootstrap_cols, cpus_per_main_job, working_dir, sa
     nni_cols = ['feature_min_ll_diff','feature_max_ll_diff', 'feature_column_variance',
                               'feature_min_ll_diff_norm','feature_max_ll_diff_norm','feature_column_variance_norm']
 
-    fast_features = [col for col in program_data.columns if
-                     'feature' in col and col not in bootstrap_cols and col not in nni_cols]  # +['partition_branch_vs_mean','partition_branch','partition_size','partition_size_ratio','partition_divergence','divergence_ratio']
+    fast_features = [col for col in full_features if col not in nni_cols]  # +['partition_branch_vs_mean','partition_branch','partition_size','partition_size_ratio','partition_divergence','divergence_ratio']
 
-    logging.info("Evaluating fast standard model- no nni feautres")
+    logging.info(f"Evaluating fast standard model- no nni feautres, number of features is {len(fast_features)}")
     fast_model_working_dir = os.path.join(working_dir, 'fast_model')
     create_dir_if_not_exists(fast_model_working_dir)
[…]

### 2023-11-08 programs/FastTree.py
Adding model mispecification

@@ -25,9 +25,12 @@ def fasttree_bootstrap_replicate(msa_path, msa_type, starting_tree, output_tree_
         final_tree = B.read()
     return final_tree
 
-def obtain_ML_fasttree(msa_path, msa_type, output_tree_file):
+def obtain_ML_fasttree(msa_path, msa_type, output_tree_file, model):
     if msa_type == 'DNA':
-        model_str = '-nt -gtr'
+        if model=='JC':
+            model_str = '-nt'
+        else:
+            model_str = '-nt -gtr'
     else:
         model_str = '-wag'
 
@@ -35,9 +38,9 @@ def obtain_ML_fasttree(msa_path, msa_type, output_tree_file):
     execute_command_and_write_to_log(search_command)
     return output_tree_file
 
-def fasttree_pipeline(curr_run_dir, results_folder, msa_path, msa_type, nb):
+def fasttree_pipeline(curr_run_dir, results_folder, msa_path, msa_type, nb, model):
     output_tree_file = os.path.join(results_folder,'fasttree_sh_bootstrap')
-    obtain_ML_fasttree(msa_path, msa_type, output_tree_file)
+    obtain_ML_fasttree(msa_path, msa_type, output_tree_file, model)
     bootstrap_trees = []
     tmp_msa_path = os.path.join(curr_run_dir, 'tmp_msa.fasta')
     bootstrap_trees_file = os.path.join(curr_run_dir,'bootstrap_trees')

### 2023-11-08 programs/raxml.py

@@ -244,7 +244,8 @@ def remove_redundant_sequences(curr_run_directory, prefix,msa_path, model, out_m
         all_msa_records = get_alignment_data(reduced_path)
         SeqIO.write(all_msa_records, out_msa_path, 'fasta')
     else:
-        shutil.copyfile(msa_path,out_msa_path)
+        if msa_path!=out_msa_path:
+            shutil.copyfile(msa_path,out_msa_path)
     for file in os.listdir(curr_run_directory):
         if prefix in file:
             os.remove(os.path.join(curr_run_directory,file))

### 2023-11-08 programs/raxml_grove.py

@@ -25,7 +25,7 @@ def extract_representative_sample(out_dir, min_n_taxa, max_n_taxa, min_n_loci, m
     d = RAxML_grove_get_tree_statistics(out_dir, min_n_taxa, max_n_taxa, min_n_loci, max_n_loci, msa_type)
     k_means_data = d[["TREE_ID", "NUM_TAXA", "BRANCH_LENGTH_MEAN", "BRANCH_LENGTH_VARIANCE"]]
     scaler = preprocessing.StandardScaler()
-    scaled_k_means_data = scaler.fit_transform(k_means_data)
+    scaled_k_means_data = scaler.fit_transform(k_means_data.drop(columns=['TREE_ID']))
     kmeans = KMeans(n_clusters=n_k_means_clusters)
     kmeans_fit = kmeans.fit(scaled_k_means_data)
     k_means_data["label"] = kmeans_fit.labels_
@@ -47,7 +47,7 @@ def extract_model_specification_from_log(log_file, param):
     return res
 
 
-def RAxML_grove_tree_simulation( out_dir, min_n_taxa, max_n_taxa, min_n_loci, max_n_loci, msa_type, tree_id):
+def RAxML_grove_tree_simulation( out_dir, min_n_taxa, max_n_taxa, min_n_loci, max_n_loci, msa_type, tree_id, model_modes):
     msa_type = f"'{msa_type}'"
     tree_id_cmd = "" if tree_id is None else f" and TREE_ID={tree_id}"
     cmd_raxml_grove = f'{RAxML_alisim_path} generate -g alisim -o {out_dir} -q " NUM_TAXA>={min_n_taxa} and NUM_TAXA<={max_n_taxa} and OVERALL_NUM_ALIGNMENT_SITES>={min_n_loci} and OVERALL_NUM_ALIGNMENT_SITES<={max_n_loci}  and DATA_TYPE=={msa_type} {tree_id_cmd}" '
@@ -57,13 +57,30 @@ def RAxML_grove_tree_simulation( out_dir, min_n_taxa, max_n_taxa, min_n_loci, ma
     folder = os.listdir(out_dir)[0]
     true_tree_path_orig = os.path.join(out_dir, folder, 'tree_best.newick')
     model_formulation_file = os.path.join(out_dir, folder, 'tree_best.newick.log')
-    model = extract_model_specification_from_log(model_formulation_file, 'model')
+
     length = extract_model_specification_from_log(model_formulation_file, 'length')
[…]

### 2023-11-08 simulation_edit/bootstrap_edit.py

@@ -203,7 +203,7 @@ def get_bootstrap_and_tree_groups(program, bootstrap_tree_details,mle_path,garba
     parsimony_trees_path = generate_n_tree_topologies(n_pars, bootstrap_tree_details["msa_path"],
                                                       curr_run_directory=garbage_dir,
                                                       seed=1, tree_type='pars',
-                                                      model=bootstrap_tree_details["model_short"])
+                                                      model=bootstrap_tree_details["tree_search_model"])
     tree_tmp_path = os.path.join(garbage_dir, "tmp.tree")
     with open(parsimony_trees_path) as trees_path:
         parsimony_trees = trees_path.read().split("\n")[:-1]
@@ -307,12 +307,12 @@ def msa_path_bootstrap_analysis(msa_path,curr_run_dir, mle_path, true_tree_path,
 
     neighbors_tmp_path = os.path.join(curr_run_dir, "neighbors_tmp.nw")
 
-    curr_pruned_tree_path, curr_pruned_msa_path = get_pruned_tree_and_msa(curr_run_dir, msa_path, bootstrap_tree_details_dict["model_short"],mle_tree_ete)
+    curr_pruned_tree_path, curr_pruned_msa_path = get_pruned_tree_and_msa(curr_run_dir, msa_path, bootstrap_tree_details_dict["tree_search_model"],mle_tree_ete)
 
     neig_ll_evaluation_time = 0
     per_site_original_tree_ll = np.array(
         raxml_compute_tree_per_site_ll(garbage_dir, curr_pruned_msa_path, curr_pruned_tree_path, ll_on_data_prefix="orig_tree_ll",
-                                       model=bootstrap_tree_details_dict["model_short"], opt=True))
+                                       model=bootstrap_tree_details_dict["tree_search_model"], opt=True))
 
     for node in mle_tree_ete.iter_descendants():
 
@@ -329,9 +329,9 @@ def msa_path_bootstrap_analysis(msa_path,curr_run_dir, mle_path, true_tree_path,
                 neighbor = Tree(neighbors_tmp_path, format=1)
                 curr_pruned_tree_path, curr_pruned_msa_path = get_pruned_tree_and_msa(curr_run_dir, msa_path,
                                                                                       bootstrap_tree_details_dict[
-                                                                                          "model_short"], neighbor)
+                                                                                          "tree_search_model"], neighbor)
 
[…]

### 2023-11-08 simulations_generation/simulations.py

@@ -42,23 +42,25 @@ def simulate_msa(output_prefix, tree_file, model, length, ret_cmd = False):
         return msa_path
 
 
-def single_simulated_MSA_pipeline(msa_path,tree_sim_dict, curr_run_dir, args):
+def obtain_tree_search_bootstrap_results(msa_path, tree_sim_dict, curr_run_dir, tree_search_model, args):
     logging.info("Evaluating likelihood on current MSA")
-    msa_results_dict = get_msa_stats(msa_path, tree_sim_dict["model"])
+    msa_results_dict = get_msa_stats(msa_path, tree_search_model)
     msa_results_dict.update(tree_sim_dict)
     tree_searches_folder = os.path.join(curr_run_dir, "all_tree_searches")
     results_folder = os.path.join(curr_run_dir, "results_folder")
     create_or_clean_dir(tree_searches_folder)
     create_or_clean_dir(results_folder)
+    logging.info(f"Model mode is {tree_sim_dict['model_mode']}, using {tree_search_model} model for tree search")
     boot_tree_fasttree = fasttree_pipeline(tree_searches_folder, results_folder, msa_path, msa_type=args.msa_type,
-                                           nb=args.nb_fasttree)
-    boot_tree_raxml =  raxml_bootstrap_pipeline(tree_searches_folder,results_folder , msa_path, prefix ="boot", model =tree_sim_dict["model_short"],  n_cpus=1,
+                                           nb=args.nb_fasttree, model = tree_search_model)
+    boot_tree_raxml =  raxml_bootstrap_pipeline(tree_searches_folder,results_folder , msa_path, prefix ="boot", model =tree_search_model,  n_cpus=1,
                                                 n_workers='auto')
-    boot_tree_iqtree = iqtree_pipeline(tree_searches_folder,results_folder , msa_path, model = tree_sim_dict["model_short"], nb=args.nb_iqtree, prefix = "iqtree_boot")
+    boot_tree_iqtree = iqtree_pipeline(tree_searches_folder,results_folder , msa_path, model = tree_search_model, nb=args.nb_iqtree, prefix = "iqtree_boot")
     for d in [boot_tree_raxml, boot_tree_iqtree, boot_tree_fasttree]:
         d.update(msa_results_dict)
         d.update(tree_sim_dict)
         d['msa_path'] = msa_path
+        d['tree_search_model'] = tree_search_model
[…]

### 2023-11-08 simulations_generation/simulations_argparser.py

@@ -1,10 +1,10 @@
 import argparse
 def main_parser():
     parser = argparse.ArgumentParser()
-    parser.add_argument('--number_of_trees', action='store', type=int, default=10)
+    parser.add_argument('--number_of_trees', action='store', type=int, default=2)
     parser.add_argument('--number_of_MSAs_per_tree', type=int, default=1)
-    parser.add_argument('--min_n_taxa', type = int, default = 20)
-    parser.add_argument('--max_n_taxa', type=int, default=30)
+    parser.add_argument('--min_n_taxa', type = int, default = 15)
+    parser.add_argument('--max_n_taxa', type=int, default=20)
     parser.add_argument('--min_n_loci', type=int, default=100)
     parser.add_argument('--max_n_loci',type=int,default = 200)
     parser.add_argument('--n_jobs', type = int, default= 1)
@@ -20,6 +20,7 @@ def main_parser():
     parser.add_argument('--calc_features', action='store_true', default=True)
     parser.add_argument('--pre_chosen_tree_ids',action='store_true', default= False)
     parser.add_argument('--n_k_means_clusters',type=int, default =3)
+    parser.add_argument('--model_modes', type=str, default='standard_downgrade_upgrade') #'mispecification
     return parser
 
 

### 2023-11-08 programs/raxml_grove.py
changing the validation

@@ -7,6 +7,7 @@ import ast
 import pandas as pd
 from sklearn.cluster import KMeans
 from sklearn import preprocessing
+from sklearn.cluster import AgglomerativeClustering
 
 def RAxML_grove_get_tree_statistics(out_dir, min_n_taxa, max_n_taxa, min_n_loci, max_n_loci, msa_type):
     msa_type = f"'{msa_type}'"
@@ -27,12 +28,13 @@ def extract_representative_sample(out_dir, min_n_taxa, max_n_taxa, min_n_loci, m
     scaler = preprocessing.StandardScaler()
     scaled_k_means_data = scaler.fit_transform(k_means_data.drop(columns=['TREE_ID']))
     kmeans = KMeans(n_clusters=n_k_means_clusters)
+    #kmeans = AgglomerativeClustering().fit(scaled_k_means_data)
     kmeans_fit = kmeans.fit(scaled_k_means_data)
     k_means_data["label"] = kmeans_fit.labels_
     scaled_k_means_data_distances = (kmeans.transform(scaled_k_means_data)**2).sum(axis=1)
     k_means_data["dist_from_center"] = scaled_k_means_data_distances
-    repersentative_sample = k_means_data.sort_values("dist_from_center").groupby("label").head(1).reset_index()
-    return list(repersentative_sample["TREE_ID"])
+    repersentative_sample = k_means_data.sort_values("dist_from_center")[["label","TREE_ID"]].groupby("label").agg(list).to_dict()['TREE_ID']
+    return repersentative_sample
 
 
 def extract_model_specification_from_log(log_file, param):

### 2023-11-08 simulations_generation/run_parallel_simulations.py

@@ -56,8 +56,8 @@ def main():
         representative_sample_trees = extract_representative_sample(out_dir, args.min_n_taxa, args.max_n_taxa, args.min_n_loci, args.max_n_loci, args.msa_type, args.n_k_means_clusters)
         logging.info(f"Chosen tree IDS are: {representative_sample_trees}")
         n_trees_per_jobs  = math.ceil(len(representative_sample_trees)/args.n_jobs)
-        for job_ind in range((args.n_jobs)):
-            curr_job_trees = representative_sample_trees[job_ind*n_trees_per_jobs:job_ind*n_trees_per_jobs+n_trees_per_jobs]
+        for job_ind in range((args.n_k_means_clusters)):
+            curr_job_trees = representative_sample_trees[job_ind]
             if len(curr_job_trees)==0:
                 break
             logging.info(f"Submitted job number {job_ind}")

### 2023-11-08 simulations_generation/simulations.py

@@ -179,11 +179,15 @@ def main():
             tree_ids = JOB_TREES_FILE.read().split("\n")
         for i,tree_id in enumerate(tree_ids):
             logging.info(f"Tree id = {tree_id}")
-            all_models_tree_sim_dict = obtain_tree_sim_dict(i, args, curr_job_folder, tree_id)
-            for model_mode in all_models_tree_sim_dict:
-                all_results_df_raxml,all_results_df_iqtree,all_results_df_fasttree = obtain_all_results_per_tree(args, all_models_tree_sim_dict[model_mode], curr_job_folder, all_results_df_raxml,
-                                            all_results_df_iqtree, all_results_df_fasttree)
-
+            try:
+                all_models_tree_sim_dict = obtain_tree_sim_dict(i, args, curr_job_folder, tree_id)
+                for model_mode in all_models_tree_sim_dict:
+                    all_results_df_raxml,all_results_df_iqtree,all_results_df_fasttree = obtain_all_results_per_tree(args, all_models_tree_sim_dict[model_mode], curr_job_folder, all_results_df_raxml,
+                                                all_results_df_iqtree, all_results_df_fasttree)
+                logging.info("Done with current tree id")
+                break
+            except:
+                logging.info(f"Could not run on id {tree_id}, trying next")
     else:
         logging.info("Using random chosen tree IDs")
         for i in range(args.number_of_trees):

### 2023-11-08 simulations_generation/simulations_argparser.py

@@ -18,9 +18,9 @@ def main_parser():
     parser.add_argument('--nb_iqtree',type = int, default= 1000)
     parser.add_argument('--nb_fasttree', type=int, default=10)
     parser.add_argument('--calc_features', action='store_true', default=True)
-    parser.add_argument('--pre_chosen_tree_ids',action='store_true', default= False)
+    parser.add_argument('--pre_chosen_tree_ids',action='store_true', default= True)
     parser.add_argument('--n_k_means_clusters',type=int, default =3)
-    parser.add_argument('--model_modes', type=str, default='standard_downgrade_upgrade') #'mispecification
+    parser.add_argument('--model_modes', type=str, default='standard') #'mispecification
     return parser
 
 

### 2023-11-12 R_code/paper_analysis.Rmd
new changes

@@
 ```{r}
-prefix = "/Users/noa/Workspace/ML_results"
+prefix = "/Users/noa/Workspace/bootstrap_results/remote_results/"
 
-fasttree_all_models_performance = read_tsv(paste(prefix,"/fasttree/all_models_performance.tsv", sep = ""))
-fasttree_final_models_performance = read_tsv(paste(prefix,"/fasttree/final_model_performance.tsv", sep = ""))
-fasttree_group_performance = read_tsv(paste(prefix,"/fasttree/groups_performance.tsv", sep = ""))
-fasttree_vi =  read_tsv(paste(prefix,"/fasttree/model_standard_vi.tsv", sep = ""))
-fasttree_enriched_test = read_tsv(paste(prefix,"/fasttree/test_model_standard_enriched.tsv", sep = ""))
+fasttree_all_models_performance = read_tsv(paste(prefix,"ML_results/fasttree/all_models_performance.tsv", sep = ""))
+fasttree_final_models_performance = read_tsv(paste(prefix,"ML_results/fasttree/final_model_performance.tsv", sep = ""))
+fasttree_group_performance = read_tsv(paste(prefix,"ML_results/fasttree/groups_performance.tsv", sep = ""))
+fasttree_vi =  read_tsv(paste(prefix,"ML_results/fasttree/model_standard_vi.tsv", sep = ""))
+fasttree_enriched_test = read_tsv(paste(prefix,"ML_results/fasttree/test_model_standard_enriched.tsv", sep = ""))
+fasttree_validation = read_tsv(paste(prefix,"validation_data/simulations_df_fasttree.tsv", sep = ""))
 #####
-iqtree_all_models_performance = read_tsv(paste(prefix,"/iqtree/all_models_performance.tsv", sep = ""))
-iqtree_final_models_performance = read_tsv(paste(prefix,"/iqtree/final_model_performance.tsv", sep = ""))
-iqtree_group_performance = read_tsv(paste(prefix,"/iqtree/groups_performance.tsv", sep = ""))
-iqtree_vi =  read_tsv(paste(prefix,"/iqtree/model_standard_vi.tsv", sep = ""))
-iqtree_enriched_test = read_tsv(paste(prefix,"/iqtree/test_model_standard_enriched.tsv", sep = ""))
+iqtree_all_models_performance = read_tsv(paste(prefix,"ML_results/iqtree/all_models_performance.tsv", sep = ""))
+iqtree_final_models_performance = read_tsv(paste(prefix,"/ML_resultsiqtree/final_model_performance.tsv", sep = ""))
+iqtree_group_performance = read_tsv(paste(prefix,"ML_results/iqtree/groups_performance.tsv", sep = ""))
+iqtree_vi =  read_tsv(paste(prefix,"ML_results/iqtree/model_standard_vi.tsv", sep = ""))
+iqtree_enriched_test = read_tsv(paste(prefix,"ML_results/iqtree/test_model_standard_enriched.tsv", sep = ""))
+iqtree_validation = read_tsv(paste(prefix,"validation_data/simulations_df_iqtree.tsv", sep = ""))
 #####
-raxml_all_models_performance = read_tsv(paste(prefix,"/raxml/all_models_performance.tsv", sep = ""))
-raxml_final_models_performance = read_tsv(paste(prefix,"/raxml/final_model_performance.tsv", sep = ""))
-raxml_group_performance = read_tsv(paste(prefix,"/raxml/groups_performance.tsv", sep = ""))
-raxml_vi =  read_tsv(paste(prefix,"/raxml/model_standard_vi.tsv", sep = ""))
-raxml_enriched_test = read_tsv(paste(prefix,"/raxml/test_model_standard_enriched.tsv", sep = ""))
+raxml_all_models_performance = read_tsv(paste(prefix,"ML_results/raxml/all_models_performance.tsv", sep = ""))
+raxml_final_models_performance = read_tsv(paste(prefix,"ML_results/raxml/final_model_performance.tsv", sep = ""))
+raxml_group_performance = read_tsv(paste(prefix,"ML_results/raxml/groups_performance.tsv", sep = ""))
+raxml_vi =  read_tsv(paste(prefix,"ML_results/raxml/model_standard_vi.tsv", sep = ""))
+raxml_enriched_test = read_tsv(paste(prefix,"ML_results/raxml/test_model_standard_enriched.tsv", sep = ""))
+raxml_validation = read_tsv(paste(prefix,"validation_data/simulations_df_raxml.tsv", sep = ""))
 
@@
 ```{r}
-program = 'iqtree'
+program = 'raxml'
 if (program=='fasttree') {
     all_models_performance = fasttree_all_models_performance
-    final_models_performance = fasttree_final_models_performance
+    final_models_performance = fasttree_final_models_performance %>% filter(analysis_type!='feature_standard_fasttree_boot_support')
     group_performance = fasttree_group_performance
@@
 ```{r}
-final_models_performance  %>% filter (dataset=='test', metric_type=='all_data') %>% dplyr::select (analysis_type,name,balanced_accuracy_score, mcc, AUC)
+
+out_path <-paste(prefix,"/",program,"/vi_text",".tsv", sep = "")
+
+paper_vi<-final_models_performance  %>% filter ((dataset=='test'|dataset=='train') , metric_type=='all_data') %>% dplyr::select (dataset,analysis_type,name,mcc,balanced_accuracy_score,precision, recall, AUC,average_precision)  %>% arrange(dataset)
+write_tsv(paper_vi,out_path)
+
+paper_vi
+
 ```
@@
 
+error_vs_size
 error_vs_size<- all_models_performance %>% filter (metric_type=='all_data', dataset =='test')
-error_vs_size_plt = error_vs_size %>% mutate(n_MSAs = sample_frac) %>% ggplot(aes(x=n_MSAs, y = AUC))+geom_line()+ expand_limits(y = 0.5)+xlab("Number of MSAs")+ theme(text = element_text(size = 13))+ylab("AUC")+geom_point()
+error_vs_size_plt = error_vs_size %>% mutate(n_MSAs = sample_frac*2007) %>% ggplot(aes(x=n_MSAs, y = mcc))+geom_line()+ expand_limits(y = 0)+xlab("Number of MSAs")+ theme(text = element_text(size = 13))+ylab("MCC")+geom_point()
 
@@
 ```
+
+```{r}
+cor.test( enriched_test$feature_msa_pypythia_msa_difficulty,  enriched_test$prob_predictions)
+cor.test( enriched_test$feature_msa_n_seq, enriched_test$prob_predictions)
+cor.test( enriched_test$feature_msa_n_loci, enriched_test$prob_predictions)
+```
+
 Validation_data
 ```{r}
-validation_data %>% dplyr::select (boot_run_time, extraction_running_time,nni_ll_eval_running_time)
+raxml_data = read_tsv("/Users/noa/Workspace/simulations_df_raxml.tsv")
 
+```
+
+```{r}
+running_time_data<- enriched_test%>% dplyr::select (boot_run_time,extraction_running_time,nni_ll_eval_running_time) %>% mutate (diff1 = boot_run_time/extraction_running_time, diff2 = boot_run_time/(extraction_running_time+nni_ll_eval_running_time
+)) 
+summary(running_time_data %>% pull (diff1))
+summary(running_time_data %>% pull (diff2))
+summary(running_time_data %>% pull (extraction_running_time))
+summary(running_time_data %>% pull (nni_ll_eval_running_time))
+#enriched_test %>% dplyr::select (boot_run_time, extraction_runni
+```
+
+```{r}
+
+running_time_data<- enriched_test%>% dplyr::select (program_boot_run_time,extraction_running_time,nni_ll_eval_running_time) %>% mutate (diff1 = program_boot_run_time/extraction_running_time, diff2 = program_boot_run_time/(extraction_running_time+nni_ll_eval_running_time
+)) 
+summary(running_time_data %>% pull (diff1))
+summary(running_time_data %>% pull (diff2))
+summary(running_time_data %>% pull (extraction_running_time))
+summary(running_time_data %>% pull (nni_ll_eval_running_time))
+summary(running_time_data %>% pull (program_boot_run_time))
+#enriched_test %>% dplyr::select (boot_run_time, extraction_running_time, nni_ll_eval_running_time)
 ```

### 2023-11-12 ML_pipeline/ML_boot_edit.py

@@ -132,8 +132,8 @@ def ML_pipeline(program_data, bootstrap_cols, cpus_per_main_job, working_dir, sa
     groups = train["tree_id"]
     logging.info(f"Number of different trees is {len(program_data['tree_id'].unique())}")
     full_features = [col for col in program_data.columns if
-                'feature' in col and 'msa_entropy' not in col and col not in bootstrap_cols]  # +['partition_branch_vs_mean','partition_branch','partition_size','partition_size_ratio','partition_divergence','divergence_ratio']
-
+                'feature' in col and 'msa_entropy' not in col and col not in bootstrap_cols and 'll_diff_norm' not in col and 'column_variance' not in col]  # +['partition_branch_vs_mean','partition_branch','partition_size','partition_size_ratio','partition_divergence','divergence_ratio']
+    logging.info(f"Full features are: {full_features}")
     logging.info(f"Evaluating full standard model- including nni feautres, number of features is {len(full_features)}")
     full_model_working_dir = os.path.join(working_dir,'full_model')
     create_dir_if_not_exists(full_model_working_dir)
@@ -181,7 +181,7 @@ def main():
     parser.add_argument('--RFE', action='store_true', default= False)
     parser.add_argument('--full_grid', action='store_true', default=False)
     parser.add_argument('--cpus_per_main_job', type = int, default=1)
-    parser.add_argument('--sample_fracs', type = str, default='0.5_1')
+    parser.add_argument('--sample_fracs', type = str, default='0.25_0.5_1')
     parser.add_argument('--inc_sample_fracs', action='store_true', default=True)
     parser.add_argument('--main_data_folder',type = str, default = '/Users/noa/Workspace/bootstrap_results/remote_results')
     parser.add_argument('--validation_data_folder', type=str,

### 2023-11-12 R_code/paper_analysis.Rmd
new changes

@@
 iqtree_all_models_performance = read_tsv(paste(prefix,"ML_results/iqtree/all_models_performance.tsv", sep = ""))
-iqtree_final_models_performance = read_tsv(paste(prefix,"/ML_resultsiqtree/final_model_performance.tsv", sep = ""))
+iqtree_final_models_performance = read_tsv(paste(prefix,"ML_results/iqtree/final_model_performance.tsv", sep = ""))
 iqtree_group_performance = read_tsv(paste(prefix,"ML_results/iqtree/groups_performance.tsv", sep = ""))
@@
     enriched_test = raxml_enriched_test
-    validation_data = raxml_validation_data
+    validation_data = raxml_validation
 }
@@
 ```{r}
-raxml_data = read_tsv("/Users/noa/Workspace/simulations_df_raxml.tsv")
+validation_data$
 
@@
 ```
+```{r}
+validation_data$model_mode
+```
 

### 2023-11-12 ML_pipeline/ML_boot_edit.py

@@ -182,10 +182,10 @@ def main():
     parser.add_argument('--full_grid', action='store_true', default=False)
     parser.add_argument('--cpus_per_main_job', type = int, default=1)
     parser.add_argument('--sample_fracs', type = str, default='0.25_0.5_1')
-    parser.add_argument('--inc_sample_fracs', action='store_true', default=True)
-    parser.add_argument('--main_data_folder',type = str, default = '/Users/noa/Workspace/bootstrap_results/remote_results')
+    parser.add_argument('--inc_sample_fracs', action='store_true', default=False)
+    parser.add_argument('--main_data_folder',type = str, default = '/Users/noa/Workspace/bootstrap_results/remote_results/full_data')
     parser.add_argument('--validation_data_folder', type=str,
-                        default='/Users/noa/Workspace/bootstrap_results/remote_results')
+                        default='/Users/noa/Workspace/bootstrap_results/remote_results/validation_data')
     args = parser.parse_args()
     log_file_path = os.path.join(args.working_dir, "ML.log")
     logging.basicConfig(filename=log_file_path, level=logging.DEBUG)
@@ -202,7 +202,9 @@ def main():
         program_data = program_data.dropna(axis=0)
         sample_fracs = [float(frac) for frac in (args.sample_fracs).split('_')]
         all_model_merics = pd.DataFrame()
-        validation_dict = {'val': program_validation_data}
+        validation_dict = {}
+        for model_mode in np.unique(program_validation_data["model_mode"]):
+            validation_dict[f'val_{model_mode}'] = program_validation_data.loc[program_validation_data.model_mode==model_mode].copy()
         if args.inc_sample_fracs:
             for sample_frac in sample_fracs:
                 logging.info(f"\n#Sample frac = {sample_frac}")

### 2023-11-12 simulations_generation/simulations_argparser.py
important fix

@@ -18,7 +18,7 @@ def main_parser():
     parser.add_argument('--nb_iqtree',type = int, default= 1000)
     parser.add_argument('--nb_fasttree', type=int, default=10)
     parser.add_argument('--calc_features', action='store_true', default=True)
-    parser.add_argument('--pre_chosen_tree_ids',action='store_true', default= True)
+    parser.add_argument('--pre_chosen_tree_ids',action='store_true', default= False)
     parser.add_argument('--n_k_means_clusters',type=int, default =3)
     parser.add_argument('--model_modes', type=str, default='standard') #'mispecification
     return parser

### 2023-11-12 ML_pipeline/ML_boot_edit.py
change

@@ -188,7 +188,7 @@ def main():
                         default='/Users/noa/Workspace/bootstrap_results/remote_results/validation_data')
     args = parser.parse_args()
     log_file_path = os.path.join(args.working_dir, "ML.log")
-    logging.basicConfig(filename=log_file_path, level=logging.DEBUG)
+    logging.basicConfig(filename=log_file_path, level=logging.INFO)
     for program in ['fasttree','iqtree','raxml']: #,'fasttree','raxml'
         logging.info(f"Program = {program}")
         program_data = pd.read_csv(os.path.join(args.main_data_folder,f'simulations_df_{program}.tsv'),sep='\t')

### 2023-11-12 simulations_generation/simulations_argparser.py

@@ -18,7 +18,7 @@ def main_parser():
     parser.add_argument('--nb_iqtree',type = int, default= 1000)
     parser.add_argument('--nb_fasttree', type=int, default=10)
     parser.add_argument('--calc_features', action='store_true', default=True)
-    parser.add_argument('--pre_chosen_tree_ids',action='store_true', default= False)
+    parser.add_argument('--pre_chosen_tree_ids',action='store_true', default= False )
     parser.add_argument('--n_k_means_clusters',type=int, default =3)
     parser.add_argument('--model_modes', type=str, default='standard') #'mispecification
     return parser

### 2023-11-13 ML_pipeline/ML_boot_edit.py
new change

@@ -180,6 +180,7 @@ def main():
     parser.add_argument('--working_dir', type = str, default = os.getcwd())
     parser.add_argument('--RFE', action='store_true', default= False)
     parser.add_argument('--full_grid', action='store_true', default=False)
+    parser.add_argument('--programs',type = str, default = 'iqtree_raxml_fasttree')
     parser.add_argument('--cpus_per_main_job', type = int, default=1)
     parser.add_argument('--sample_fracs', type = str, default='0.25_0.5_1')
     parser.add_argument('--inc_sample_fracs', action='store_true', default=False)
@@ -189,7 +190,7 @@ def main():
     args = parser.parse_args()
     log_file_path = os.path.join(args.working_dir, "ML.log")
     logging.basicConfig(filename=log_file_path, level=logging.INFO)
-    for program in ['fasttree','iqtree','raxml']: #,'fasttree','raxml'
+    for program in args.programs.split('_'): 
         logging.info(f"Program = {program}")
         program_data = pd.read_csv(os.path.join(args.main_data_folder,f'simulations_df_{program}.tsv'),sep='\t')
         transform_data(program_data)

### 2023-11-14 ML_pipeline/ML_boot_edit.py
new changes

@@ -9,6 +9,7 @@ sys.path.append(PROJECT_ROOT_DIRECRTORY)
 
 from ML_utils.ML_algorithms_and_hueristics import *
 from side_code.file_handling import create_or_clean_dir, create_dir_if_not_exists
+from side_code.combine_current_results import unify_results_across_jobs
 from side_code.MSA_manipulation import get_MSA_seq_names
 from sklearn.metrics import balanced_accuracy_score
 import os
@@ -60,34 +61,45 @@ def get_bootstrap_col(program):
     return bootstrap_cols
 
 
-def bootstrap_model_pipeline(working_dir,train, y_train,test, y_test,features,bootstrap_col,groups,cpus_per_main_job,sample_frac,do_RFE , large_grid, name):
-    X_train_only_boot = train[[bootstrap_col]]
-    X_train_only_boot["ignore"] = 1
-    X_test_only_boot = test[[bootstrap_col]]
-    X_test_only_boot["ignore"] = 1
-    X_train_inc_boot = train[[col for col in train.columns if col in features] + [bootstrap_col]]
-    X_test_inc_boot = test[[col for col in train.columns if col in features] + [bootstrap_col]]
-    model_only_boot = ML_model(X_train_only_boot, groups, y_train, n_jobs=cpus_per_main_job,
+def generate_data_dict_from_column_subset(data_dict, column_subset):
+    res = {}
+    for dataset in data_dict:
+        data = data_dict[dataset]
+        X = data[column_subset]
+        if len(column_subset)==1:
+            X["ignore"] = 1
+        res[dataset] = {'X': X, 'y': data['true_binary_support'], 'full_data': data}
+    return res
+
+
+
+def bootstrap_model_pipeline(working_dir,train,test,validation_dict,features,bootstrap_col,groups,cpus_per_main_job,sample_frac,do_RFE , large_grid, name):
+
+
+    data_dict = {'test':test,'train':train}
+    data_dict.update(validation_dict)
+    only_boot_dict = generate_data_dict_from_column_subset(data_dict, column_subset=[bootstrap_col])
+
+    inc_boot_dict = generate_data_dict_from_column_subset(data_dict, column_subset = [col for col in train.columns if col in features] + [bootstrap_col])
+
+    model_only_boot = ML_model(only_boot_dict["train"]["X"], groups, only_boot_dict["train"]["y"], n_jobs=cpus_per_main_job,
                                path=os.path.join(working_dir, f'model_only_boot'),
[…]

### 2023-11-15 R_code/paper_analysis.Rmd
new changes

--- 2023-11-06 R_code/paper_analysis.Rmd
@@
 ```{r}
-prefix = "/Users/noa/Workspace/ML_results"
+prefix = "/Users/noa/Workspace/bootstrap_results/remote_results/"
 
-fasttree_all_models_performance = read_tsv(paste(prefix,"/fasttree/all_models_performance.tsv", sep = ""))
-fasttree_final_models_performance = read_tsv(paste(prefix,"/fasttree/final_model_performance.tsv", sep = ""))
-fasttree_group_performance = read_tsv(paste(prefix,"/fasttree/groups_performance.tsv", sep = ""))
-fasttree_vi =  read_tsv(paste(prefix,"/fasttree/model_standard_vi.tsv", sep = ""))
-fasttree_enriched_test = read_tsv(paste(prefix,"/fasttree/test_model_standard_enriched.tsv", sep = ""))
+fasttree_all_models_performance = read_tsv(paste(prefix,"ML_results/fasttree/all_models_performance.tsv", sep = ""))
+fasttree_final_models_performance = read_tsv(paste(prefix,"ML_results/fasttree/final_model_performance.tsv", sep = ""))
+fasttree_group_performance = read_tsv(paste(prefix,"ML_results/fasttree/groups_performance.tsv", sep = ""))
+fasttree_vi =  read_tsv(paste(prefix,"ML_results/fasttree/final_model/full_model/model_standard_vi.tsv", sep = ""))
+fasttree_enriched_test = read_tsv(paste(prefix,"ML_results/fasttree/final_model/full_model/test_model_standard_enriched.tsv", sep = ""))
+fasttree_validation = read_tsv(paste(prefix,"validation_data/simulations_df_fasttree.tsv", sep = ""))
 #####
-iqtree_all_models_performance = read_tsv(paste(prefix,"/iqtree/all_models_performance.tsv", sep = ""))
-iqtree_final_models_performance = read_tsv(paste(prefix,"/iqtree/final_model_performance.tsv", sep = ""))
-iqtree_group_performance = read_tsv(paste(prefix,"/iqtree/groups_performance.tsv", sep = ""))
-iqtree_vi =  read_tsv(paste(prefix,"/iqtree/model_standard_vi.tsv", sep = ""))
-iqtree_enriched_test = read_tsv(paste(prefix,"/iqtree/test_model_standard_enriched.tsv", sep = ""))
+iqtree_all_models_performance = read_tsv(paste(prefix,"ML_results/iqtree/all_models_performance.tsv", sep = ""))
+iqtree_final_models_performance = read_tsv(paste(prefix,"ML_results/iqtree/final_model_performance.tsv", sep = ""))
+iqtree_group_performance = read_tsv(paste(prefix,"ML_results/iqtree/groups_performance.tsv", sep = ""))
+iqtree_vi =  read_tsv(paste(prefix,"ML_results/iqtree/final_model/full_model/model_standard_vi.tsv", sep = ""))
+iqtree_enriched_test = read_tsv(paste(prefix,"ML_results/iqtree/final_model/full_model/test_model_standard_enriched.tsv", sep = ""))
+iqtree_validation = read_tsv(paste(prefix,"validation_data/simulations_df_iqtree.tsv", sep = ""))
 #####
-raxml_all_models_performance = read_tsv(paste(prefix,"/raxml/all_models_performance.tsv", sep = ""))
-raxml_final_models_performance = read_tsv(paste(prefix,"/raxml/final_model_performance.tsv", sep = ""))
-raxml_group_performance = read_tsv(paste(prefix,"/raxml/groups_performance.tsv", sep = ""))
-raxml_vi =  read_tsv(paste(prefix,"/raxml/model_standard_vi.tsv", sep = ""))
-raxml_enriched_test = read_tsv(paste(prefix,"/raxml/test_model_standard_enriched.tsv", sep = ""))
+raxml_all_models_performance = read_tsv(paste(prefix,"ML_results/raxml/all_models_performance.tsv", sep = ""))
+raxml_final_models_performance = read_tsv(paste(prefix,"ML_results/raxml/final_model_performance.tsv", sep = ""))
+raxml_group_performance = read_tsv(paste(prefix,"ML_results/raxml/groups_performance.tsv", sep = ""))
+raxml_vi =  read_tsv(paste(prefix,"ML_results/raxml/final_model/full_model/model_standard_vi.tsv", sep = ""))
+raxml_enriched_test = read_tsv(paste(prefix,"ML_results/raxml/final_model/full_model/test_model_standard_enriched.tsv", sep = ""))
+raxml_validation = read_tsv(paste(prefix,"validation_data/simulations_df_raxml.tsv", sep = ""))
 
@@
 ```{r}
-program = 'iqtree'
+program = 'fasttree'
 if (program=='fasttree') {
     all_models_performance = fasttree_all_models_performance
-    final_models_performance = fasttree_final_models_performance
+    final_models_performance = fasttree_final_models_performance %>% filter(analysis_type!='feature_standard_fasttree_boot_support')
     group_performance = fasttree_group_performance
@@
     enriched_test = fasttree_enriched_test
-    validation_data = fasttree_validation_data
+    #validation_data = fasttree_validation_data
 
@@
     group_performance = iqtree_group_performance
-    vi = fasttree_vi
+    vi = iqtree_vi
     enriched_test = iqtree_enriched_test
-    validation_data = iqtree_validation_data
+    #validation_data = iqtree_validation_data
 } else if (program=='raxml') {
@@
     group_performance = raxml_group_performance
-    vi = fasttree_vi
+    vi = raxml_vi
     enriched_test = raxml_enriched_test
-    validation_data = raxml_validation_data
+    #validation_data = raxml_validation
 }
@@
 ```{r}
-final_models_performance  %>% filter (dataset=='test', metric_type=='all_data') %>% dplyr::select (analysis_type,name,balanced_accuracy_score, mcc, AUC)
+
+out_path <-paste(prefix,"ML_results/",program,"/performance_text",".tsv", sep = "")
+
+paper_performance<-final_models_performance  %>% filter (metric_type=='all_data') %>% dplyr::select (dataset,analysis_type,name,mcc,balanced_accuracy_score,precision, recall, AUC,average_precision)  %>% arrange(dataset) %>% mutate (description = case_when(analysis_type == "full_standard" & name =="model_standard" ~ "all_features",
+                                analysis_type == "fast_standard" & name =="model_standard" ~ "All features except NNI features",
+                                analysis_type == "bootstrap_support" & name =="raw_only_boot" ~ "Raw bootstrap support",
+                                analysis_type == "bootstrap_support" & name =="only_boot" ~ "Calibrated bootstrap support",
+                                analysis_type == "bootstrap_support" & name =="inc_boot" ~ "All features+ bootstrap support",
+                                    analysis_type == "feature_aLRT_iqtree_support" & name =="raw_only_boot" ~ "Raw aLRT support",
+                                analysis_type == "feature_aLRT_iqtree_support" & name =="only_boot" ~ "Calibrated aLRT support",
+                                analysis_type == "feature_aLRT_iqtree_support" & name =="inc_boot" ~ "All features+ aLRT support",
+                                                                    analysis_type == "feature_aBayes_iqtree_support" & name =="raw_only_boot" ~ "Raw aBayes support",
+                                analysis_type == "feature_aBayes_iqtree_support" & name =="only_boot" ~ "Calibrated aBayes support",
+                                analysis_type == "feature_aBayes_iqtree_support" & name =="inc_boot" ~ "All features+ aBayes support",
+                                )) %>% dplyr::select (dataset, description, mcc, balanced_accuracy_score,precision, recall)  %>% arrange(dataset)
+
+paper_performance
+
 ```
@@
 ```{r}
-vi %>% arrange (-`Gini-importance`)
+out_path <-paste(prefix,"ML_results/",program,"/vi_text",".tsv", sep = "")
+
+vi_data<- vi %>% arrange (`...1`) %>% mutate (feature_name = case_when(
+  `...1` == "feature_min_ll_diff"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
+  `...1` =="feature_max_ll_diff"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",
+
+  `...1` =='feature_mean_parsimony_trees_binary'~"Fraction of parsimoy trees in which the bipartition exists",
+  `...1` =='feature_mean_parsimony_trees_neighbors_binary'~"Fraction of parsimony trees in which the neighbouring bipartitions exist",
+
+    `...1`=='feature_mean_all_ML_boot_raxml_neighbors_binary'~'Fraction of final ML trees in which the bipartition exists',
+
+   `...1` =='feature_mean_all_ML_boot_raxml_binary'~"Fraction of parsimony trees in which the neighbouring bipartitions exist",
+
+`...1` =='feature_mean_parsimony_trees'~"Mean transfer distance of the bipartition across parsimoy trees",
+`...1` =='feature_mean_parsimony_trees_neighbors'~"Mean transfer distance of the neighbouring bipartitions across parsimoy trees",
+
+`...1` =='feature_mean_all_ML_boot_raxml'~'Mean transfer distance of the bipartition across final ML trees',
+`...1` =='feature_mean_all_ML_boot_raxml_neighbors'~'Mean transfer distance of the neighbouring bipartitions across final ML trees',
+
+`...1` =='feature_partition_branch'~'Branch length at the parititon',
+`...1` =='feature_partition_branch_vs_mean'~'Branch length at the parititon divided by total tree divergence',
+
+`...1` =='feature_partition_divergence'~'Total divergence in the smaller subtree defined by the bipartition',
+`...1` =='feature_divergence_ratio'~'Total divergence in the smaller subtree defined by the bipartition divided by total tree divergence',
+`...1` =='feature_partition_size'~'Number of leaves in the smaller subtree defined by the bipartition',
+`...1` =='feature_partition_size_ratio'~'Fraction of leaves in the smaller subtree defined by the bipartition',
+
+`...1` =='feature_min_neighbor_brlen'~'Minimal branch length among the neighbouring branches',
+`...1` =='feature_mean_neighbor_brlen'~'Mean branch length among the neighbouring branches',
+`...1` =='feature_msa_constant_sites_pct'~'Fraction of constant sites among the MSA sites',
+`...1` =='feature_msa_n_loci'~'Number of positions in the MSA',
+`...1` =='feature_msa_n_unique_sites'~'Number of unique positions in the MSA',
+`...1` =='feature_msa_n_seq'~'Number of sequences in the MSA',
+`...1` =='feature_msa_pypythia_msa_difficulty'~'MSA difficulty',
+
+)) %>% dplyr::select (`...1`,feature_name,`Gini-importance`, mcc ) %>% arrange(feature_name,-`Gini-importance`)
+vi_data
+write_tsv(vi_data,out_path)
+
 ```
@@
 ```{r}
+all_models_performance
+```
+
+```{r}
 library(predtools)
@@
 
-error_vs_size<- all_models_performance %>% filter (metric_type=='all_data', dataset =='test')
-error_vs_size_plt = error_vs_size %>% mutate(n_MSAs = sample_frac) %>% ggplot(aes(x=n_MSAs, y = AUC))+geom_line()+ expand_limits(y = 0.5)+xlab("Number of MSAs")+ theme(text = element_text(size = 13))+ylab("AUC")+geom_point()
+error_vs_size<- all_models_performance %>% filter (metric_type=='all_data', dataset =='test', analysis_type=='full_standard')
+error_vs_size_plt = error_vs_size %>% mutate(n_MSAs = sample_frac*2007) %>% ggplot(aes(x=n_MSAs, y = mcc))+geom_line()+ expand_limits(y = 0)+xlab("Number of MSAs")+ theme(text = element_text(size = 13))+ylab("MCC")+geom_point()
 
@@
 ```
+```{r}
+test_group_performance %>% filter (grouping_col_name =='feature_msa_n_loci') 
+```
+
 Group performance
@@
 
-Pypythia_error<-test_group_performance %>% filter (grouping_col_name =='msa_difficulty_group') %>% ggplot(aes(y=grouping_col, x=Balanced_Accuracy))+geom_col()+ expand_limits(y = c(0.5,1))+ylab('MSA\ndifficulty')+ xlab('Balanced\naccuracy')+theme(text = element_text(size = 11))
+Pypythia_error<-test_group_performance %>% filter (grouping_col_name =='msa_difficulty_group') %>%  mutate(grouping_col = factor(grouping_col, levels = c('(0.00328, 0.303]','(0.303, 0.538]','(0.538, 0.571]','	(0.571, 0.865]'))) %>%
+ggplot(aes(y=grouping_col, x=Balanced_Accuracy))+geom_col()+ expand_limits(y = c(0.5,1))+ylab('MSA\ndifficulty')+ xlab('Balanced\naccuracy')+theme(text = element_text(size = 11))
 
-n_seq_error<-test_group_performance %>% filter (grouping_col_name =='n_seq_group')  %>% ggplot(aes(y=grouping_col, x=Balanced_Accuracy))+geom_col()+ expand_limits(y = c(0.5,1))+ylab('Number\nof sequences')+xlab('Balanced\naccuracy')+theme(text = element_text(size = 11))
+n_seq_error<-test_group_performance %>% filter (grouping_col_name =='n_seq_group')   %>%  mutate(grouping_col = factor(grouping_col, levels = c('(4.999, 69.0]','(69.0, 129.0]','129.0, 236.0]','	(236.0, 827.0]]'))) %>% ggplot(aes(y=grouping_col, x=Balanced_Accuracy))+geom_col()+ expand_limits(y = c(0.5,1))+ylab('Number\nof sequences')+xlab('Balanced\naccuracy')+theme(text = element_text(size = 11))
 
-#>%  mutate(grouping_col = factor(grouping_col, levels = c('(99.999, 219.0]','(219.0, 546.0]','(546.0, 1519.0]','(1519.0, 16115.0]')))
-
-n_loci_error<-test_group_performance %>% filter (grouping_col_name =='feature_msa_n_loci') %>% ggplot(aes(y=grouping_col, x=Balanced_Accuracy))+geom_col()+ expand_limits(y = c(0.5,1))+ylab('Number\nof positions')+xlab('Balanced\naccuracy')+theme(text = element_text(size = 11))
+n_loci_error<-test_group_performance %>% filter (grouping_col_name =='feature_msa_n_loci') %>%  mutate(grouping_col = factor(grouping_col, levels = c('(4.999, 69.0]','(69.0, 129.0]','129.0, 236.0]','	(236.0, 827.0]]')))  %>% ggplot(aes(y=grouping_col, x=Balanced_Accuracy))+geom_col()+ expand_limits(y = c(0.5,1))+ylab('Number\nof positions')+xlab('Balanced\naccuracy')+theme(text = element_text(size = 11))
 
@@
 ```
-Validation_data
+
 ```{r}
-validation_data %>% dplyr::select (boot_run_time, extraction_running_time,nni_ll_eval_running_time)
+cor.test( enriched_test$feature_msa_pypythia_msa_difficulty,  enriched_test$prob_predictions)
+cor.test( enriched_test$feature_msa_n_seq, enriched_test$prob_predictions)
+cor.test( enriched_test$feature_msa_n_loci, enriched_test$prob_predictions)
+```
 
+```{r}
+running_time_data<- enriched_test%>% dplyr::select (boot_run_time,extraction_running_time,nni_ll_eval_running_time) %>% mutate (diff1 = boot_run_time/extraction_running_time, diff2 = boot_run_time/(extraction_running_time+nni_ll_eval_running_time
+)) 
+summary(running_time_data %>% pull (diff1))
+summary(running_time_data %>% pull (diff2))
+summary(running_time_data %>% pull (extraction_running_time))
+summary(running_time_data %>% pull (nni_ll_eval_running_time))
+#enriched_test %>% dplyr::select (boot_run_time, extraction_runni
+```
+
+```{r}
+
+running_time_data<- enriched_test%>% dplyr::select (program_boot_run_time,extraction_running_time,nni_ll_eval_running_time) %>% mutate (diff1 = program_boot_run_time/extraction_running_time, diff2 = program_boot_run_time/(extraction_running_time+nni_ll_eval_running_time
+)) 
+summary(running_time_data %>% pull (diff1))
+summary(running_time_data %>% pull (diff2))
+summary(running_time_data %>% pull (extraction_running_time))
+summary(running_time_data %>% pull (nni_ll_eval_running_time))
+summary(running_time_data %>% pull (program_boot_run_time))
+#enriched_test %>% dplyr::select (boot_run_time, extraction_running_time, nni_ll_eval_running_time)
+```
+```{r}
+summary(validation_data$feature_n_unique_seq)
+summary(enriched_test$feature_msa_n_seq)
 ```

### 2023-11-15 ML_pipeline/ML_boot_edit.py

@@ -28,7 +28,7 @@ def per_tree_analysis(test, features,model):
         curr_X_test = tree_data_test[[col for col in test.columns if col in features]]
         #prob_predictions = model.predict_proba(curr_X_test)[:, 1]
         predictions = model.predict(( curr_X_test))
-        metrics = model_evaluation_metrics(curr_y_true, predictions, None)
+        metrics = model_evaluation_metrics(curr_y_true, None)
         balanced_accuracies.append(metrics["balanced_accuracy_score"])
     plt.hist(balanced_accuracies)
     plt.show()
@@ -136,7 +136,7 @@ def ML_pipeline(program_data, bootstrap_cols, cpus_per_main_job, working_dir, sa
     groups = train["tree_id"]
     logging.info(f"Number of different trees is {len(program_data['tree_id'].unique())}")
     full_features = [col for col in program_data.columns if
-                'feature' in col and 'msa_entropy' not in col and col not in bootstrap_cols and 'll_diff_norm' not in col and 'column_variance' not in col]  # +['partition_branch_vs_mean','partition_branch','partition_size','partition_size_ratio','partition_divergence','divergence_ratio']
+                'feature' in col and 'msa_entropy' not in col and col not in bootstrap_cols and 'column_variance' not in col]  # +['partition_branch_vs_mean','partition_branch','partition_size','partition_size_ratio','partition_divergence','divergence_ratio']
     logging.info(f"Full features are: {full_features}")
     logging.info(f"Evaluating full standard model- including nni feautres, number of features is {len(full_features)}")
     full_model_working_dir = os.path.join(working_dir,'full_model')
@@ -188,11 +188,12 @@ def main():
     parser.add_argument('--cpus_per_main_job', type = int, default=1)
     parser.add_argument('--sample_fracs', type = str, default='0.25_0.5_1')
     parser.add_argument('--inc_sample_fracs', action='store_true', default=False)
-    parser.add_argument('--reunite_val_data', action='store_true', default=True)
+    parser.add_argument('--reunite_val_data', action='store_true', default=False)
     parser.add_argument('--reunite_training_data', action='store_true', default=False)
     parser.add_argument('--main_data_folder',type = str, default = '/Users/noa/Workspace/bootstrap_results/remote_results/full_data')
     parser.add_argument('--validation_data_folder', type=str,
                         default='/Users/noa/Workspace/bootstrap_results/remote_results/validation_data')
[…]

### 2023-11-15 ML_utils/ML_algorithms_and_hueristics.py

@@ -18,6 +18,7 @@ from sklearn.linear_model import LogisticRegression
 from sklearn.linear_model import SGDClassifier
 from sklearn.preprocessing import StandardScaler
 from sklearn.pipeline import make_pipeline
+from sklearn.metrics import confusion_matrix
 
 
 def regression_per_group(df):
@@ -29,7 +30,7 @@ def regression_per_group(df):
 def performance_per_group(df):
     y_pred = list(df.y_pred)
     y = list(df.y)
-    return balanced_accuracy_score(y, y_pred)
+    return matthews_corrcoef(y, y_pred)
 
 
 def AUPRC_per_group(df):
@@ -43,7 +44,7 @@ def score_func(y, y_pred, classification, groups_data):
     for group in groups_data:
         df = pd.DataFrame({'y': y, 'y_pred': y_pred, 'grouping_col': groups_data[group]})
         if classification:
-            df = df.groupby('grouping_col').apply(performance_per_group).reset_index(name='Balanced_Accuracy')
+            df = df.groupby('grouping_col').apply(performance_per_group).reset_index(name='mcc')
         else:
             df = df.groupby('grouping_col').apply(regression_per_group).reset_index(name='R2')
         df["grouping_col_name"] = group
@@ -126,27 +127,27 @@ def calibration_plot(model, test_data, y_test):
     pyplot.show()
 
 def enrich_with_single_feature_metrics(var_impt, train_X, y_train, test_X, y_test):
-    mcc_scores = []
-    bacc_scores = []
-    for feature in var_impt.index:
-            #lg = LogisticRegression(random_state=0).fit(train_X[[feature]], y_train)
-            lg = lightgbm.LGBMClassifier(importance_type='gain').fit(train_X[[feature]], y_train)
-            pred = lg.predict(test_X[[feature]])
-            mcc = matthews_corrcoef(y_test, pred)
-            bacc = balanced_accuracy_score(y_test, pred)
-            mcc_scores.append(mcc)
-            bacc_scores.append(bacc)
-    var_impt["mcc"] = mcc_scores
-    var_impt["balanced_accuracy"] = bacc_scores
+    mcc_thresholds = [0.5,0.9,0.95,0.99]
+    for threshold in mcc_thresholds:
+        curr_mcc_scores = []
+        for feature in var_impt.index:
+                #lg = LogisticRegression(random_state=0).fit(train_X[[feature]], y_train)
+                lg = lightgbm.LGBMClassifier(importance_type='gain').fit(train_X[[feature]], y_train)
[…]

### 2023-11-15 simulations_generation/simulations_argparser.py

@@ -20,7 +20,7 @@ def main_parser():
     parser.add_argument('--calc_features', action='store_true', default=True)
     parser.add_argument('--pre_chosen_tree_ids',action='store_true', default= False )
     parser.add_argument('--n_k_means_clusters',type=int, default =3)
-    parser.add_argument('--model_modes', type=str, default='standard') #'mispecification
+    parser.add_argument('--model_modes', type=str, default='upgrade') #'mispecification
     return parser
 
 

### 2023-11-15 ML_pipeline/ML_boot_edit.py
new changes

@@ -206,7 +206,7 @@ def main():
                                                                 name=f'simulations_df_{program}', n_jobs=1000)
         else:
             logging.info(f"Using existing training data in {training_data_path} ")
-            program_data = pd.read_csv(training_data_path,sep='\t')
+            program_data = pd.read_csv(training_data_path,sep='\t')#.sample(frac = 0.1)
         transform_data(program_data)
         validation_data_path = os.path.join(args.validation_data_folder, f'simulations_df_{program}.tsv')
         if args.reunite_val_data or not os.path.exists(validation_data_path):

### 2023-11-15 ML_utils/ML_algorithms_and_hueristics.py

@@ -180,8 +180,8 @@ def overall_model_performance_analysis(working_dir,model, data_dict, name,extrac
             predictions = model['calibrated_model'].predict((model['selector']).transform(data_dict[dataset]["X"]))
         else:
             raw_bootstrap_values = data_dict[dataset]["X"].drop(['ignore'],axis=1).iloc[:,0]
-            prob_predictions = raw_bootstrap_values / 100
-            predictions = raw_bootstrap_values > 0.5
+            prob_predictions = raw_bootstrap_values
+            predictions = raw_bootstrap_values >= 0.5
         true_label = data_dict[dataset]["y"]
         evaluation_metrics = model_evaluation_metrics(true_label, prob_predictions)
         evaluation_metrics["dataset"] = dataset

### 2023-11-17 ML_pipeline/ML_boot_edit.py
new changes to bootstrap edit

@@ -86,13 +86,13 @@ def bootstrap_model_pipeline(working_dir,train,test,validation_dict,features,boo
                                path=os.path.join(working_dir, f'model_only_boot'),
                                classifier=True, model='lightgbm', calibrate=True, name=name,
                                large_grid=large_grid, do_RFE=do_RFE,
-                               n_cv_folds=5)
+                               n_cv_folds=3)
     model_inc_boot = ML_model(inc_boot_dict["train"]["X"], groups, inc_boot_dict["train"]["y"], n_jobs=cpus_per_main_job,
                               path=os.path.join(working_dir, f'model_inc_boot'
                               f''),
                               classifier=True, model='lightgbm', calibrate=True, name=name, large_grid=large_grid,
                               do_RFE=do_RFE,
-                              n_cv_folds=5)
+                              n_cv_folds=3)
 
     raw_boot_performance,groups_data_raw_boot = overall_model_performance_analysis(working_dir, None, only_boot_dict,
                                                               name="raw_only_boot",extract_predictions = False)
@@ -136,7 +136,7 @@ def ML_pipeline(program_data, bootstrap_cols, cpus_per_main_job, working_dir, sa
     groups = train["tree_id"]
     logging.info(f"Number of different trees is {len(program_data['tree_id'].unique())}")
     full_features = [col for col in program_data.columns if
-                'feature' in col and 'msa_entropy' not in col and col not in bootstrap_cols and 'column_variance' not in col]  # +['partition_branch_vs_mean','partition_branch','partition_size','partition_size_ratio','partition_divergence','divergence_ratio']
+                'feature' in col and 'msa_entropy' not in col and col not in bootstrap_cols and 'column_variance' not in col]# +['partition_branch_vs_mean','partition_branch','partition_size','partition_size_ratio','partition_divergence','divergence_ratio']
     logging.info(f"Full features are: {full_features}")
     logging.info(f"Evaluating full standard model- including nni feautres, number of features is {len(full_features)}")
     full_model_working_dir = os.path.join(working_dir,'full_model')
@@ -172,11 +172,42 @@ def ML_pipeline(program_data, bootstrap_cols, cpus_per_main_job, working_dir, sa
 
 
 
+def get_n_free_parameters(x):
+    if x=='GTR':
+        n = 8
+    elif x=='K80':
+        n=2
+    elif x=='JC':
+        n = 1
[…]

### 2023-11-17 ML_utils/ML_algorithms_and_hueristics.py

@@ -77,7 +77,7 @@ def ML_model(X_train, groups, y_train, n_jobs, path, classifier=False, model='li
         group_splitter = list(GroupKFold(n_splits=n_cv_folds).split(X_train, y_train.ravel(), groups=groups))
         if classifier:
             if model=='lightgbm':
-                model =lightgbm.LGBMClassifier(importance_type='gain')#importance_type='gain'
+                model =lightgbm.LGBMClassifier(importance_type='gain',metric='binary_logloss')#importance_type='gain'
                 param_grid = LIGHTGBM_CLASSIFICATION_PARAM_GRID
                 if large_grid:
                     param_grid.update(GENERAL_PARAM_GRID)

### 2023-11-17 programs/raxml.py

@@ -318,7 +318,7 @@ def standard_raxml_search(curr_run_directory, msa_path, prefix,model,n_pars, n_r
 
 
 def raxml_optimize_trees_for_given_msa(full_data_path, ll_on_data_prefix, tree_file,
-                                       curr_run_directory, model, opt_model_and_brlen=True,n_cpus = 1, n_workers = 'auto'
+                                       curr_run_directory, model, opt_model_and_brlen=True,n_cpus = 1, n_workers = 'auto', return_opt_tree = False
                                        ):
     curr_run_directory = os.path.join(curr_run_directory, ll_on_data_prefix)
     if os.path.exists(curr_run_directory):
@@ -332,13 +332,16 @@ def raxml_optimize_trees_for_given_msa(full_data_path, ll_on_data_prefix, tree_f
         raxml_exe_path=RAXML_NG_EXE, msa_path=full_data_path, tree_file=tree_file, seed=SEED,
         prefix=prefix, brlen_command=brlen_command, model=model, n_cpus = n_cpus, n_workers = n_workers)
     #optimized_trees_path = prefix + ".raxml.mlTrees"
-    #best_tree_path = prefix + ".raxml.bestTree"
+    best_tree_path = prefix + ".raxml.bestTree"
     raxml_log_file = prefix + ".raxml.log"
     execute_command_and_write_to_log(compute_ll_run_command)
     trees_ll_on_data = extract_param_from_raxmlNG_log(raxml_log_file, "ll")
     #tree_alpha = extract_param_from_raxmlNG_log(raxml_log_file, "alpha")
     #optimized_trees_final_path = optimized_trees_path if os.path.exists(optimized_trees_path) else best_tree_path
-    return trees_ll_on_data#, tree_alpha, optimized_trees_final_path
+    if return_opt_tree:
+        return trees_ll_on_data,best_tree_path
+    else:
+        return trees_ll_on_data#, tree_alpha, optimized_trees_final_path
 
 def raxml_extract_sitelh(sitelh_file):
     logging.debug("Extracting sitelh from sitelh_file in {}".format(sitelh_file))

### 2023-11-17 simulation_edit/bootstrap_edit.py

@@ -14,9 +14,9 @@ from side_code.basic_trees_manipulation import *
 from side_code.MSA_manipulation import bootstrap_MSA,get_MSA_seq_names, add_unique_seq
 from side_code.spr_prune_and_regraft import *
 from side_code.file_handling import *
-from programs.raxml import raxml_compute_tree_per_site_ll, generate_n_tree_topologies,remove_redundant_sequences
+from programs.raxml import raxml_compute_tree_per_site_ll, generate_n_tree_topologies,remove_redundant_sequences,raxml_optimize_trees_for_given_msa
 from side_code.code_submission import execute_command_and_write_to_log
-
+from math import exp
 
 
 
@@ -156,6 +156,7 @@ def generate_partition_statistics(node, mle_tree_obj, extra_tree_groups, extra_b
         node_cp = mle_tree_ete_cp & node.name
         removed_node = node_cp.detach()
         mean_bl = np.mean(get_branch_lengths(mle_tree_obj))
+        var_bl = np.var(get_branch_lengths(mle_tree_obj))
         remaining_tree = mle_tree_ete_cp
         min_partition_divergence = min(get_tree_divergence(removed_node), get_tree_divergence(remaining_tree))
         divergence_ratio = min_partition_divergence/total_tree_divergence
@@ -164,7 +165,7 @@ def generate_partition_statistics(node, mle_tree_obj, extra_tree_groups, extra_b
         neighboring_nodes= get_neighboring_nodes(mle_tree_obj, node.name)
         childs_brlen = [c_node.dist for c_node in neighboring_nodes]
 
-        statistics.update({'feature_mean_neighbor_brlen': np.mean(childs_brlen),'feature_min_neighbor_brlen': np.min(childs_brlen),'feature_partition_branch': node.dist,'feature_partition_branch_vs_mean': node.dist/mean_bl,'bootstrap_support': node.support, 'true_support': true_support, 'true_binary_support': true_binary_support, 'feature_partition_size': min_partition_size,'feature_partition_size_ratio': partition_size_ratio,
[…]

### 2023-11-17 simulations_generation/simulations.py

@@ -15,7 +15,7 @@ from programs.IQTREE import iqtree_pipeline
 from simulations_generation.simulations_argparser import  job_parser
 from simulations_generation.msa_features import get_msa_stats
 from side_code.MSA_manipulation import get_MSA_seq_names
-from simulation_edit.bootstrap_edit import msa_path_bootstrap_analysis
+from simulation_edit.bootstrap_edit import msa_path_edit_analysis
 from side_code.file_handling import create_dir_if_not_exists, create_or_clean_dir
 from programs.raxml_grove import RAxML_grove_tree_simulation
 import pandas as pd
@@ -66,28 +66,22 @@ def obtain_tree_search_bootstrap_results(msa_path, tree_sim_dict, curr_run_dir, 
 
 
 def add_features_to_data(msa_path,msa_dir,boot_tree_raxml_metrics,boot_tree_iqtree_metrics,boot_tree_fasttree_metrics,tree_sim_dict):
-        curr_boot_tree_raxml_df, raxml_feature_running_time, raxml_ll_eval_running_time = msa_path_bootstrap_analysis(msa_path,msa_dir, mle_path=
+        curr_boot_tree_raxml_df = msa_path_edit_analysis(msa_path, msa_dir, mle_path=
         boot_tree_raxml_metrics['final_tree_topology_path'], true_tree_path=tree_sim_dict['true_tree_path'],
-                                                                                          program='raxml',
-                                                                                          bootstrap_tree_details_dict=boot_tree_raxml_metrics,
-                                                                                          n_pars=100)
-        curr_boot_tree_raxml_df["extraction_running_time"] = raxml_feature_running_time
-        curr_boot_tree_raxml_df["nni_ll_eval_running_time"] = raxml_ll_eval_running_time
+                                                                                                                 program='raxml',
+                                                                                                                 bootstrap_tree_details_dict=boot_tree_raxml_metrics,
+                                                                                                                 n_pars=100)
 
-        curr_boot_tree_iqtree_df, iqtree_feature_running_time,iqtree_ll_eval_running_time = msa_path_bootstrap_analysis(msa_path,msa_dir, mle_path=
+        curr_boot_tree_iqtree_df = msa_path_edit_analysis(msa_path, msa_dir, mle_path=
         boot_tree_iqtree_metrics['final_tree_ultrafast'], true_tree_path=tree_sim_dict['true_tree_path'],
-                                                                                            program='iqtree',
[…]

### 2023-11-17 simulation_edit/parallel_boot_edit.py
small fix

@@ -70,6 +70,10 @@ def main():
     raw_data_raxml["program"] = 'raxml'
     raw_data_iqtree["program"] = 'iqtree'
     raw_data_fasttree["program"] = 'fasttree'
+    if 'tree_search_model' not in raw_data_raxml.columns:
+        raw_data_raxml['tree_search_model'] = raw_data_raxml['model_short']
+
+
 
 
     raw_data = pd.concat([raw_data_fasttree, raw_data_iqtree, raw_data_raxml])

### 2023-11-17 simulation_edit/parallel_boot_edit.py
small fix

@@ -72,6 +72,8 @@ def main():
     raw_data_fasttree["program"] = 'fasttree'
     if 'tree_search_model' not in raw_data_raxml.columns:
         raw_data_raxml['tree_search_model'] = raw_data_raxml['model_short']
+        raw_data_iqtree['tree_search_model'] = raw_data_iqtree['model_short']
+        raw_data_fasttree['tree_search_model'] = raw_data_fasttree['model_short']
 
 
 

### 2023-11-18 simulation_edit/bootstrap_edit.py
fix

@@ -225,10 +225,11 @@ def get_bootstrap_and_tree_groups(program, bootstrap_tree_details,mle_path,garba
         add_internal_names(aBayes_ete)
         extra_boot = {'aLRT_iqtree': aLRT_ete, 'aBayes_iqtree': aBayes_ete}
     elif program == 'fasttree':
-        standard_bootstrap = bootstrap_tree_details['standard_bootstrap']
-        standard_ete = Tree(newick=standard_bootstrap, format=0)
-        add_internal_names(standard_ete)
-        extra_boot = {'standard_fasttree_boot': standard_ete}
+            standard_bootstrap = bootstrap_tree_details['standard_bootstrap']
+            if os.path.exists(standard_bootstrap):
+                standard_ete = Tree(newick=standard_bootstrap, format=0)
+                add_internal_names(standard_ete)
+                extra_boot = {'standard_fasttree_boot': standard_ete}
     return extra_boot, extra_tree_groups
 
 

### 2023-11-18 simulation_edit/bootstrap_edit.py
small fix

@@ -210,12 +210,12 @@ def get_bootstrap_and_tree_groups(program, bootstrap_tree_details,mle_path,garba
         parsimony_trees = trees_path.read().split("\n")[:-1]
     all_pars_ete = generate_booster_trees(mle_path, parsimony_trees, garbage_dir, tree_tmp_path)
     extra_tree_groups = {'parsimony_trees': all_pars_ete}
+    extra_boot = {}
     if program == 'raxml':
         all_mle_path = bootstrap_tree_details['all_final_tree_topologies_path']
         all_ML_nw = get_file_rows(all_mle_path)
         all_ML_ete = generate_booster_trees(mle_path, all_ML_nw, garbage_dir, tree_tmp_path)
         extra_tree_groups.update( {'all_ML_boot_raxml': all_ML_ete})
-        extra_boot = {}
     elif program == 'iqtree':
         final_tree_aLRT = bootstrap_tree_details['final_tree_aLRT']
         aLRT_ete = Tree(newick=final_tree_aLRT, format=0)
@@ -225,11 +225,12 @@ def get_bootstrap_and_tree_groups(program, bootstrap_tree_details,mle_path,garba
         add_internal_names(aBayes_ete)
         extra_boot = {'aLRT_iqtree': aLRT_ete, 'aBayes_iqtree': aBayes_ete}
     elif program == 'fasttree':
-            standard_bootstrap = bootstrap_tree_details['standard_bootstrap']
-            if os.path.exists(standard_bootstrap):
-                standard_ete = Tree(newick=standard_bootstrap, format=0)
-                add_internal_names(standard_ete)
-                extra_boot = {'standard_fasttree_boot': standard_ete}
+        standard_bootstrap = bootstrap_tree_details['standard_bootstrap']
+        if os.path.exists(standard_bootstrap):
+            standard_ete = Tree(newick=standard_bootstrap, format=0)
+            add_internal_names(standard_ete)
+            extra_boot = {'standard_fasttree_boot': standard_ete}
+
     return extra_boot, extra_tree_groups
 
 

### 2023-11-18 ML_pipeline/ML_boot_edit.py
new changes

@@ -220,11 +220,12 @@ def main():
     parser.add_argument('--sample_fracs', type = str, default='0.25_0.5_1')
     parser.add_argument('--inc_sample_fracs', action='store_true', default=False)
     parser.add_argument('--reunite_val_data', action='store_true', default=False)
+    parser.add_argument('--use_val_data', action='store_true', default=False)
     parser.add_argument('--reunite_training_data', action='store_true', default=False)
     parser.add_argument('--main_data_folder',type = str, default = '/Users/noa/Workspace/bootstrap_results/remote_results/full_data')
     parser.add_argument('--validation_data_folder', type=str,
                         default='/Users/noa/Workspace/bootstrap_results/remote_results/validation_data')
-    parser.add_argument('--sample_val', action='store_true', default=True)
+    parser.add_argument('--sample_val', action='store_true', default=False)
     args = parser.parse_args()
     log_file_path = os.path.join(args.working_dir, "ML.log")
     logging.basicConfig(filename=log_file_path, level=logging.INFO)
@@ -260,8 +261,9 @@ def main():
         sample_fracs = [float(frac) for frac in (args.sample_fracs).split('_')]
         all_model_merics = pd.DataFrame()
         validation_dict = {}
-        for model_mode in np.unique(program_validation_data["model_mode"]):
-            validation_dict[f'val_{model_mode}'] = program_validation_data.loc[program_validation_data.model_mode==model_mode].copy()
+        if args.use_val_data:
+            for model_mode in np.unique(program_validation_data["model_mode"]):
+                validation_dict[f'val_{model_mode}'] = program_validation_data.loc[program_validation_data.model_mode==model_mode].copy()
         if args.inc_sample_fracs:
             for sample_frac in sample_fracs:
                 logging.info(f"\n#Sample frac = {sample_frac}")

### 2023-11-18 ML_pipeline/ML_boot_edit.py
new change

@@ -241,7 +241,7 @@ def main():
             program_data = pd.read_csv(training_data_path,sep='\t')
         program_data=transform_data(program_data)
         validation_data_path = os.path.join(args.validation_data_folder, f'simulations_df_{program}.tsv')
-        if args.reunite_val_data or not os.path.exists(validation_data_path):
+        if args.use_val_data and (args.reunite_val_data or not os.path.exists(validation_data_path)):
             logging.info(f"Re-uniting validation data and saving to {validation_data_path}")
             program_validation_data = unify_results_across_jobs(args.validation_data_folder, name =f'simulations_df_{program}' , n_jobs = 1000)
         else:

### 2023-11-18 ML_pipeline/ML_boot_edit.py
new fix

@@ -204,6 +204,8 @@ def transform_data(df):
     df['feature_G'] = df['model_short'].str.contains('+G', regex = False)
     df['feature_I'] = df['model_short'].str.contains('+I',regex = False)
     df['feature_F'] = df['model_short'].str.contains('+F',regex = False)
+    if 'model_short' not in df.columns:
+        df['model_short'] = df['tree_search_model']
     df['model_short'] = df['model_short'].apply(lambda x: x.split('+')[0])
     df['feature_free_parameters'] = df['model_short'].apply(lambda x: get_n_free_parameters(x))
     #df = pd.get_dummies(df,prefix='feature_model_',columns=['model_short']) #
@@ -229,6 +231,7 @@ def main():
     args = parser.parse_args()
     log_file_path = os.path.join(args.working_dir, "ML.log")
     logging.basicConfig(filename=log_file_path, level=logging.INFO)
+    create_or_clean_dir(args.working_dir)
     for program in args.programs.split('_'):
         logging.info(f"Program = {program}")
         training_data_path= os.path.join(args.main_data_folder,f'simulations_df_{program}.tsv')
@@ -240,19 +243,19 @@ def main():
             logging.info(f"Using existing training data in {training_data_path} ")
             program_data = pd.read_csv(training_data_path,sep='\t')
         program_data=transform_data(program_data)
+        validation_dict = {}
         validation_data_path = os.path.join(args.validation_data_folder, f'simulations_df_{program}.tsv')
-        if args.use_val_data and (args.reunite_val_data or not os.path.exists(validation_data_path)):
-            logging.info(f"Re-uniting validation data and saving to {validation_data_path}")
-            program_validation_data = unify_results_across_jobs(args.validation_data_folder, name =f'simulations_df_{program}' , n_jobs = 1000)
-        else:
-
-            logging.info("Using existing validation data")
-            program_validation_data = pd.read_csv(validation_data_path, sep='\t')
-        if args.sample_val:
-            program_validation_data = program_validation_data[program_validation_data['true_tree_path'].str.contains("iqtree_msa_0")]
-            program_validation_data['model_short']= program_validation_data["tree_search_model"]
-
-        program_validation_data = transform_data(program_validation_data)
+        if args.use_val_data:
+            logging.info("Using validation data")
+            if (args.reunite_val_data or not os.path.exists(validation_data_path)):
+                logging.info(f"Re-uniting validation data and saving to {validation_data_path}")
[…]

### 2023-11-19 simulation_edit/parallel_boot_edit.py
important fix

@@ -70,11 +70,11 @@ def main():
                          'true_tree_path_orig','tree_folder'
                          ]
     raw_data_raxml = unify_results_across_jobs(raw_results_folder, name = 'simulations_df_raxml')
-    raw_data_raxml = raw_data_raxml[[col for col in raw_data_raxml.columns if col in remaining_columns or 'feature' in col]].drop_duplicates()
+    raw_data_raxml = raw_data_raxml[[col for col in raw_data_raxml.columns if col in remaining_columns or 'feature_msa' in col]].drop_duplicates()
     raw_data_iqtree = unify_results_across_jobs(raw_results_folder, name='simulations_df_iqtree')
-    raw_data_iqtree = raw_data_iqtree[[col for col in raw_data_iqtree.columns if col in remaining_columns or 'feature' in col]].drop_duplicates()
+    raw_data_iqtree = raw_data_iqtree[[col for col in raw_data_iqtree.columns if col in remaining_columns or 'feature_msa' in col]].drop_duplicates()
     raw_data_fasttree = unify_results_across_jobs(raw_results_folder, name='simulations_df_fasttree')
-    raw_data_fasttree = raw_data_fasttree[[col for col in raw_data_fasttree.columns if col in remaining_columns or 'feature' in col]].drop_duplicates()
+    raw_data_fasttree = raw_data_fasttree[[col for col in raw_data_fasttree.columns if col in remaining_columns or 'feature_msa' in col]].drop_duplicates()
     raw_data_raxml["program"] = 'raxml'
     raw_data_iqtree["program"] = 'iqtree'
     raw_data_fasttree["program"] = 'fasttree'

### 2023-11-19 ML_pipeline/ML_boot_edit.py
new changes

@@ -88,21 +88,23 @@ def bootstrap_model_pipeline(working_dir, train, test, validation_dict, features
                               do_RFE=do_RFE,
                               n_cv_folds=3)
 
-    raw_boot_performance, groups_data_raw_boot = overall_model_performance_analysis(working_dir, None, only_boot_dict,
+    raw_boot_performance, groups_data_raw_boot,test_predictions_raw_boot = overall_model_performance_analysis(working_dir, None, only_boot_dict,
                                                                                     name="raw_only_boot",
                                                                                     extract_predictions=False)
 
-    only_boot_performance, groups_data_only_boot = overall_model_performance_analysis(working_dir, model_only_boot,
+    only_boot_performance, groups_data_only_boot,test_predictions_only_boot = overall_model_performance_analysis(working_dir, model_only_boot,
                                                                                       only_boot_dict,
                                                                                       name="only_boot",
                                                                                       extract_predictions=False)
-    inc_boot_performance, groups_data_inc_boot = overall_model_performance_analysis(working_dir, model_inc_boot,
+    inc_boot_performance, groups_data_inc_boot,test_predictions_inc_boot = overall_model_performance_analysis(working_dir, model_inc_boot,
                                                                                     inc_boot_dict,
                                                                                     name="inc_boot",
                                                                                     extract_predictions=False)
     bootstrap_models_performance = pd.concat([raw_boot_performance, only_boot_performance, inc_boot_performance])
     bootstrap_models_performance["analysis_type"] = bootstrap_col
-    return bootstrap_models_performance
+
+    test_performance_dict = {f'{bootstrap_col}_only_boot' : test_predictions_only_boot,f'{bootstrap_col}_raw_boot' : test_predictions_raw_boot,f'{bootstrap_col}_inc_boot' : test_predictions_inc_boot}
+    return bootstrap_models_performance,test_performance_dict
 
 
 def standard_model_pipeline(train, test, validation_dict, groups, features, cpus_per_main_job, working_dir, do_RFE,
@@ -119,16 +121,17 @@ def standard_model_pipeline(train, test, validation_dict, groups, features, cpus
                      n_cv_folds=3)
 
     logging.info("Evaluating model performance")
-    model_performance, group_performance = overall_model_performance_analysis(working_dir, model, datasets_dict,
[…]

### 2023-11-19 ML_utils/ML_algorithms_and_hueristics.py

@@ -193,8 +193,8 @@ def overall_model_performance_analysis(working_dir,model, data_dict, name,extrac
         enriched_dataset["prob_predictions"] = prob_predictions
         enriched_dataset["predictions"] = predictions
         enriched_dataset["true_label"] = true_label
-        if extract_predictions and dataset=='test':
-            enriched_dataset.to_csv(os.path.join(working_dir,f"{dataset}_{name}_enriched.tsv"),sep= CSV_SEP)
+        #if extract_predictions and dataset=='test':
+        #    enriched_dataset.to_csv(os.path.join(working_dir,f"{dataset}_{name}_enriched.tsv"),sep= CSV_SEP)
 
         #per_tree_evaluation_metrics = pd.DataFrame()
         #for tree_ind in enriched_dataset["tree_id"].unique():
@@ -210,7 +210,7 @@ def overall_model_performance_analysis(working_dir,model, data_dict, name,extrac
         group_performance["dataset"] = dataset
         group_performance["name"] = name
         all_group_metrics = pd.concat([all_group_metrics,group_performance])
-    return all_metrics,all_group_metrics
+    return all_metrics,all_group_metrics,prob_predictions
 
 
 

### 2023-11-20 simulation_edit/bootstrap_edit.py
new fix

@@ -407,18 +407,22 @@ def main():
     data = pd.read_csv(args.job_data_path, sep='\t')
     all_splits = pd.DataFrame()
     log_file_path = os.path.join(args.job_work_path, "general_features.log")
-    logging.basicConfig(filename=log_file_path, level=logging.DEBUG)
-    for true_tree_path in data['true_tree_path'].unique():
+    logging.basicConfig(filename=log_file_path, level=logging.INFO)
+    logging.info(f"Number of different trees is {len(data['true_tree_path'].unique())}")
+    for i,true_tree_path in enumerate(data['true_tree_path'].unique()):
+        logging.info(f"Tree {i} out of {len(data['true_tree_path'].unique())}")
         tree_data = data.loc[data.true_tree_path == true_tree_path]
         for program in tree_data['program'].unique():
             tree_program_data = tree_data.loc[tree_data.program==program]
-            for msa_path in tree_data['msa_path'].unique():
+            for j,msa_path in tree_data['msa_path'].unique():
+                logging.info(f"MSA {j} out of {len(tree_data['msa_path'].unique())}")
                 bootstrap_tree_details = tree_program_data.loc[tree_program_data.msa_path == msa_path].head(1).squeeze()
 
                 mle_path =  bootstrap_tree_details[get_program_default_ML_tree(program)]
                 msa_splits = msa_path_edit_analysis(msa_path, args.job_work_path, mle_path, true_tree_path, program, bootstrap_tree_details.to_dict(), args.n_pars)
                 all_splits = pd.concat([all_splits, msa_splits])
                 all_splits.to_csv(args.job_final_output_path, sep='\t')
+        logging.info("Done with all trees!")
         raxml_data = all_splits.loc[all_splits.program=='raxml']
         raxml_data.to_csv(os.path.join(args.job_work_path,'simulations_df_raxml.tsv'), sep='\t')
         iqtree_data = all_splits.loc[all_splits.program == 'iqtree']

### 2023-11-20 simulation_edit/bootstrap_edit.py
change boot edit

@@ -155,8 +155,7 @@ def generate_partition_statistics(node, mle_tree_obj, extra_tree_groups, extra_b
         mle_tree_ete_cp = mle_tree_obj.copy()
         node_cp = mle_tree_ete_cp & node.name
         removed_node = node_cp.detach()
-        mean_bl = np.mean(get_branch_lengths(mle_tree_obj))
-        var_bl = np.var(get_branch_lengths(mle_tree_obj))
+        branch_lengths = get_branch_lengths(mle_tree_obj)
         remaining_tree = mle_tree_ete_cp
         min_partition_divergence = min(get_tree_divergence(removed_node), get_tree_divergence(remaining_tree))
         divergence_ratio = min_partition_divergence/total_tree_divergence
@@ -165,7 +164,7 @@ def generate_partition_statistics(node, mle_tree_obj, extra_tree_groups, extra_b
         neighboring_nodes= get_neighboring_nodes(mle_tree_obj, node.name)
         childs_brlen = [c_node.dist for c_node in neighboring_nodes]
 
-        statistics.update({'feature_mean_bl': mean_bl,'feature_var_bl': var_bl,'feature_mean_neighbor_brlen': np.mean(childs_brlen),'feature_min_neighbor_brlen': np.min(childs_brlen),'feature_partition_branch': node.dist,'feature_partition_branch_vs_mean': node.dist/mean_bl,'bootstrap_support': node.support, 'true_support': true_support, 'true_binary_support': true_binary_support, 'feature_partition_size': min_partition_size,'feature_partition_size_ratio': partition_size_ratio,
+        statistics.update({'feaure_25_pct_bl':np.percentile(branch_lengths,25),'feaure_75_pct_bl':np.percentile(branch_lengths,25), 'feature_mean_bl': np.mean(branch_lengths),'feature_var_bl': np.var(branch_lengths),'feature_mean_neighbor_brlen': np.mean(childs_brlen),'feature_min_neighbor_brlen': np.min(childs_brlen),'feature_partition_branch': node.dist,'feature_partition_branch_vs_mean': node.dist/np.mean(branch_lengths),'bootstrap_support': node.support, 'true_support': true_support, 'true_binary_support': true_binary_support, 'feature_partition_size': min_partition_size,'feature_partition_size_ratio': partition_size_ratio,
                     'feature_partition_divergence': min_partition_divergence, 'feature_divergence_ratio': divergence_ratio})
         for b_method in extra_boot_ete:
[…]

### 2023-11-20 simulation_edit/bootstrap_edit.py
fix

@@ -164,7 +164,7 @@ def generate_partition_statistics(node, mle_tree_obj, extra_tree_groups, extra_b
         neighboring_nodes= get_neighboring_nodes(mle_tree_obj, node.name)
         childs_brlen = [c_node.dist for c_node in neighboring_nodes]
 
-        statistics.update({'feaure_25_pct_bl':np.percentile(branch_lengths,25),'feaure_75_pct_bl':np.percentile(branch_lengths,25), 'feature_mean_bl': np.mean(branch_lengths),'feature_var_bl': np.var(branch_lengths),'feature_mean_neighbor_brlen': np.mean(childs_brlen),'feature_min_neighbor_brlen': np.min(childs_brlen),'feature_partition_branch': node.dist,'feature_partition_branch_vs_mean': node.dist/np.mean(branch_lengths),'bootstrap_support': node.support, 'true_support': true_support, 'true_binary_support': true_binary_support, 'feature_partition_size': min_partition_size,'feature_partition_size_ratio': partition_size_ratio,
+        statistics.update({'feature_25_pct_bl':np.percentile(branch_lengths,25),'feature_75_pct_bl':np.percentile(branch_lengths,25), 'feature_mean_bl': np.mean(branch_lengths),'feature_var_bl': np.var(branch_lengths),'feature_mean_neighbor_brlen': np.mean(childs_brlen),'feature_min_neighbor_brlen': np.min(childs_brlen),'feature_partition_branch': node.dist,'feature_partition_branch_vs_mean': node.dist/np.mean(branch_lengths),'bootstrap_support': node.support, 'true_support': true_support, 'true_binary_support': true_binary_support, 'feature_partition_size': min_partition_size,'feature_partition_size_ratio': partition_size_ratio,
                     'feature_partition_divergence': min_partition_divergence, 'feature_divergence_ratio': divergence_ratio})
         for b_method in extra_boot_ete:
             statistics.update({f'feature_{b_method}_support':(extra_boot_ete[b_method]&node.name).support})
@@ -413,7 +413,7 @@ def main():
         tree_data = data.loc[data.true_tree_path == true_tree_path]
         for program in tree_data['program'].unique():
             tree_program_data = tree_data.loc[tree_data.program==program]
-            for j,msa_path in tree_data['msa_path'].unique():
+            for j,msa_path in enumerate(tree_data['msa_path'].unique()):
[…]

### 2023-11-20 simulation_edit/bootstrap_edit.py
new change

@@ -412,6 +412,7 @@ def main():
         logging.info(f"Tree {i} out of {len(data['true_tree_path'].unique())}")
         tree_data = data.loc[data.true_tree_path == true_tree_path]
         for program in tree_data['program'].unique():
+            logging.info(f"Program = {program}")
             tree_program_data = tree_data.loc[tree_data.program==program]
             for j,msa_path in enumerate(tree_data['msa_path'].unique()):
                 logging.info(f"MSA {j} out of {len(tree_data['msa_path'].unique())}")

### 2023-11-20 ML_pipeline/ML_boot_edit.py
new change

@@ -260,7 +260,7 @@ def main():
         if args.reunite_training_data or not os.path.exists(training_data_path):
             logging.info(f"Re-uniting training data and saving to {training_data_path}")
             program_data = unify_results_across_jobs(args.main_data_folder,
-                                                     name=f'simulations_df_{program}', n_jobs=1000).sample(n=300)
+                                                     name=f'simulations_df_{program}', n_jobs=1000)
         else:
             logging.info(f"Using existing training data in {training_data_path} ")
             program_data = pd.read_csv(training_data_path, sep='\t')

### 2023-11-20 ML_pipeline/ML_boot_edit.py
reset index

@@ -183,8 +183,8 @@ def ML_pipeline(program_data, bootstrap_cols, cpus_per_main_job, working_dir, sa
     all_models_performance["sample_frac"] = sample_frac
     group_performance_full["sample_frac"] = sample_frac
     if sample_frac==-1:
-        enriched_test = test.copy()
-        test_performance_df = pd.DataFrame(test_performance_dict)
+        enriched_test = test.copy().reset_index()
+        test_performance_df = pd.DataFrame(test_performance_dict).reset_index()
         test_with_predictions = pd.concat([enriched_test,test_performance_df],axis=1)
     return all_models_performance, group_performance_full,test_with_predictions
 

### 2023-11-20 ML_pipeline/ML_boot_edit.py
fix

@@ -50,7 +50,7 @@ def get_bootstrap_col(program):
     elif program == 'iqtree':
         bootstrap_cols = ['bootstrap_support', 'feature_aLRT_iqtree_support', 'feature_aBayes_iqtree_support']
     elif program == 'fasttree':
-        bootstrap_cols = ['feature_standard_fasttree_boot_support', 'bootstrap_support']
+        bootstrap_cols = ['bootstrap_support']
     return bootstrap_cols
 
 

### 2023-11-20 ML_pipeline/ML_boot_edit.py
new changes

@@ -103,7 +103,7 @@ def bootstrap_model_pipeline(working_dir, train, test, validation_dict, features
     bootstrap_models_performance = pd.concat([raw_boot_performance, only_boot_performance, inc_boot_performance])
     bootstrap_models_performance["analysis_type"] = bootstrap_col
 
-    test_performance_dict = {f'{bootstrap_col}_only_boot' : test_predictions_only_boot,f'{bootstrap_col}_raw_boot' : test_predictions_raw_boot,f'{bootstrap_col}_inc_boot' : test_predictions_inc_boot}
+    test_performance_dict = {f'predictions_{bootstrap_col}_only_boot' : test_predictions_only_boot,f'predictions_{bootstrap_col}_raw_boot' : test_predictions_raw_boot,f'predictions_{bootstrap_col}_inc_boot' : test_predictions_inc_boot}
     return bootstrap_models_performance,test_performance_dict
 
 
@@ -184,7 +184,7 @@ def ML_pipeline(program_data, bootstrap_cols, cpus_per_main_job, working_dir, sa
     group_performance_full["sample_frac"] = sample_frac
     if sample_frac==-1:
         enriched_test = test.copy().reset_index()
-        test_performance_df = pd.DataFrame(test_performance_dict).reset_index()
+        test_performance_df = pd.DataFrame(all_test_perdictions).reset_index()
         test_with_predictions = pd.concat([enriched_test,test_performance_df],axis=1)
     return all_models_performance, group_performance_full,test_with_predictions
 

### 2023-11-20 ML_utils/ML_algorithms_and_hueristics.py

@@ -174,6 +174,7 @@ def overall_model_performance_analysis(working_dir,model, data_dict, name,extrac
         var_impt.to_csv(vi_path, sep=CSV_SEP)
     all_metrics = pd.DataFrame()
     all_group_metrics = pd.DataFrame()
+    prob_predictions_per_dataset = {}
     for dataset in data_dict:
         if model:
             prob_predictions =  model['calibrated_model'].predict_proba((model['selector']).transform(data_dict[dataset]["X"]))[:, 1]
@@ -210,7 +211,8 @@ def overall_model_performance_analysis(working_dir,model, data_dict, name,extrac
         group_performance["dataset"] = dataset
         group_performance["name"] = name
         all_group_metrics = pd.concat([all_group_metrics,group_performance])
-    return all_metrics,all_group_metrics,prob_predictions
+        prob_predictions_per_dataset[dataset] = prob_predictions
+    return all_metrics,all_group_metrics, prob_predictions_per_dataset["test"]
 
 
 

### 2023-11-21 simulation_edit/parallel_boot_edit.py
adding model mode

@@ -67,7 +67,7 @@ def main():
     remaining_columns = ['all_final_tree_topologies_path','tree_search_model','model_short','model',
                          'final_tree_aLRT', 'final_tree_aBayes', 'final_tree_ultrafast','final_tree_topology_path',
                          'boot_run_time','sh_bootstrap','standard_bootstrap','true_tree_path','msa_path','tree_id',
-                         'true_tree_path_orig','tree_folder'
+                         'true_tree_path_orig','tree_folder','model_mode'
                          ]
     raw_data_raxml = unify_results_across_jobs(raw_results_folder, name = 'simulations_df_raxml')
     raw_data_raxml = raw_data_raxml[[col for col in raw_data_raxml.columns if col in remaining_columns or 'feature_msa' in col]].drop_duplicates()

### 2023-11-22 R_code/paper_analysis.Rmd
small fix

@@
 fasttree_vi =  read_tsv(paste(prefix,"ML_results/fasttree/final_model/full_model/model_standard_vi.tsv", sep = ""))
-fasttree_enriched_test = read_tsv(paste(prefix,"ML_results/fasttree/final_model/full_model/test_model_standard_enriched.tsv", sep = ""))
+fasttree_enriched_test = read_tsv(paste(prefix,"ML_results/fasttree/test_with_predictions.tsv", sep = ""))
 fasttree_validation = read_tsv(paste(prefix,"validation_data/simulations_df_fasttree.tsv", sep = ""))
@@
 iqtree_vi =  read_tsv(paste(prefix,"ML_results/iqtree/final_model/full_model/model_standard_vi.tsv", sep = ""))
-iqtree_enriched_test = read_tsv(paste(prefix,"ML_results/iqtree/final_model/full_model/test_model_standard_enriched.tsv", sep = ""))
+iqtree_enriched_test = read_tsv(paste(prefix,"ML_results/iqtree/test_with_predictions.tsv", sep = ""))
 iqtree_validation = read_tsv(paste(prefix,"validation_data/simulations_df_iqtree.tsv", sep = ""))
@@
 raxml_vi =  read_tsv(paste(prefix,"ML_results/raxml/final_model/full_model/model_standard_vi.tsv", sep = ""))
-raxml_enriched_test = read_tsv(paste(prefix,"ML_results/raxml/final_model/full_model/test_model_standard_enriched.tsv", sep = ""))
+raxml_enriched_test = read_tsv(paste(prefix,"ML_results/raxml//test_with_predictions.tsv", sep = ""))
 raxml_validation = read_tsv(paste(prefix,"validation_data/simulations_df_raxml.tsv", sep = ""))
@@
 ```{r}
-program = 'fasttree'
+program = 'iqtree'
 if (program=='fasttree') {
@@
 
+```{r}
+
+summm<- enriched_test %>% group_by(tree_id) %>% summarise(mean_true = mean(true_binary_support), mean_pred = mean(prob_predictions), mean_bayes = mean(feature_aBayes_iqtree_support), mean_bayes_cal = mean(predictions_feature_aBayes_iqtree_support_only_boot),
+                                                mean_alrt = mean(feature_aLRT_iqtree_support), mean_alrt_cal = mean(predictions_feature_aLRT_iqtree_support_only_boot)
+
+                                                )
+
+summm%>% ggplot(aes(x=mean_true,y=mean_bayes))+geom_point()+geom_abline(slope=1, intercept = 0)
+summm%>% ggplot(aes(x=mean_true,y=mean_alrt))+geom_point()+geom_abline(slope=1, intercept = 0)
+
+summm %>% ggplot(aes(x=mean_true,y=mean_bayes_cal))+geom_point()+geom_abline(slope=1, intercept = 0)
+summm %>% ggplot(aes(x=mean_true,y=mean_alrt_cal))+geom_point()+geom_abline(slope=1, intercept = 0)
+
+summm  %>% ggplot(aes(x=mean_true,y=mean_pred))+geom_point()+geom_abline(slope=1, intercept = 0)
+
+pivoted<- summ %>% pivot_longer()
+```
+
 General performance
@@
 
-paper_performance<-final_models_performance  %>% filter (metric_type=='all_data') %>% dplyr::select (dataset,analysis_type,name,mcc,balanced_accuracy_score,precision, recall, AUC,average_precision)  %>% arrange(dataset) %>% mutate (description = case_when(analysis_type == "full_standard" & name =="model_standard" ~ "all_features",
+paper_performance<-final_models_performance  %>% filter (metric_type=='all_data') %>% dplyr::select (dataset,analysis_type,name,mcc_0.5,logloss, AUC,average_precision, tp_0.5, tn_0.5, fp_0.5, fn_0.5)  %>% arrange(dataset) %>% mutate (description = case_when(analysis_type == "full_standard" & name =="model_standard" ~ "all features",
                                 analysis_type == "fast_standard" & name =="model_standard" ~ "All features except NNI features",
@@
                                 analysis_type == "feature_aBayes_iqtree_support" & name =="inc_boot" ~ "All features+ aBayes support",
-                                )) %>% dplyr::select (dataset, description, mcc, balanced_accuracy_score,precision, recall)  %>% arrange(dataset)
+                                )) %>% mutate(TPR =`tp_0.5`/(`tp_0.5`+`fn_0.5`),TNR =`tn_0.5`/(`tn_0.5`+`fp_0.5`),FPR =`fp_0.5`/(`tn_0.5`+`fp_0.5`),FNR =`fn_0.5`/(`tp_0.5`+`fn_0.5`)) %>% dplyr::select (dataset, description,logloss, mcc_0.5,FPR,FNR, AUC,average_precision)  %>% arrange(dataset)
 
 paper_performance
+write_tsv(paper_performance,out_path)
 
@@
 ```{r}
-all_models_performance
-```
-
-```{r}
 library(predtools)
@@
 
-n_loci_error<-test_group_performance %>% filter (grouping_col_name =='feature_msa_n_loci') %>%  mutate(grouping_col = factor(grouping_col, levels = c('(4.999, 69.0]','(69.0, 129.0]','129.0, 236.0]','	(236.0, 827.0]]')))  %>% ggplot(aes(y=grouping_col, x=Balanced_Accuracy))+geom_col()+ expand_limits(y = c(0.5,1))+ylab('Number\nof positions')+xlab('Balanced\naccuracy')+theme(text = element_text(size = 11))
+n_loci_error<-test_group_performance %>% filter (grouping_col_name =='feature_msa_n_loci') %>%  mutate(grouping_col = factor(grouping_col, levels = c('(61.999, 727.0]','(69.0, 129.0]','129.0, 236.0]','	(236.0, 827.0]]')))  %>% ggplot(aes(y=grouping_col, x=Balanced_Accuracy))+geom_col()+ expand_limits(y = c(0.5,1))+ylab('Number\nof positions')+xlab('Balanced\naccuracy')+theme(text = element_text(size = 11))
 
@@
 summary(validation_data$feature_n_unique_seq)
-summary(enriched_test$feature_msa_n_seq)
+raxml_validation_data$
 ```

### 2023-11-22 ML_pipeline/ML_boot_edit.py

@@ -182,10 +182,9 @@ def ML_pipeline(program_data, bootstrap_cols, cpus_per_main_job, working_dir, sa
 
     all_models_performance["sample_frac"] = sample_frac
     group_performance_full["sample_frac"] = sample_frac
-    if sample_frac==-1:
-        enriched_test = test.copy().reset_index()
-        test_performance_df = pd.DataFrame(all_test_perdictions).reset_index()
-        test_with_predictions = pd.concat([enriched_test,test_performance_df],axis=1)
+    enriched_test = test.copy().reset_index()
+    test_performance_df = pd.DataFrame(all_test_perdictions).reset_index()
+    test_with_predictions = pd.concat([enriched_test,test_performance_df],axis=1)
     return all_models_performance, group_performance_full,test_with_predictions
 
 
@@ -231,6 +230,7 @@ def transform_data(df):
     # df = pd.get_dummies(df,prefix='feature_model_',columns=['model_short']) #
     return df
     # +[col for col in df.columns if 'msa_entropy' in col]
+
 
 
 def main():
@@ -292,7 +292,7 @@ def main():
                 logging.info(f"\n#Sample frac = {sample_frac}")
                 sample_frac_working_dir = os.path.join(working_dir, f"frac_{sample_frac}")
                 create_dir_if_not_exists(sample_frac_working_dir)
-                curr_model_metrics, groups_analysis = ML_pipeline(program_data, bootstrap_cols, args.cpus_per_main_job,
+                curr_model_metrics, groups_analysis, curr_test_with_predictions = ML_pipeline(program_data, bootstrap_cols, args.cpus_per_main_job,
                                                                   sample_frac_working_dir, sample_frac,
                                                                   compare_to_bootstrap_models=False,
                                                                   subsample_train=True, do_RFE=args.RFE,

### 2023-11-25 R_code/paper_analysis.Rmd
subsampling validation adding brier score fixing stuff

@@
 library(pROC)
+library(predtools)
+library(MASS)
+library(caret)
 ```
@@
 ```{r}
-program = 'iqtree'
+program = 'fasttree'
 if (program=='fasttree') {
@@
 
+IQTREE
 ```{r}
 
-summm<- enriched_test %>% group_by(tree_id) %>% summarise(mean_true = mean(true_binary_support), mean_pred = mean(prob_predictions), mean_bayes = mean(feature_aBayes_iqtree_support), mean_bayes_cal = mean(predictions_feature_aBayes_iqtree_support_only_boot),
-                                                mean_alrt = mean(feature_aLRT_iqtree_support), mean_alrt_cal = mean(predictions_feature_aLRT_iqtree_support_only_boot)
+summm_iqtree<- iqtree_enriched_test %>% group_by(tree_id) %>% summarise(mean_true = mean(true_binary_support), mean_pred = mean(prob_predictions), mean_bayes = mean(feature_aBayes_iqtree_support), mean_ultrafast = mean(bootstrap_support), mean_bayes_cal = mean(predictions_feature_aBayes_iqtree_support_only_boot),
+                                                mean_alrt = mean(feature_aLRT_iqtree_support), mean_alrt_cal = mean(predictions_feature_aLRT_iqtree_support_only_boot))
 
-                                                )
+bayes<-summm_iqtree%>% ggplot(aes(y=mean_true,x=mean_bayes))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("aBayes test")+theme(text = element_text(size = 11))+ coord_flip()
 
-summm%>% ggplot(aes(x=mean_true,y=mean_bayes))+geom_point()+geom_abline(slope=1, intercept = 0)
-summm%>% ggplot(aes(x=mean_true,y=mean_alrt))+geom_point()+geom_abline(slope=1, intercept = 0)
+alrt<-summm_iqtree %>% ggplot(aes(y=mean_true,x=mean_alrt))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("aLRT test ")+theme(text = element_text(size = 11))+ coord_flip()
 
-summm %>% ggplot(aes(x=mean_true,y=mean_bayes_cal))+geom_point()+geom_abline(slope=1, intercept = 0)
-summm %>% ggplot(aes(x=mean_true,y=mean_alrt_cal))+geom_point()+geom_abline(slope=1, intercept = 0)
+ultrafast<-summm_iqtree %>% ggplot(aes(y=mean_true,x=mean_ultrafast/100))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("Ultrafast\nbootstrap")+theme(text = element_text(size = 11))+ coord_flip()
 
-summm  %>% ggplot(aes(x=mean_true,y=mean_pred))+geom_point()+geom_abline(slope=1, intercept = 0)
+pred_iqtree<-summm_iqtree  %>% ggplot(aes(y=mean_true,x=mean_pred))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("IQTREE\nmodel")+theme(text = element_text(size = 11))+ coord_flip()
 
-pivoted<- summ %>% pivot_longer()
+iqtree_plot<-ggarrange(pred_iqtree,ultrafast,bayes,alrt, labels = c("A","B","C"),align = "h", nrow = 2, ncol = 2 )
+iqtree_plot
+```
+FASTTREE
+
+```{r}
+summm_fasttree<- fasttree_enriched_test %>% group_by(tree_id) %>% summarise(mean_true = mean(true_binary_support), mean_pred = mean(prob_predictions), mean_sh = mean(bootstrap_support))
+
+pred_fasttree<-summm_fasttree  %>% ggplot(aes(y=mean_true,x=mean_pred))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("Fasttree\nmodel")+theme(text = element_text(size = 11))+ coord_flip()
+
+sh<-summm_fasttree  %>% ggplot(aes(y=mean_true,x=mean_sh))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("SH test")+theme(text = element_text(size = 11))+ coord_flip()
+
+fasttree_plot<-ggarrange(pred_fasttree,sh, labels = c("A","B"),align = "h", nrow = 1, ncol = 2 )
+fasttree_plot
+```
+
+RAXML
+```{r}
+summm_raxml<- raxml_enriched_test %>% group_by(tree_id) %>% summarise(mean_true = mean(true_binary_support), mean_pred = mean(prob_predictions), mean_sh = mean(bootstrap_support))
+
+pred_raxml<-summm_raxml  %>% ggplot(aes(y=mean_true,x=mean_pred))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0, color = "green")+ylab("True fraction")+xlab("RAxML-NG\nmodel")+theme(text = element_text(size = 11))+ coord_flip()
+
+rapid<-summm_fasttree  %>% ggplot(aes(y=mean_true,x=mean_sh))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("Bootstrap\nsupport")+theme(text = element_text(size = 11))+ coord_flip()
+
+raxml_plot<-ggarrange(pred_raxml,rapid, labels = c("A","B"),align = "h", nrow = 1, ncol = 2 )
+raxml_plot
+```
+
+```{r}
+ggarrange(pred_iqtree,alrt, bayes, ultrafast,pred_raxml,rapid,"","",pred_fasttree,sh,align = "h", nrow = 3, ncol = 4 )
+
 ```
@@
 General performance
+IQTREE
 ```{r}
 
-out_path <-paste(prefix,"ML_results/",program,"/performance_text",".tsv", sep = "")
+out_path <-paste(prefix,"ML_results/","iqtree","/performance_text",".tsv", sep = "")
 
-paper_performance<-final_models_performance  %>% filter (metric_type=='all_data') %>% dplyr::select (dataset,analysis_type,name,mcc_0.5,logloss, AUC,average_precision, tp_0.5, tn_0.5, fp_0.5, fn_0.5)  %>% arrange(dataset) %>% mutate (description = case_when(analysis_type == "full_standard" & name =="model_standard" ~ "all features",
+paper_performance_iqtree<-iqtree_final_models_performance  %>% filter (metric_type=='all_data') %>% dplyr::select (dataset,analysis_type,name,mcc_0.5,logloss, AUC,average_precision, tp_0.5, tn_0.5, fp_0.5, fn_0.5)  %>% arrange(dataset) %>% mutate (description = case_when(analysis_type == "full_standard" & name =="model_standard" ~ "all features",
+                                analysis_type == "fast_standard" & name =="model_standard" ~ "All features except NNI features",
+                                analysis_type == "bootstrap_support" & name =="raw_only_boot" ~ "Raw ultrafast bootstrap support",
+                                analysis_type == "bootstrap_support" & name =="only_boot" ~ "Calibrated ultrafast bootstrap support",
+                                analysis_type == "bootstrap_support" & name =="inc_boot" ~ "All features+ ultrafast bootstrap support",
+                                    analysis_type == "feature_aLRT_iqtree_support" & name =="raw_only_boot" ~ "Raw aLRT support",
+                                analysis_type == "feature_aLRT_iqtree_support" & name =="only_boot" ~ "Calibrated aLRT support",
+                                analysis_type == "feature_aLRT_iqtree_support" & name =="inc_boot" ~ "All features+ aLRT support",
+                                                                    analysis_type == "feature_aBayes_iqtree_support" & name =="raw_only_boot" ~ "Raw aBayes support",
+                                analysis_type == "feature_aBayes_iqtree_support" & name =="only_boot" ~ "Calibrated aBayes support",
+                                analysis_type == "feature_aBayes_iqtree_support" & name =="inc_boot" ~ "All features+ aBayes support",
+                                )) %>% mutate(TPR =`tp_0.5`/(`tp_0.5`+`fn_0.5`),TNR =`tn_0.5`/(`tn_0.5`+`fp_0.5`),FPR =`fp_0.5`/(`tn_0.5`+`fp_0.5`),FNR =`fn_0.5`/(`tp_0.5`+`fn_0.5`)) %>% dplyr::select (dataset, description,logloss, mcc_0.5,FPR,FNR, AUC,average_precision)  %>% arrange(dataset)
+
+paper_performance_iqtree
+write_tsv(paper_performance_iqtree,out_path)
+
+paper_performance_iqtree %>% arrange(description)
+
+```
+FASTTREE
+```{r}
+
+out_path <-paste(prefix,"ML_results/","fasttree","/performance_text",".tsv", sep = "")
+
+paper_performance_fasttree<-fasttree_final_models_performance  %>% filter (metric_type=='all_data') %>% dplyr::select (dataset,analysis_type,name,mcc_0.5,logloss, AUC,average_precision, tp_0.5, tn_0.5, fp_0.5, fn_0.5)  %>% arrange(dataset) %>% mutate (description = case_when(analysis_type == "full_standard" & name =="model_standard" ~ "all features",
+                                analysis_type == "fast_standard" & name =="model_standard" ~ "All features except NNI features",
+                                analysis_type == "bootstrap_support" & name =="raw_only_boot" ~ "Raw SH support",
+                                analysis_type == "bootstrap_support" & name =="only_boot" ~ "Calibrated SH support",
+                                analysis_type == "bootstrap_support" & name =="inc_boot" ~ "All features+ SH support",
+                                    analysis_type == "feature_aLRT_iqtree_support" & name =="raw_only_boot" ~ "Raw aLRT support",
+                                analysis_type == "feature_aLRT_iqtree_support" & name =="only_boot" ~ "Calibrated aLRT support",
+                                analysis_type == "feature_aLRT_iqtree_support" & name =="inc_boot" ~ "All features+ aLRT support",
+                                                                    analysis_type == "feature_aBayes_iqtree_support" & name =="raw_only_boot" ~ "Raw aBayes support",
+                                analysis_type == "feature_aBayes_iqtree_support" & name =="only_boot" ~ "Calibrated aBayes support",
+                                analysis_type == "feature_aBayes_iqtree_support" & name =="inc_boot" ~ "All features+ aBayes support",
+                                )) %>% mutate(TPR =`tp_0.5`/(`tp_0.5`+`fn_0.5`),TNR =`tn_0.5`/(`tn_0.5`+`fp_0.5`),FPR =`fp_0.5`/(`tn_0.5`+`fp_0.5`),FNR =`fn_0.5`/(`tp_0.5`+`fn_0.5`)) %>% dplyr::select (dataset, description,logloss, mcc_0.5,FPR,FNR, AUC,average_precision)  %>% arrange(dataset)
+
+paper_performance_fasttree
+write_tsv(paper_performance_fasttree,out_path)
+paper_performance_fasttree %>% arrange(description)
+
+```
+RAXML
+```{r}
+
+out_path <-paste(prefix,"ML_results/","raxml","/performance_text",".tsv", sep = "")
+
+paper_performance_raxml<-raxml_final_models_performance  %>% filter (metric_type=='all_data') %>% dplyr::select (dataset,analysis_type,name,mcc_0.5,logloss, AUC,average_precision, tp_0.5, tn_0.5, fp_0.5, fn_0.5)  %>% arrange(dataset) %>% mutate (description = case_when(analysis_type == "full_standard" & name =="model_standard" ~ "all features",
                                 analysis_type == "fast_standard" & name =="model_standard" ~ "All features except NNI features",
@@
 
-paper_performance
-write_tsv(paper_performance,out_path)
+paper_performance_raxml
+write_tsv(paper_performance_raxml,out_path)
+paper_performance_raxml %>% arrange(description)
 
@@
 
+Results on test data
+```{r}
+total_model_performance = rbind(paper_performance_iqtree %>% mutate(program='IQTREE'), paper_performance_fasttree %>% mutate(program='FastTree'), paper_performance_raxml %>% mutate(program='RAxML-NG')) %>% filter (dataset %in% c("test"), description %in% c("all features","Raw SH support","Raw aLRT support","Raw aBayes support","Raw bootstrap support","Raw ultrafast bootstrap support")) %>% mutate(Support = description) %>% rename(MCC = mcc_0.5)
+
+mcc_plot<-total_model_performance %>% ggplot(aes(y=MCC, x = program, fill = Support))+geom_col(position = "dodge")
+logloss_plot<-total_model_performance %>% ggplot(aes(y=logloss, x = program, fill = Support))+geom_col(position = "dodge")
+#total_model_performance %>% ggplot(aes)
+
+ggarrange(mcc_plot,logloss_plot, labels = c("A","B"),align = "h", nrow = 2, ncol = 1, common.legend = TRUE )
+```
+
 Variable importance 
+IQTREE
+
 ```{r}
-out_path <-paste(prefix,"ML_results/",program,"/vi_text",".tsv", sep = "")
+out_path <-paste(prefix,"ML_results/","iqtree","/vi_text",".tsv", sep = "")
 
-vi_data<- vi %>% arrange (`...1`) %>% mutate (feature_name = case_when(
-  `...1` == "feature_min_ll_diff"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
-  `...1` =="feature_max_ll_diff"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",
+vi_data<- iqtree_vi %>% arrange (`...1`) %>% mutate (feature_name = case_when(
+  `...1` == "feature_min_ll_diff_opt=True_model_False"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
+  `...1` =="feature_max_ll_diff_opt=True_model_False"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",
+
+    `...1` == "feature_min_ll_diff_opt=False_model_False"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
+  `...1` =="feature_max_ll_diff_opt=False_model_False"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",
 
@@
 `...1` =='feature_partition_branch_vs_mean'~'Branch length at the parititon divided by total tree divergence',
+`...1` =='feature_var_bl'~'Variance of branch lengths across the tree',
+`...1` =='feature_25_pct_bl'~'25th percentile of tree branch lengths',
+`...1` =='feature_75_pct_bl'~'75th percentile of tree branch lengths',
+`...1` =='feature_mean_bl'~'Average of tree branch length',
 
@@
 `...1` =='feature_msa_pypythia_msa_difficulty'~'MSA difficulty',
+`...1` =='feature_max_mean_branch_length_opt=True_model_False'~'Difference in branch lengths',
 
-)) %>% dplyr::select (`...1`,feature_name,`Gini-importance`, mcc ) %>% arrange(feature_name,-`Gini-importance`)
+)) %>% dplyr::select (`...1`,feature_name,`Gini-importance`, mcc_0.5 ) %>% arrange(-`Gini-importance`,feature_name)
 vi_data
@@
 
+```
+
+FASTTREE
+
+```{r}
+out_path <-paste(prefix,"ML_results/","fasttree","/vi_text",".tsv", sep = "")
+
+vi_data<- fasttree_vi %>% arrange (`...1`) %>% mutate (feature_name = case_when(
+  `...1` == "feature_min_ll_diff_opt=True_model_False"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
+  `...1` =="feature_max_ll_diff_opt=True_model_False"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",
+
+    `...1` == "feature_min_ll_diff_opt=False_model_False"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
+  `...1` =="feature_max_ll_diff_opt=False_model_False"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",
+
+  `...1` =='feature_mean_parsimony_trees_binary'~"Fraction of parsimoy trees in which the bipartition exists",
+  `...1` =='feature_mean_parsimony_trees_neighbors_binary'~"Fraction of parsimony trees in which the neighbouring bipartitions exist",
+
+    `...1`=='feature_mean_all_ML_boot_raxml_neighbors_binary'~'Fraction of final ML trees in which the bipartition exists',
+
+   `...1` =='feature_mean_all_ML_boot_raxml_binary'~"Fraction of parsimony trees in which the neighbouring bipartitions exist",
+
+`...1` =='feature_mean_parsimony_trees'~"Mean transfer distance of the bipartition across parsimoy trees",
+`...1` =='feature_mean_parsimony_trees_neighbors'~"Mean transfer distance of the neighbouring bipartitions across parsimoy trees",
+
+`...1` =='feature_mean_all_ML_boot_raxml'~'Mean transfer distance of the bipartition across final ML trees',
+`...1` =='feature_mean_all_ML_boot_raxml_neighbors'~'Mean transfer distance of the neighbouring bipartitions across final ML trees',
+
+`...1` =='feature_partition_branch'~'Branch length at the parititon',
+`...1` =='feature_partition_branch_vs_mean'~'Branch length at the parititon divided by total tree divergence',
+`...1` =='feature_var_bl'~'Variance of branch lengths across the tree',
+`...1` =='feature_25_pct_bl'~'25th percentile of tree branch lengths',
+`...1` =='feature_75_pct_bl'~'75th percentile of tree branch lengths',
+`...1` =='feature_mean_bl'~'Average of tree branch length',
+
+`...1` =='feature_partition_divergence'~'Total divergence in the smaller subtree defined by the bipartition',
+`...1` =='feature_divergence_ratio'~'Total divergence in the smaller subtree defined by the bipartition divided by total tree divergence',
+`...1` =='feature_partition_size'~'Number of leaves in the smaller subtree defined by the bipartition',
+`...1` =='feature_partition_size_ratio'~'Fraction of leaves in the smaller subtree defined by the bipartition',
+
+`...1` =='feature_min_neighbor_brlen'~'Minimal branch length among the neighbouring branches',
+`...1` =='feature_mean_neighbor_brlen'~'Mean branch length among the neighbouring branches',
+`...1` =='feature_msa_constant_sites_pct'~'Fraction of constant sites among the MSA sites',
+`...1` =='feature_msa_n_loci'~'Number of positions in the MSA',
+`...1` =='feature_msa_n_unique_sites'~'Number of unique positions in the MSA',
+`...1` =='feature_msa_n_seq'~'Number of sequences in the MSA',
+`...1` =='feature_msa_pypythia_msa_difficulty'~'MSA difficulty',
+`...1` =='feature_max_mean_branch_length_opt=True_model_False'~'Difference in branch lengths',
+
+)) %>% dplyr::select (`...1`,feature_name,`Gini-importance`, mcc_0.5 ) %>% arrange(-`Gini-importance`)
+vi_data 
+write_tsv(vi_data,out_path)
+```
+
+RAXML
+
+```{r}
+out_path <-paste(prefix,"ML_results/","raxml","/vi_text",".tsv", sep = "")
+
+vi_data<- raxml_vi %>% arrange (`...1`) %>% mutate (feature_name = case_when(
+  `...1` == "feature_min_ll_diff_opt=True_model_False"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
+  `...1` =="feature_max_ll_diff_opt=True_model_False"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",
+
+    `...1` == "feature_min_ll_diff_opt=False_model_False"~ "Minimum log-likelihood difference between an NNI neighbor near the bipartition and current tree",
+  `...1` =="feature_max_ll_diff_opt=False_model_False"~ "Maximum log-likelihood difference between an NNI neighbor near the split and current tree",
+
+  `...1` =='feature_mean_parsimony_trees_binary'~"Fraction of parsimoy trees in which the bipartition exists",
+  `...1` =='feature_mean_parsimony_trees_neighbors_binary'~"Fraction of parsimony trees in which the neighbouring bipartitions exist",
+
+    `...1`=='feature_mean_all_ML_boot_raxml_neighbors_binary'~'Fraction of final ML trees in which the bipartition exists',
+
+   `...1` =='feature_mean_all_ML_boot_raxml_binary'~"Fraction of parsimony trees in which the neighbouring bipartitions exist",
+
+`...1` =='feature_mean_parsimony_trees'~"Mean transfer distance of the bipartition across parsimoy trees",
+`...1` =='feature_mean_parsimony_trees_neighbors'~"Mean transfer distance of the neighbouring bipartitions across parsimoy trees",
+
+`...1` =='feature_mean_all_ML_boot_raxml'~'Mean transfer distance of the bipartition across final ML trees',
+`...1` =='feature_mean_all_ML_boot_raxml_neighbors'~'Mean transfer distance of the neighbouring bipartitions across final ML trees',
+
+`...1` =='feature_partition_branch'~'Branch length at the parititon',
+`...1` =='feature_partition_branch_vs_mean'~'Branch length at the parititon divided by total tree divergence',
+`...1` =='feature_var_bl'~'Variance of branch lengths across the tree',
+`...1` =='feature_25_pct_bl'~'25th percentile of tree branch lengths',
+`...1` =='feature_75_pct_bl'~'75th percentile of tree branch lengths',
+`...1` =='feature_mean_bl'~'Average of tree branch length',
+
+`...1` =='feature_partition_divergence'~'Total divergence in the smaller subtree defined by the bipartition',
+`...1` =='feature_divergence_ratio'~'Total divergence in the smaller subtree defined by the bipartition divided by total tree divergence',
+`...1` =='feature_partition_size'~'Number of leaves in the smaller subtree defined by the bipartition',
+`...1` =='feature_partition_size_ratio'~'Fraction of leaves in the smaller subtree defined by the bipartition',
+
+`...1` =='feature_min_neighbor_brlen'~'Minimal branch length among the neighbouring branches',
+`...1` =='feature_mean_neighbor_brlen'~'Mean branch length among the neighbouring branches',
+`...1` =='feature_msa_constant_sites_pct'~'Fraction of constant sites among the MSA sites',
+`...1` =='feature_msa_n_loci'~'Number of positions in the MSA',
+`...1` =='feature_msa_n_unique_sites'~'Number of unique positions in the MSA',
+`...1` =='feature_msa_n_seq'~'Number of sequences in the MSA',
+`...1` =='feature_msa_pypythia_msa_difficulty'~'MSA difficulty',
+`...1` =='feature_max_mean_branch_length_opt=True_model_False'~'Difference in branch lengths',
+
+)) %>% dplyr::select (`...1`,feature_name,`Gini-importance`, mcc_0.5 ) %>% arrange(-`Gini-importance`)
+vi_data %>% arrange(feature_name)
+write_tsv(vi_data,out_path)
 ```
@@
 }
-conf_data<-enriched_test %>% dplyr::select (prob_predictions, true_label) %>% mutate (pred_status = prob_predictions>0.5)
+conf_data<-iqtree_enriched_test %>% dplyr::select (prob_predictions, true_label) %>% mutate (pred_status = prob_predictions>0.5)
 
@@
 confusion_plot<-ggplotConfusionMatrix(confusionMatrix(data = pred, reference= obs))+theme(legend.position = "none")
-
-```
-Prediction metrics
-
-```{r}
-library(predtools)
-library(MASS)
-library(caret)
 
@@
 
-```
-```{r}
-test_group_performance %>% filter (grouping_col_name =='feature_msa_n_loci') 
 ```
@@
 
-```
-
-```{r}
 cor.test( enriched_test$feature_msa_pypythia_msa_difficulty,  enriched_test$prob_predictions)
@@
 cor.test( enriched_test$feature_msa_n_loci, enriched_test$prob_predictions)
+
 ```
+
+Running time
 
@@
 ```
+
 ```{r}

### 2023-11-25 ML_pipeline/ML_boot_edit.py

@@ -13,6 +13,7 @@ from side_code.MSA_manipulation import get_MSA_seq_names
 from sklearn.metrics import balanced_accuracy_score
 import os
 import matplotlib.pyplot as plt
+from random import sample
 import argparse
 
 
@@ -88,23 +89,24 @@ def bootstrap_model_pipeline(working_dir, train, test, validation_dict, features
                               do_RFE=do_RFE,
                               n_cv_folds=3)
 
-    raw_boot_performance, groups_data_raw_boot,test_predictions_raw_boot = overall_model_performance_analysis(working_dir, None, only_boot_dict,
-                                                                                    name="raw_only_boot",
-                                                                                    extract_predictions=False)
+    raw_boot_performance, groups_data_raw_boot,predictions_raw_boot = overall_model_performance_analysis(working_dir,
+                                                                                                         None,
+                                                                                                         only_boot_dict,
+                                                                                                         name="raw_only_boot")
 
-    only_boot_performance, groups_data_only_boot,test_predictions_only_boot = overall_model_performance_analysis(working_dir, model_only_boot,
-                                                                                      only_boot_dict,
-                                                                                      name="only_boot",
-                                                                                      extract_predictions=False)
-    inc_boot_performance, groups_data_inc_boot,test_predictions_inc_boot = overall_model_performance_analysis(working_dir, model_inc_boot,
-                                                                                    inc_boot_dict,
-                                                                                    name="inc_boot",
-                                                                                    extract_predictions=False)
+    only_boot_performance, groups_data_only_boot,predictions_only_boot = overall_model_performance_analysis(working_dir,
+                                                                                                            model_only_boot,
+                                                                                                            only_boot_dict,
+                                                                                                            name="only_boot")
+    inc_boot_performance, groups_data_inc_boot,predictions_inc_boot = overall_model_performance_analysis(working_dir,
[…]

### 2023-11-25 ML_utils/ML_algorithms_and_hueristics.py

@@ -4,8 +4,7 @@ import pandas as pd
 from sklearn.ensemble import RandomForestClassifier
 import pickle
 from sklearn.metrics import mean_squared_error, r2_score, mean_absolute_error
-from sklearn.metrics import matthews_corrcoef, balanced_accuracy_score, log_loss, roc_auc_score, average_precision_score, accuracy_score, precision_score, \
-    recall_score
+from sklearn.metrics import matthews_corrcoef, log_loss, brier_score_loss, roc_auc_score, average_precision_score
 import numpy as np
 import os
 from sklearn.metrics import balanced_accuracy_score
@@ -148,7 +147,7 @@ def model_evaluation_metrics(y_true, prob_predictions, estimate_auc = True):
         t_metrics = {f'tn_{threshold}': tn, f'fp_{threshold}': fp,f'fn_{threshold}': fn,f'tp_{threshold}': tp, f'mcc_{threshold}': mcc}
         metrics.update(t_metrics)
     if estimate_auc:
-        prob_metrics = {'AUC': roc_auc_score(y_true, prob_predictions),
+        prob_metrics = {'AUC': roc_auc_score(y_true, prob_predictions),'brier_loss': brier_score_loss(y_true, prob_predictions),
                         'logloss': log_loss(y_true, prob_predictions), 'average_precision': average_precision_score(y_true, prob_predictions), }
         metrics.update(prob_metrics)
     return metrics
@@ -166,7 +165,7 @@ def model_groups_anlaysis(X,y_true, predictions):
     return performance_per_group
 
 
-def overall_model_performance_analysis(working_dir,model, data_dict, name,extract_predictions = False):
+def overall_model_performance_analysis(working_dir, model, data_dict, name):
     if model:
         var_impt = variable_importance((data_dict["train"]["X"]).columns[model['selector'].get_support(indices=True)], model['best_model'])
         enrich_with_single_feature_metrics(var_impt, data_dict["train"]["X"], data_dict["train"]["y"], data_dict["test"]["X"], data_dict["test"]["y"])
@@ -178,11 +177,9 @@ def overall_model_performance_analysis(working_dir,model, data_dict, name,extrac
     for dataset in data_dict:
         if model:
             prob_predictions =  model['calibrated_model'].predict_proba((model['selector']).transform(data_dict[dataset]["X"]))[:, 1]
-            predictions = model['calibrated_model'].predict((model['selector']).transform(data_dict[dataset]["X"]))
[…]

### 2023-11-26 ML_pipeline/ML_boot_edit.py
new changes

@@ -286,12 +286,14 @@ def main():
 
 
     if args.use_val_data:
-        fasttree_validation = pd.read_csv(os.path.join(args.validation_data_folder, f'simulations_df_fasttree.tsv'))
-        iqtree_validation = pd.read_csv(os.path.join(args.validation_data_folder, f'simulations_df_iqtree.tsv'))
-        raxml_validation = pd.read_csv(os.path.join(args.validation_data_folder, f'simulations_df_raxml.tsv'))
+        fasttree_validation = pd.read_csv(os.path.join(args.validation_data_folder, f'simulations_df_fasttree.tsv'), sep='\t')
+        iqtree_validation = pd.read_csv(os.path.join(args.validation_data_folder, f'simulations_df_iqtree.tsv'),sep='\t')
+        raxml_validation = pd.read_csv(os.path.join(args.validation_data_folder, f'simulations_df_raxml.tsv'),sep='\t')
         common_tree_ids = list(set(fasttree_validation['tree_id']) & set(raxml_validation['tree_id']) & set(
            iqtree_validation['tree_id']))
         selected_tree_ids = sample(common_tree_ids,args.n_val_samp)
+    else:
+        selected_tree_ids = []
 
     for program in args.programs.split('_'):
         logging.info(f"Program = {program}")
@@ -304,6 +306,7 @@ def main():
             logging.info(f"Using existing training data in {training_data_path} ")
             program_data = pd.read_csv(training_data_path, sep='\t')
         program_data = transform_data(program_data, program)
+        program_data = program_data.loc[~program_data.tree_id.isin(selected_tree_ids)]
         validation_dict = {}
         validation_data_path = os.path.join(args.validation_data_folder, f'simulations_df_{program}.tsv')
         if args.use_val_data:

### 2023-11-26 ML_pipeline/ML_boot_edit.py
new change

@@ -318,7 +318,7 @@ def main():
             else:
                 logging.info("Using existing validation data")
                 program_validation_data = pd.read_csv(validation_data_path, sep='\t')
-            program_validation_data = transform_data(program_validation_data)
+            program_validation_data = transform_data(program_validation_data,program)
             logging.info("Subsampling tree inds in validation to the common tree inds")
             program_validation_data = program_validation_data.loc[program_validation_data.tree_id.isin(selected_tree_ids)]
             for model_mode in np.unique(program_validation_data["model_mode"]):

### 2023-11-26 ML_pipeline/ML_boot_edit.py
log

@@ -318,8 +318,10 @@ def main():
             else:
                 logging.info("Using existing validation data")
                 program_validation_data = pd.read_csv(validation_data_path, sep='\t')
+
             program_validation_data = transform_data(program_validation_data,program)
             logging.info("Subsampling tree inds in validation to the common tree inds")
+            logging.info(f"Number of MSAs in validation is {len(program_data['tree_id'].unqiue())}")
             program_validation_data = program_validation_data.loc[program_validation_data.tree_id.isin(selected_tree_ids)]
             for model_mode in np.unique(program_validation_data["model_mode"]):
                 validation_dict[f'val_{model_mode}'] = program_validation_data.loc[

### 2023-11-26 ML_pipeline/ML_boot_edit.py
fix

@@ -321,7 +321,7 @@ def main():
 
             program_validation_data = transform_data(program_validation_data,program)
             logging.info("Subsampling tree inds in validation to the common tree inds")
-            logging.info(f"Number of MSAs in validation is {len(program_data['tree_id'].unqiue())}")
+            logging.info(f"Number of MSAs in validation is {len(np.unique(program_data['tree_id']))}")
             program_validation_data = program_validation_data.loc[program_validation_data.tree_id.isin(selected_tree_ids)]
             for model_mode in np.unique(program_validation_data["model_mode"]):
                 validation_dict[f'val_{model_mode}'] = program_validation_data.loc[

### 2023-11-26 ML_pipeline/ML_boot_edit.py
new fix

@@ -196,8 +196,7 @@ def ML_pipeline(program_data, bootstrap_cols, cpus_per_main_job, working_dir, sa
     group_performance_full["sample_frac"] = sample_frac
     if extract_predictions:
         generate_enriched_datasets(working_dir, all_perdictions, train, test, validation_dict)
-        all_models_performance.to_csv(os.path.join(working_dir, 'final_model_performance.tsv'), sep=CSV_SEP)
-        group_performance_full.to_csv(os.path.join(working_dir, 'groups_performance.tsv'), sep=CSV_SEP)
+    return all_models_performance, group_performance_full
 
 
 
@@ -321,7 +320,7 @@ def main():
 
             program_validation_data = transform_data(program_validation_data,program)
             logging.info("Subsampling tree inds in validation to the common tree inds")
-            logging.info(f"Number of MSAs in validation is {len(np.unique(program_data['tree_id']))}")
+            logging.info(f"Number of MSAs in validation is {len(np.unique(program_validation_data['tree_id']))}")
             program_validation_data = program_validation_data.loc[program_validation_data.tree_id.isin(selected_tree_ids)]
             for model_mode in np.unique(program_validation_data["model_mode"]):
                 validation_dict[f'val_{model_mode}'] = program_validation_data.loc[
@@ -342,7 +341,7 @@ def main():
 
 
 
-                curr_model_metrics, groups_analysis, curr_test_with_predictions = ML_pipeline(program_data, bootstrap_cols, args.cpus_per_main_job,
+                curr_model_metrics, groups_analysis, curr_predictions = ML_pipeline(program_data, bootstrap_cols, args.cpus_per_main_job,
                                                                   sample_frac_working_dir, sample_frac,
                                                                   compare_to_bootstrap_models=False,
                                                                   subsample_train=True, do_RFE=args.RFE,
@@ -353,13 +352,14 @@ def main():
         logging.info(f"Generating optimized final model")
         final_model_working_dir = os.path.join(working_dir, f"final_model")
         create_dir_if_not_exists(final_model_working_dir)
-        ML_pipeline(program_data, bootstrap_cols, args.cpus_per_main_job,
+        final_models_performance,final_group_performance_full= ML_pipeline(program_data, bootstrap_cols, args.cpus_per_main_job,
                                                            final_model_working_dir, sample_frac=-1,
                                                            subsample_train=False, do_RFE=args.RFE,
                                                            large_grid=args.full_grid, name=f"final_model",
[…]

### 2023-11-26 ML_pipeline/ML_boot_edit.py
fix

@@ -341,7 +341,7 @@ def main():
 
 
 
-                curr_model_metrics, groups_analysis, curr_predictions = ML_pipeline(program_data, bootstrap_cols, args.cpus_per_main_job,
+                curr_model_metrics, groups_analysis = ML_pipeline(program_data, bootstrap_cols, args.cpus_per_main_job,
                                                                   sample_frac_working_dir, sample_frac,
                                                                   compare_to_bootstrap_models=False,
                                                                   subsample_train=True, do_RFE=args.RFE,

### 2023-11-27 R_code/paper_analysis.Rmd
fix

@@
 ```
-```{r}
-validation_data_prefix = "/Users/noa/Workspace/bootstrap_results/validation_data"
-fasttree_validation_data = read_tsv(paste(validation_data_prefix,"/simulations_df_fasttree.tsv", sep = ""))
-raxml_validation_data = read_tsv(paste(validation_data_prefix,"/simulations_df_raxml.tsv", sep = ""))
-iqtree_validation_data = read_tsv(paste(validation_data_prefix,"/simulations_df_iqtree.tsv", sep = ""))
-```
 
@@
 fasttree_vi =  read_tsv(paste(prefix,"ML_results/fasttree/final_model/full_model/model_standard_vi.tsv", sep = ""))
-fasttree_enriched_test = read_tsv(paste(prefix,"ML_results/fasttree/test_with_predictions.tsv", sep = ""))
+fasttree_enriched_test = read_tsv(paste(prefix,"ML_results/fasttree/final_model/test.tsv", sep = ""))
 fasttree_validation = read_tsv(paste(prefix,"validation_data/simulations_df_fasttree.tsv", sep = ""))
+fasttree_enriched_val_downgrade = read_tsv(paste(prefix,"ML_results/fasttree/final_model/val_downgrade.tsv", sep = ""))
+fasttree_enriched_val_standard = read_tsv(paste(prefix,"ML_results/fasttree/final_model/val_upgrade.tsv", sep = ""))
+fasttree_enriched_val_upgrade = read_tsv(paste(prefix,"ML_results/fasttree/final_model/val_standard.tsv", sep = ""))
 #####
@@
 iqtree_vi =  read_tsv(paste(prefix,"ML_results/iqtree/final_model/full_model/model_standard_vi.tsv", sep = ""))
-iqtree_enriched_test = read_tsv(paste(prefix,"ML_results/iqtree/test_with_predictions.tsv", sep = ""))
-iqtree_validation = read_tsv(paste(prefix,"validation_data/simulations_df_iqtree.tsv", sep = ""))
+iqtree_enriched_test = read_tsv(paste(prefix,"ML_results/iqtree/final_model/test.tsv", sep = ""))
+iqtree_enriched_val_downgrade = read_tsv(paste(prefix,"ML_results/iqtree/final_model/val_downgrade.tsv", sep = ""))
+iqtree_enriched_val_standard = read_tsv(paste(prefix,"ML_results/iqtree/final_model/val_upgrade.tsv", sep = ""))
+iqtree_enriched_val_upgrade = read_tsv(paste(prefix,"ML_results/iqtree/final_model/val_standard.tsv", sep = ""))
 #####
@@
 raxml_vi =  read_tsv(paste(prefix,"ML_results/raxml/final_model/full_model/model_standard_vi.tsv", sep = ""))
-raxml_enriched_test = read_tsv(paste(prefix,"ML_results/raxml//test_with_predictions.tsv", sep = ""))
-raxml_validation = read_tsv(paste(prefix,"validation_data/simulations_df_raxml.tsv", sep = ""))
-
-```
-
-Choose specific program
-```{r}
-program = 'fasttree'
-if (program=='fasttree') {
-    all_models_performance = fasttree_all_models_performance
-    final_models_performance = fasttree_final_models_performance %>% filter(analysis_type!='feature_standard_fasttree_boot_support')
-    group_performance = fasttree_group_performance
-    vi = fasttree_vi
-    enriched_test = fasttree_enriched_test
-    #validation_data = fasttree_validation_data
-
-} else if (program=='iqtree') {
-    all_models_performance = iqtree_all_models_performance
-    final_models_performance = iqtree_final_models_performance
-    group_performance = iqtree_group_performance
-    vi = iqtree_vi
-    enriched_test = iqtree_enriched_test
-    #validation_data = iqtree_validation_data
-} else if (program=='raxml') {
-    all_models_performance = raxml_all_models_performance
-    final_models_performance = raxml_final_models_performance
-    group_performance = raxml_group_performance
-    vi = raxml_vi
-    enriched_test = raxml_enriched_test
-    #validation_data = raxml_validation
-}
-```
-
-IQTREE
-```{r}
-
-summm_iqtree<- iqtree_enriched_test %>% group_by(tree_id) %>% summarise(mean_true = mean(true_binary_support), mean_pred = mean(prob_predictions), mean_bayes = mean(feature_aBayes_iqtree_support), mean_ultrafast = mean(bootstrap_support), mean_bayes_cal = mean(predictions_feature_aBayes_iqtree_support_only_boot),
-                                                mean_alrt = mean(feature_aLRT_iqtree_support), mean_alrt_cal = mean(predictions_feature_aLRT_iqtree_support_only_boot))
-
-bayes<-summm_iqtree%>% ggplot(aes(y=mean_true,x=mean_bayes))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("aBayes test")+theme(text = element_text(size = 11))+ coord_flip()
-
-alrt<-summm_iqtree %>% ggplot(aes(y=mean_true,x=mean_alrt))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("aLRT test ")+theme(text = element_text(size = 11))+ coord_flip()
-
-ultrafast<-summm_iqtree %>% ggplot(aes(y=mean_true,x=mean_ultrafast/100))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("Ultrafast\nbootstrap")+theme(text = element_text(size = 11))+ coord_flip()
-
-pred_iqtree<-summm_iqtree  %>% ggplot(aes(y=mean_true,x=mean_pred))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("IQTREE\nmodel")+theme(text = element_text(size = 11))+ coord_flip()
-
-iqtree_plot<-ggarrange(pred_iqtree,ultrafast,bayes,alrt, labels = c("A","B","C"),align = "h", nrow = 2, ncol = 2 )
-iqtree_plot
-```
-FASTTREE
-
-```{r}
-summm_fasttree<- fasttree_enriched_test %>% group_by(tree_id) %>% summarise(mean_true = mean(true_binary_support), mean_pred = mean(prob_predictions), mean_sh = mean(bootstrap_support))
-
-pred_fasttree<-summm_fasttree  %>% ggplot(aes(y=mean_true,x=mean_pred))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("Fasttree\nmodel")+theme(text = element_text(size = 11))+ coord_flip()
-
-sh<-summm_fasttree  %>% ggplot(aes(y=mean_true,x=mean_sh))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("SH test")+theme(text = element_text(size = 11))+ coord_flip()
-
-fasttree_plot<-ggarrange(pred_fasttree,sh, labels = c("A","B"),align = "h", nrow = 1, ncol = 2 )
-fasttree_plot
-```
-
-RAXML
-```{r}
-summm_raxml<- raxml_enriched_test %>% group_by(tree_id) %>% summarise(mean_true = mean(true_binary_support), mean_pred = mean(prob_predictions), mean_sh = mean(bootstrap_support))
-
-pred_raxml<-summm_raxml  %>% ggplot(aes(y=mean_true,x=mean_pred))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0, color = "green")+ylab("True fraction")+xlab("RAxML-NG\nmodel")+theme(text = element_text(size = 11))+ coord_flip()
-
-rapid<-summm_fasttree  %>% ggplot(aes(y=mean_true,x=mean_sh))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("Bootstrap\nsupport")+theme(text = element_text(size = 11))+ coord_flip()
-
-raxml_plot<-ggarrange(pred_raxml,rapid, labels = c("A","B"),align = "h", nrow = 1, ncol = 2 )
-raxml_plot
-```
-
-```{r}
-ggarrange(pred_iqtree,alrt, bayes, ultrafast,pred_raxml,rapid,"","",pred_fasttree,sh,align = "h", nrow = 3, ncol = 4 )
+raxml_enriched_test = read_tsv(paste(prefix,"ML_results/raxml/final_model/test.tsv", sep = ""))
+raxml_enriched_val_downgrade = read_tsv(paste(prefix,"ML_results/fasttree/final_model/val_downgrade.tsv", sep = ""))
+raxml_enriched_val_standard = read_tsv(paste(prefix,"ML_results/fasttree/final_model/val_upgrade.tsv", sep = ""))
+raxml_enriched_val_upgrade = read_tsv(paste(prefix,"ML_results/fasttree/final_model/val_standard.tsv", sep = ""))
 
@@
 
-paper_performance_iqtree<-iqtree_final_models_performance  %>% filter (metric_type=='all_data') %>% dplyr::select (dataset,analysis_type,name,mcc_0.5,logloss, AUC,average_precision, tp_0.5, tn_0.5, fp_0.5, fn_0.5)  %>% arrange(dataset) %>% mutate (description = case_when(analysis_type == "full_standard" & name =="model_standard" ~ "all features",
+paper_performance_iqtree<-iqtree_final_models_performance  %>% filter (metric_type=='all_data') %>% dplyr::select (dataset,analysis_type,name,brier_loss,logloss, AUC,average_precision,mcc_0.5, tp_0.5, tn_0.5, fp_0.5, fn_0.5)  %>% arrange(dataset) %>% mutate (description = case_when(analysis_type == "full_standard" & name =="model_standard" ~ "all features",
                                 analysis_type == "fast_standard" & name =="model_standard" ~ "All features except NNI features",
@@
                                 analysis_type == "feature_aBayes_iqtree_support" & name =="inc_boot" ~ "All features+ aBayes support",
-                                )) %>% mutate(TPR =`tp_0.5`/(`tp_0.5`+`fn_0.5`),TNR =`tn_0.5`/(`tn_0.5`+`fp_0.5`),FPR =`fp_0.5`/(`tn_0.5`+`fp_0.5`),FNR =`fn_0.5`/(`tp_0.5`+`fn_0.5`)) %>% dplyr::select (dataset, description,logloss, mcc_0.5,FPR,FNR, AUC,average_precision)  %>% arrange(dataset)
+                                )) %>% mutate(TPR =`tp_0.5`/(`tp_0.5`+`fn_0.5`),TNR =`tn_0.5`/(`tn_0.5`+`fp_0.5`),FPR =`fp_0.5`/(`tn_0.5`+`fp_0.5`),FNR =`fn_0.5`/(`tp_0.5`+`fn_0.5`)) %>% dplyr::select (dataset, description, brier_loss,logloss,AUC,average_precision, mcc_0.5,FPR,FNR, )  %>% arrange(dataset)
 

### 2023-11-27 ML_pipeline/ML_boot_edit.py

@@ -240,7 +240,7 @@ def transform_data(df, program):
     #df['model_short'] = df['model_short'].apply(lambda x: x.split('+')[0])
     #df['feature_free_parameters'] = df['model_short'].apply(lambda x: get_n_free_parameters(x))
     df = df[[col for col in df.columns if
-             'msa_entropy' not in col and 'extraction_of_features_time' not in col and 'feature_abayes_opt' not in col and 'column_variance' not in col and 'feature_min_mean_branch_length' not in col and 'feature_max_mean_branch_length' and 'll_diff_norm' not in col ]]
+             'msa_entropy' not in col and 'extraction_of_features_time' not in col and 'feature_abayes_opt' not in col and 'column_variance' not in col and 'feature_min_mean_branch_length' not in col and 'feature_max_mean_branch_length' not in col and 'll_diff_norm' not in col ]]
     # df = pd.get_dummies(df,prefix='feature_model_',columns=['model_short']) #
     if program=='raxml':
         df['bootstrap_support'] =df['bootstrap_support']/100
@@ -249,7 +249,7 @@ def transform_data(df, program):
         df['feature_aLRT_iqtree_support']= np.minimum(df['feature_aLRT_iqtree_support'],1)
         df['feature_aBayes_iqtree_support'] = np.minimum(df['feature_aBayes_iqtree_support'],1)
     elif program=='fasttree':
-        df['bootstrap_support'] = np.maximum(df['bootstrap_support'], 1)
+        df['bootstrap_support'] = np.minimum(df['bootstrap_support'], 1)
     return df
     # +[col for col in df.columns if 'msa_entropy' in col]
 

### 2023-11-30 R_code/paper_analysis.Rmd
new changes for ML pipeline

@@
 library(tidyverse)
-library(caret)
 library(ggpubr)
@@
 library(MASS)
-library(caret)
 ```
@@
 fasttree_enriched_test = read_tsv(paste(prefix,"ML_results/fasttree/final_model/test.tsv", sep = ""))
+fasttree_enriched_train = read_tsv(paste(prefix,"ML_results/fasttree/final_model/train.tsv", sep = ""))
 fasttree_validation = read_tsv(paste(prefix,"validation_data/simulations_df_fasttree.tsv", sep = ""))
@@
 iqtree_enriched_test = read_tsv(paste(prefix,"ML_results/iqtree/final_model/test.tsv", sep = ""))
+iqtree_enriched_train = read_tsv(paste(prefix,"ML_results/iqtree/final_model/train.tsv", sep = ""))
 iqtree_enriched_val_downgrade = read_tsv(paste(prefix,"ML_results/iqtree/final_model/val_downgrade.tsv", sep = ""))
@@
 
-General performance
-IQTREE
 ```{r}
+iqtree_enriched_train %>% ggplot(aes(x=feature_msa_pypythia_msa_difficulty))+geom_histogram()
+iqtree_enriched_test %>% ggplot(aes(x=feature_msa_pypythia_msa_difficulty))+geom_histogram()
+iqtree_enriched_val_standard %>% ggplot(aes(x=feature_msa_pypythia_msa_difficulty))+geom_histogram()
+iqtree_enriched_val_upgrade %>% ggplot(aes(x=feature_msa_pypythia_msa_difficulty))+geom_histogram()
+iqtree_enriched_val_downgrade %>% ggplot(aes(x=feature_msa_pypythia_msa_difficulty))+geom_histogram()
 
-out_path <-paste(prefix,"ML_results/","iqtree","/performance_text",".tsv", sep = "")
-
-paper_performance_iqtree<-iqtree_final_models_performance  %>% filter (metric_type=='all_data') %>% dplyr::select (dataset,analysis_type,name,brier_loss,logloss, AUC,average_precision,mcc_0.5, tp_0.5, tn_0.5, fp_0.5, fn_0.5)  %>% arrange(dataset) %>% mutate (description = case_when(analysis_type == "full_standard" & name =="model_standard" ~ "all features",
-                                analysis_type == "fast_standard" & name =="model_standard" ~ "All features except NNI features",
-                                analysis_type == "bootstrap_support" & name =="raw_only_boot" ~ "Raw ultrafast bootstrap support",
-                                analysis_type == "bootstrap_support" & name =="only_boot" ~ "Calibrated ultrafast bootstrap support",
-                                analysis_type == "bootstrap_support" & name =="inc_boot" ~ "All features+ ultrafast bootstrap support",
-                                    analysis_type == "feature_aLRT_iqtree_support" & name =="raw_only_boot" ~ "Raw aLRT support",
-                                analysis_type == "feature_aLRT_iqtree_support" & name =="only_boot" ~ "Calibrated aLRT support",
-                                analysis_type == "feature_aLRT_iqtree_support" & name =="inc_boot" ~ "All features+ aLRT support",
-                                                                    analysis_type == "feature_aBayes_iqtree_support" & name =="raw_only_boot" ~ "Raw aBayes support",
-                                analysis_type == "feature_aBayes_iqtree_support" & name =="only_boot" ~ "Calibrated aBayes support",
-                                analysis_type == "feature_aBayes_iqtree_support" & name =="inc_boot" ~ "All features+ aBayes support",
-                                )) %>% mutate(TPR =`tp_0.5`/(`tp_0.5`+`fn_0.5`),TNR =`tn_0.5`/(`tn_0.5`+`fp_0.5`),FPR =`fp_0.5`/(`tn_0.5`+`fp_0.5`),FNR =`fn_0.5`/(`tp_0.5`+`fn_0.5`)) %>% dplyr::select (dataset, description, brier_loss,logloss,AUC,average_precision, mcc_0.5,FPR,FNR, )  %>% arrange(dataset)
-
-paper_performance_iqtree
-write_tsv(paper_performance_iqtree,out_path)
-
-paper_performance_iqtree %>% arrange(description)
-
-```
-FASTTREE
-```{r}
-
-out_path <-paste(prefix,"ML_results/","fasttree","/performance_text",".tsv", sep = "")
-
-paper_performance_fasttree<-fasttree_final_models_performance  %>% filter (metric_type=='all_data') %>% dplyr::select (dataset,analysis_type,name,mcc_0.5,logloss, AUC,average_precision, tp_0.5, tn_0.5, fp_0.5, fn_0.5)  %>% arrange(dataset) %>% mutate (description = case_when(analysis_type == "full_standard" & name =="model_standard" ~ "all features",
-                                analysis_type == "fast_standard" & name =="model_standard" ~ "All features except NNI features",
-                                analysis_type == "bootstrap_support" & name =="raw_only_boot" ~ "Raw SH support",
-                                analysis_type == "bootstrap_support" & name =="only_boot" ~ "Calibrated SH support",
-                                analysis_type == "bootstrap_support" & name =="inc_boot" ~ "All features+ SH support",
-                                    analysis_type == "feature_aLRT_iqtree_support" & name =="raw_only_boot" ~ "Raw aLRT support",
-                                analysis_type == "feature_aLRT_iqtree_support" & name =="only_boot" ~ "Calibrated aLRT support",
-                                analysis_type == "feature_aLRT_iqtree_support" & name =="inc_boot" ~ "All features+ aLRT support",
-                                                                    analysis_type == "feature_aBayes_iqtree_support" & name =="raw_only_boot" ~ "Raw aBayes support",
-                                analysis_type == "feature_aBayes_iqtree_support" & name =="only_boot" ~ "Calibrated aBayes support",
-                                analysis_type == "feature_aBayes_iqtree_support" & name =="inc_boot" ~ "All features+ aBayes support",
-                                )) %>% mutate(TPR =`tp_0.5`/(`tp_0.5`+`fn_0.5`),TNR =`tn_0.5`/(`tn_0.5`+`fp_0.5`),FPR =`fp_0.5`/(`tn_0.5`+`fp_0.5`),FNR =`fn_0.5`/(`tp_0.5`+`fn_0.5`)) %>% dplyr::select (dataset, description,logloss, mcc_0.5,FPR,FNR, AUC,average_precision)  %>% arrange(dataset)
-
-paper_performance_fasttree
-write_tsv(paper_performance_fasttree,out_path)
-paper_performance_fasttree %>% arrange(description)
-
-```
-RAXML
-```{r}
-
-out_path <-paste(prefix,"ML_results/","raxml","/performance_text",".tsv", sep = "")
-
-paper_performance_raxml<-raxml_final_models_performance  %>% filter (metric_type=='all_data') %>% dplyr::select (dataset,analysis_type,name,mcc_0.5,logloss, AUC,average_precision, tp_0.5, tn_0.5, fp_0.5, fn_0.5)  %>% arrange(dataset) %>% mutate (description = case_when(analysis_type == "full_standard" & name =="model_standard" ~ "all features",
-                                analysis_type == "fast_standard" & name =="model_standard" ~ "All features except NNI features",
-                                analysis_type == "bootstrap_support" & name =="raw_only_boot" ~ "Raw bootstrap support",
-                                analysis_type == "bootstrap_support" & name =="only_boot" ~ "Calibrated bootstrap support",
-                                analysis_type == "bootstrap_support" & name =="inc_boot" ~ "All features+ bootstrap support",
-                                    analysis_type == "feature_aLRT_iqtree_support" & name =="raw_only_boot" ~ "Raw aLRT support",
-                                analysis_type == "feature_aLRT_iqtree_support" & name =="only_boot" ~ "Calibrated aLRT support",
-                                analysis_type == "feature_aLRT_iqtree_support" & name =="inc_boot" ~ "All features+ aLRT support",
-                                                                    analysis_type == "feature_aBayes_iqtree_support" & name =="raw_only_boot" ~ "Raw aBayes support",
-                                analysis_type == "feature_aBayes_iqtree_support" & name =="only_boot" ~ "Calibrated aBayes support",
-                                analysis_type == "feature_aBayes_iqtree_support" & name =="inc_boot" ~ "All features+ aBayes support",
-                                )) %>% mutate(TPR =`tp_0.5`/(`tp_0.5`+`fn_0.5`),TNR =`tn_0.5`/(`tn_0.5`+`fp_0.5`),FPR =`fp_0.5`/(`tn_0.5`+`fp_0.5`),FNR =`fn_0.5`/(`tp_0.5`+`fn_0.5`)) %>% dplyr::select (dataset, description,logloss, mcc_0.5,FPR,FNR, AUC,average_precision)  %>% arrange(dataset)
-
-paper_performance_raxml
-write_tsv(paper_performance_raxml,out_path)
-paper_performance_raxml %>% arrange(description)
-
+iqtree_enriched_test %>% ggplot(aes(x=feature_msa_pypythia_msa_difficulty))+geom_histogram()
+iqtree_enriched_val_standard %>% ggplot(aes(x=feature_msa_pypythia_msa_difficulty))+geom_histogram()
+iqtree_enriched_val_upgrade %>% ggplot(aes(x=feature_msa_pypythia_msa_difficulty))+geom_histogram()
+iqtree_enriched_val_downgrade %>% ggplot(aes(x=feature_msa_pypythia_msa_difficulty))+geom_histogram()
 ```
@@
 )) %>% dplyr::select (`...1`,feature_name,`Gini-importance`, mcc_0.5 ) %>% arrange(-`Gini-importance`)
-vi_data %>% arrange(feature_name)
+vi_data 
 write_tsv(vi_data,out_path)
@@
 
-Confusion matrix
+Performance per sample fraction
 
 ```{r}
-ggplotConfusionMatrix <- function(m){
-  mytitle <- ""#paste("Accuracy", percent_format()(m$overall[1]))
-                  # "Kappa", percent_format()(m$overall[2]))
-  p <-
-    ggplot(data = as.data.frame(m$table) ,
-           aes(x = Reference, y = Prediction)) +
-    geom_tile(aes(fill = log(Freq)), colour = "white") +
-    scale_fill_gradient(low = "white", high = "steelblue") +
-    geom_text(aes(x = Reference, y = Prediction, label = scales::comma(Freq))) +
-    theme(legend.position = "none") +
-    ggtitle(mytitle)+theme(text = element_text(size = 13))
-  return(p)
-}
-conf_data<-iqtree_enriched_test %>% dplyr::select (prob_predictions, true_label) %>% mutate (pred_status = prob_predictions>0.5)
+iqtree_data<- iqtree_all_models_performance %>% filter (dataset=='test', name =='model_standard', analysis_type=='full_standard')  %>% dplyr::select (sample_frac, AUC,logloss) %>% mutate (Program = 'IQTREE')
 
-obs<- factor(conf_data %>% pull (true_label))
-pred<- factor(conf_data %>% pull (pred_status))
+raxml_data<- raxml_all_models_performance %>% filter (dataset=='test', name =='model_standard', analysis_type=='full_standard')  %>% dplyr::select (sample_frac, AUC,logloss) %>% mutate (Program = 'RAxML-NG')
 
-confusion_plot<-ggplotConfusionMatrix(confusionMatrix(data = pred, reference= obs))+theme(legend.position = "none")
+fasttree_data<- fasttree_all_models_performance %>% filter (dataset=='test', name =='model_standard', analysis_type=='full_standard')  %>% dplyr::select (sample_frac, AUC,logloss) %>% mutate (Program = 'FastTree')
 
-error_vs_size<- all_models_performance %>% filter (metric_type=='all_data', dataset =='test', analysis_type=='full_standard')
-error_vs_size_plt = error_vs_size %>% mutate(n_MSAs = sample_frac*2007) %>% ggplot(aes(x=n_MSAs, y = mcc))+geom_line()+ expand_limits(y = 0)+xlab("Number of MSAs")+ theme(text = element_text(size = 13))+ylab("MCC")+geom_point()
+sample_frac_vs_result_plot<-rbind(iqtree_data,raxml_data,fasttree_data) %>% ggplot(aes(x=sample_frac*4652, y = logloss, color = Program))+geom_line()+expand_limits(y = c(0,0))+xlab("Number of training MSAs")+theme_classic()
+sample_frac_vs_result_plot
+```
 
-rocobj <- roc(enriched_test$true_label, enriched_test$prob_predictions)
-auc <- round(auc(enriched_test$true_label, enriched_test$prob_predictions),3)
+Performance per group
 
-global_max_auc_plt<-ggroc(rocobj,colour = 'steelblue', size = 2)+
-  ggtitle(paste0('AUC = ', auc))+theme(text = element_text(size = 13))+xlab("Specificity")
+```{r}
+library(yardstick)
+per_group_data_iqtree<-iqtree_enriched_test%>% dplyr::select (predictions_full_standard, true_binary_support, model_short, feature_msa_pypythia_msa_difficulty,tree_search_model,feature_msa_n_loci, tree_search_model,feature_msa_n_seq) %>% mutate(Pypythia_group = ntile(feature_msa_pypythia_msa_difficulty,n = 20),
+n_seq_group  = ntile(feature_msa_n_seq, 20), n_loci_group = ntile(feature_msa_n_loci,n = 20), true = as.factor(true_binary_support), pred = 1-predictions_full_standard) %>% mutate(Program='IQTREE') %>% group_by(n_seq_group) %>% mutate(median_seq_per_group = median(feature_msa_n_seq)) %>% ungroup() %>% group_by(n_loci_group) %>% mutate(median_loci_per_group = median(feature_msa_n_loci)) %>% ungroup() %>% group_by(Pypythia_group) %>% mutate(median_pythia_per_group = median(feature_msa_pypythia_msa_difficulty)) %>% ungroup()
 
-ggarrange(global_max_auc_plt,confusion_plot,error_vs_size_plt, labels = c("A","B","C"),align = "h", nrow = 2, ncol = 2 )
+per_group_data_fasttree<-fasttree_enriched_test%>% dplyr::select (predictions_full_standard, true_binary_support, model_short, feature_msa_pypythia_msa_difficulty,tree_search_model,feature_msa_n_loci, tree_search_model,feature_msa_n_seq) %>% mutate(Pypythia_group = ntile(feature_msa_pypythia_msa_difficulty,n = 20),
+n_seq_group  = ntile(feature_msa_n_seq, 20), n_loci_group = ntile(feature_msa_n_loci,n = 20), true = as.factor(true_binary_support), pred = 1-predictions_full_standard) %>% mutate(Program='FastTree') %>% group_by(n_seq_group) %>% mutate(median_seq_per_group = median(feature_msa_n_seq)) %>% ungroup() %>% group_by(n_loci_group) %>% mutate(median_loci_per_group = median(feature_msa_n_loci)) %>% ungroup() %>% group_by(Pypythia_group) %>% mutate(median_pythia_per_group = median(feature_msa_pypythia_msa_difficulty)) %>% ungroup()
+
+per_group_data_raxml<-raxml_enriched_test%>% dplyr::select (predictions_full_standard, true_binary_support, model_short, feature_msa_pypythia_msa_difficulty,tree_search_model,feature_msa_n_loci, tree_search_model,feature_msa_n_seq) %>% mutate(Pypythia_group = ntile(feature_msa_pypythia_msa_difficulty,n = 20),
+n_seq_group  = ntile(feature_msa_n_seq, 20), n_loci_group = ntile(feature_msa_n_loci,n = 20), true = as.factor(true_binary_support), pred = 1-predictions_full_standard) %>% mutate(Program='RAxML-NG') %>% group_by(n_seq_group) %>% mutate(median_seq_per_group = median(feature_msa_n_seq)) %>% ungroup() %>% group_by(n_loci_group) %>% mutate(median_loci_per_group = median(feature_msa_n_loci)) %>% ungroup() %>% group_by(Pypythia_group) %>% mutate(median_pythia_per_group = median(feature_msa_pypythia_msa_difficulty)) %>% ungroup()
+
+total_data <- rbind(per_group_data_iqtree, per_group_data_fasttree, per_group_data_raxml)
+
+pythia<- total_data  %>% group_by(median_pythia_per_group, Program) %>% roc_auc( true, pred) %>% mutate(AUC=.estimate) %>% ggplot(aes(x=median_pythia_per_group, y = AUC, color = Program))+expand_limits(y = c(0.5,1))+geom_line()+xlab("MSA difficulty")+ theme_classic()
+loci<-total_data %>% group_by(median_loci_per_group,Program) %>% roc_auc( true, pred) %>% mutate(AUC=.estimate) %>% ggplot(aes(x=median_loci_per_group, y = AUC, color = Program))+expand_limits(y = c(0.5,1))+geom_line()+xlab("Number of positions")+ theme_classic()
+n_seq<-total_data  %>% group_by(median_seq_per_group, Program) %>% roc_auc( true, pred) %>% mutate(AUC=.estimate) %>% ggplot(aes(x=median_seq_per_group, y = AUC, color = Program))+expand_limits(y = c(0.5,1))+geom_line()+xlab("Number of sequences")+ theme_classic()
 
@@
 
-Group performance
+```{r}
+ggarrange(sample_frac_vs_result_plot,pythia,loci,n_seq, labels = c("A","B","C","D"),align = "h", nrow = 2, ncol = 2, common.legend = TRUE )
+```
 
 ```{r}
-test_group_performance<- group_performance %>% filter (dataset=='test')
+cor.test(iqtree_enriched_test$predictions_full_standard, iqtree_enriched_test$feature_msa_pypythia_msa_difficulty)
+cor.test(raxml_enriched_test$predictions_full_standard, raxml_enriched_test$feature_msa_pypythia_msa_difficulty)
+cor.test(fasttree_enriched_test$predictions_full_standard, fasttree_enriched_test$feature_msa_pypythia_msa_difficulty)
 
-Pypythia_error<-test_group_performance %>% filter (grouping_col_name =='msa_difficulty_group') %>%  mutate(grouping_col = factor(grouping_col, levels = c('(0.00328, 0.303]','(0.303, 0.538]','(0.538, 0.571]','	(0.571, 0.865]'))) %>%
-ggplot(aes(y=grouping_col, x=Balanced_Accuracy))+geom_col()+ expand_limits(y = c(0.5,1))+ylab('MSA\ndifficulty')+ xlab('Balanced\naccuracy')+theme(text = element_text(size = 11))
+cor.test(iqtree_enriched_test$predictions_full_standard, iqtree_enriched_test$feature_msa_n_loci)
+cor.test(raxml_enriched_test$predictions_full_standard, raxml_enriched_test$feature_msa_n_loci)
+cor.test(fasttree_enriched_test$predictions_full_standard, fasttree_enriched_test$feature_msa_n_loci)
 
-n_seq_error<-test_group_performance %>% filter (grouping_col_name =='n_seq_group')   %>%  mutate(grouping_col = factor(grouping_col, levels = c('(4.999, 69.0]','(69.0, 129.0]','129.0, 236.0]','	(236.0, 827.0]]'))) %>% ggplot(aes(y=grouping_col, x=Balanced_Accuracy))+geom_col()+ expand_limits(y = c(0.5,1))+ylab('Number\nof sequences')+xlab('Balanced\naccuracy')+theme(text = element_text(size = 11))
-
-n_loci_error<-test_group_performance %>% filter (grouping_col_name =='feature_msa_n_loci') %>%  mutate(grouping_col = factor(grouping_col, levels = c('(61.999, 727.0]','(69.0, 129.0]','129.0, 236.0]','	(236.0, 827.0]]')))  %>% ggplot(aes(y=grouping_col, x=Balanced_Accuracy))+geom_col()+ expand_limits(y = c(0.5,1))+ylab('Number\nof positions')+xlab('Balanced\naccuracy')+theme(text = element_text(size = 11))
-
-Pypythia_scatter<-enriched_test %>% dplyr::select (msa_path,prob_predictions,  feature_msa_pypythia_msa_difficulty) %>% group_by(msa_path,feature_msa_pypythia_msa_difficulty) %>% summarise(mean_prob = mean(prob_predictions)) %>% ggplot(aes(x=feature_msa_pypythia_msa_difficulty, y = mean_prob))+geom_point(size = 0.7)+xlab("MSA\ndifficulty")+ylab("Predicted\nprobability")+theme(text = element_text(size = 11))
-
-seq_scatter<-enriched_test %>% dplyr::select (msa_path,prob_predictions, feature_msa_n_seq, feature_msa_pypythia_msa_difficulty) %>% group_by(msa_path,feature_msa_n_seq) %>% summarise(mean_prob = mean(prob_predictions)) %>% ggplot(aes(x=feature_msa_n_seq, y = mean_prob))+geom_point(size = 0.7)+xlab("Number of\nsequences")+ylab("Predicted\nprobability")+theme(text = element_text(size = 11))
-
-n_loci_scatter<-enriched_test %>% dplyr::select (msa_path,prob_predictions, feature_msa_n_loci) %>% group_by(msa_path,feature_msa_n_loci) %>% summarise(mean_prob = mean(prob_predictions)) %>% ggplot(aes(x=feature_msa_n_loci, y = mean_prob))+geom_point(size = 0.7)+xlab("Number of\n positions")+ylab("Predicted\nprobability")+theme(text = element_text(size = 11))
-
-ggarrange(seq_scatter,Pypythia_scatter,n_loci_scatter, n_seq_error,Pypythia_error, n_loci_error,labels = c("A","C","E","B","D","F"),align = "h", nrow = 3, ncol = 3, legend = "bottom",vjust= 1)
-#error_vs_group_5_5_including_plt
-
-cor.test( enriched_test$feature_msa_pypythia_msa_difficulty,  enriched_test$prob_predictions)
-cor.test( enriched_test$feature_msa_n_seq, enriched_test$prob_predictions)
-cor.test( enriched_test$feature_msa_n_loci, enriched_test$prob_predictions)
+cor.test(iqtree_enriched_test$predictions_full_standard, iqtree_enriched_test$feature_msa_n_seq)
+cor.test(raxml_enriched_test$predictions_full_standard, raxml_enriched_test$feature_msa_n_seq)
+cor.test(fasttree_enriched_test$predictions_full_standard, fasttree_enriched_test$feature_msa_n_seq)
 
@@
 
-Running time
-
 ```{r}
-running_time_data<- enriched_test%>% dplyr::select (boot_run_time,extraction_running_time,nni_ll_eval_running_time) %>% mutate (diff1 = boot_run_time/extraction_running_time, diff2 = boot_run_time/(extraction_running_time+nni_ll_eval_running_time
-)) 
-summary(running_time_data %>% pull (diff1))
-summary(running_time_data %>% pull (diff2))
-summary(running_time_data %>% pull (extraction_running_time))
-summary(running_time_data %>% pull (nni_ll_eval_running_time))
-#enriched_test %>% dplyr::select (boot_run_time, extraction_runni
+iqtree_val
 ```
@@
 
-running_time_data<- enriched_test%>% dplyr::select (program_boot_run_time,extraction_running_time,nni_ll_eval_running_time) %>% mutate (diff1 = program_boot_run_time/extraction_running_time, diff2 = program_boot_run_time/(extraction_running_time+nni_ll_eval_running_time
+running_time_data<- raxml_enriched_test%>% dplyr::select (raxml_enriched_test$time) %>% mutate (diff1 = program_boot_run_time/extraction_running_time, diff2 = program_boot_run_time/(extraction_running_time+nni_ll_eval_running_time
 )) 
@@
 
-```{r}
-summary(validation_data$feature_n_unique_seq)
-raxml_validation_data$
-```
-
 Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

### 2023-11-30 ML_pipeline/ML_argparser.py

@@ -1,0 +1,23 @@
+import argparse
+import os
+
+def main_parser():
+    parser = argparse.ArgumentParser()
+    parser.add_argument('--working_dir', type=str, default=os.getcwd())
+    parser.add_argument('--RFE', action='store_true', default=False)
+    parser.add_argument('--full_grid', action='store_true', default=False)
+    parser.add_argument('--programs', type=str, default='iqtree_raxml_fasttree')
+    parser.add_argument('--cpus_per_main_job', type=int, default=1)
+    parser.add_argument('--sample_fracs', type=str, default='0.25_0.5_1')
+    parser.add_argument('--inc_sample_fracs', action='store_true', default=False)
+    parser.add_argument('--reunite_val_data', action='store_true', default=False)
+    parser.add_argument('--use_val_data', action='store_true', default=False)
+    parser.add_argument('--reunite_training_data', action='store_true', default=False)
+    parser.add_argument('--main_data_folder', type=str,
+                        default='/Users/noa/Workspace/bootstrap_results/remote_results/full_data')
+    parser.add_argument('--validation_data_folder', type=str,
+                        default='/Users/noa/Workspace/bootstrap_results/remote_results/validation_data')
+    parser.add_argument('--n_val_samp', type=int, default=250)
+    parser.add_argument('--n_main_samp', type=int, default=6000)
+    parser.add_argument('--ML_model', action='store_true', default='rf')
+    return parser

### 2023-11-30 ML_pipeline/ML_boot_edit.py

@@ -1,5 +1,5 @@
 import sys
-
+import random
 if sys.platform == "linux" or sys.platform == "linux2":
     PROJECT_ROOT_DIRECRTORY = "/groups/pupko/noaeker/bootstrap_repo"
 else:
@@ -7,14 +7,14 @@ else:
 sys.path.append(PROJECT_ROOT_DIRECRTORY)
 
 from ML_utils.ML_algorithms_and_hueristics import *
+from ML_pipeline.ML_argparser import main_parser
 from side_code.file_handling import create_or_clean_dir, create_dir_if_not_exists
 from side_code.combine_current_results import unify_results_across_jobs
-from side_code.MSA_manipulation import get_MSA_seq_names
-from sklearn.metrics import balanced_accuracy_score
 import os
 import matplotlib.pyplot as plt
 from random import sample
 import argparse
+from functools import reduce
 
 
 def per_tree_analysis(test, features, model):
@@ -66,7 +66,7 @@ def generate_data_dict_from_column_subset(data_dict, column_subset):
     return res
 
 
-def bootstrap_model_pipeline(working_dir, train, test, validation_dict, features, bootstrap_col, groups,
+def bootstrap_model_pipeline(working_dir, train, test, validation_dict,ML_model, features, bootstrap_col, groups,
                              cpus_per_main_job, sample_frac, do_RFE, large_grid, name):
     data_dict = {'test': test, 'train': train}
     data_dict.update(validation_dict)
@@ -75,32 +75,35 @@ def bootstrap_model_pipeline(working_dir, train, test, validation_dict, features
     inc_boot_dict = generate_data_dict_from_column_subset(data_dict, column_subset=[col for col in train.columns if
                                                                                     col in features] + [bootstrap_col])
 
-    model_only_boot = ML_model(only_boot_dict["train"]["X"], groups, only_boot_dict["train"]["y"],
+    model_only_boot = ML_training(only_boot_dict["train"]["X"], groups, only_boot_dict["train"]["y"],
                                n_jobs=cpus_per_main_job,
                                path=os.path.join(working_dir, f'model_only_boot'),
-                               classifier=True, model='lightgbm', calibrate=True, name=name,
+                               classifier=True, model=ML_model, calibrate=True, name=name,
                                large_grid=large_grid, do_RFE=do_RFE,
                                n_cv_folds=3)
-    model_inc_boot = ML_model(inc_boot_dict["train"]["X"], groups, inc_boot_dict["train"]["y"],
[…]

### 2023-11-30 ML_utils/ML_algorithms_and_hueristics.py

@@ -66,7 +66,7 @@ def RFE(model, X, y, group_splitter, n_jobs, scoring, do_RFE):
 
 
 
-def ML_model(X_train, groups, y_train, n_jobs, path, classifier=False, model='lightgbm', calibrate=True, name="", large_grid = False, do_RFE = False, n_cv_folds = 3):
+def ML_training(X_train, groups, y_train, n_jobs, path, classifier=False, model='lightgbm', calibrate=True, name="", large_grid = False, do_RFE = False, n_cv_folds = 3):
     logging.info(f"Building a {name} model and saving to {path}")
     if path and os.path.exists(path):
         logging.info(f"Using existing model in {path}")
@@ -126,16 +126,18 @@ def calibration_plot(model, test_data, y_test):
     pyplot.show()
 
 def enrich_with_single_feature_metrics(var_impt, train_X, y_train, test_X, y_test):
-    mcc_thresholds = [0.5,0.95]
-    for threshold in mcc_thresholds:
         curr_mcc_scores = []
+        curr_auc_scores = []
         for feature in var_impt.index:
-                #lg = LogisticRegression(random_state=0).fit(train_X[[feature]], y_train)
-                lg = lightgbm.LGBMClassifier(importance_type='gain').fit(train_X[[feature]], y_train)
-                pred = lg.predict_proba(test_X[[feature]])[:, 1] >= threshold
-                mcc = matthews_corrcoef(y_test, pred)
-                curr_mcc_scores.append(mcc)
-        var_impt[f"mcc_{threshold}"] = curr_mcc_scores
+            lg = lightgbm.LGBMClassifier(importance_type='gain').fit(train_X[[feature]], y_train)
+            pred_prob= lg.predict_proba(test_X[[feature]])[:, 1]
+            pred = lg.predict_proba(test_X[[feature]])[:, 1] >= 0.5
+            curr_mcc_scores.append(matthews_corrcoef(y_test, pred))
+            auc = roc_auc_score(y_test,pred_prob)
+            curr_auc_scores.append(auc)
+
+        var_impt[f"mcc"] = curr_mcc_scores
+        var_impt[f"auc"] = curr_auc_scores
 
 
 def model_evaluation_metrics(y_true, prob_predictions, estimate_auc = True):
@@ -165,8 +167,8 @@ def model_groups_anlaysis(X,y_true, predictions):
     return performance_per_group
 
 
-def overall_model_performance_analysis(working_dir, model, data_dict, name):
-    if model:
[…]

### 2023-11-30 ML_pipeline/ML_argparser.py
fix defaults

@@ -19,5 +19,5 @@ def main_parser():
                         default='/Users/noa/Workspace/bootstrap_results/remote_results/validation_data')
     parser.add_argument('--n_val_samp', type=int, default=250)
     parser.add_argument('--n_main_samp', type=int, default=6000)
-    parser.add_argument('--ML_model', action='store_true', default='rf')
+    parser.add_argument('--ML_model', action='store_true', default='lightgbm')
     return parser
